### 2019-03-23 response.tex
copying the files from google drive to github repo

\documentclass[11pt]{article}

\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.5in}
\setlength{\evensidemargin}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\topmargin}{0in}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0.1in}

%\usepackage[dvips]{graphicx}

\usepackage{pgfplots,float}
\pgfplotsset{compat=1.11}

\usepackage{amsmath,amssymb,amsthm,epsfig}
\usepackage{enumerate,color,soul}
\def\H2{{{\mathsf H}_2}}
\def\G{{\mathcal G}}
%\input{../vf_def}

\def\serkan#1{\textcolor{blue}{{#1}}}
\def\caleb#1{\textcolor{blue}{{#1}}}
\def\chris#1{\textcolor{blue}{{#1}}}
\def\serkanlast#1{\textcolor{blue}{{#1}}}

\title{Authors' Response to the Referee Comments}
\author{}
\date{}
\begin{document}
\maketitle

We thank the referees {for their detailed reading of the manuscript and their valuable comments}.
 We have incorporated the changes that the referees have suggested into the revised
manuscript.   We list below the referees' comments, followed by our responses, denoted by \serkan{\textsf{AR}}.

\section*{Referee \#1}%change this about referee #1

\begin{itemize}
\item My only concern is the fact that the term H2 is used on time-domain
function, which might be inappropriate from a mathematical point of
view?\\
\serkan{\textsf{AR}: We thank the reviewer for this comment. Since the time domain and frequency domain functions are equivalent, we believe this notation is appropriate. Furthermore, since the infinite horizon, and finite horizon case appear to be connected, the notation $\mathcal{H}_2(t_f)$ illustrates this connection.}
\end{itemize}

\section*{Referee \#2}

\begin{itemize}
\item The introduction of the paper does not allow to position the problem
accurately even though valuable elements of this positioning are given
in section 3.3.

\serkan{\textsf{AR}:   }

\item Some particular application domains are cited in the beginning without
justification. \\
\serkan{\textsf{AR}:   }  
\item What about the choice of the finite horizon tf ?

\serkan{\textsf{AR}: The choice of the finite horizon depends on the problem. We added a clarification for this issue at the end of section 3.2}

\item Justify the choice of simple poles for the reduced order model. 
\serkan{\textsf{AR}:  The choice of simple poles was motivated by the Iterative Rational Krylov Algorithm (IRKA). Since for many applications, the poles of the model are simple, we believe this choice is justified. }

\item What is the reference of the existing result given by Theorem 2.1 ? \\
\serkan{\textsf{AR}:  We have added the citations immediately after the theorem in order to clarify any confusion. }

\item Notation are not fixed in the paper. For example H'(s) is the
derivative of H with respect to s. This can be stated at the end of the
introduction section (with all other notations...)\\
\serkan{\textsf{AR}:  We have added a clarification for this notation immediately after Theorem 2.1, which corresponds to the first occurrence of $\bf{H}'(s)$ .}

\item  Section 2.2 : the 5 introduction lines are very important and can be
used to state the purpose of the paper.  \\
\serkan{\textsf{AR}:  Should we move these lines to the introduction section?}

\item section 3 : the first 5 lines are a "useless repetition" and can be
removed.

\serkan{\textsf{AR}:  Should we remove these lines completely, or argue that it makes the paper easier to read for someone who is not as familiar with the concepts in the paper? }

\item Section 3 : lines 6 to 10 can be given as a Remark (important remark!)

\serkan{\textsf{AR}:  How should we organize (reorganize?) this section? }

\item Remark 3.2 can be given after the proof of the main result and can be
a part of the discussion of the given result. \\
\serkan{\textsf{AR}:  We have moved the remark after the proof. }

\item Lemma 3.3 : "Let G(s) and Gr(s) be as defined in (3.5) and (3.7)"
instead of "Let G(s) and Gr(s) be as defined in (3.5) and (3.5)"

\serkan{\textsf{AR}:  We thank the referee for pointing out this issue. We have addressed it in the revised version.}

\item Page 8 : (3.18) is one equation no need to (3.19).
\serkan{\textsf{AR}:  We thank the referee for pointing out this issue. We have addressed it in the revised version.}

\item The computation of (3.18) is not clear. One can give more details
allowing to obtain such result. \\
\serkan{\textsf{AR}:  What details? Should I include the trivial differentiations?}

\item Sentence after (3.19) : "...in the parentheses in (3.18) and (3.19)"
instead of "...in the parentheses in (3.18) and (3.18)". \\

\serkan{\textsf{AR}:  We thank the referee for pointing out this issue. We have addressed it in the revised version.}

\item 3.3 section can be a (concise) part of the introduction of the paper. \\
\serkan{\textsf{AR}:  We thank the referee for the comment but we respectfully disagree???.}

\item Section 4 : should state how to use the main result in constructing
the reduced order model which is not the case in the current version.\\
\serkan{\textsf{AR}:  We thank the referee for the comment but it is the case in the current version???.}

\item The numerical examples try to show the "supremacy" of the proposed
result which is obviously not the case. Instead, it will be useful to
explain clearly the aim of each numerical experimentation before giving
a clear figure (with only one comparison aim at each time). The given
figures are almost indecipherable !\\

\serkan{\textsf{AR}:  How should we proceed here?.}

\end{itemize}

\section*{Referee \#3}

\begin{itemize}
\item As stated in Remark 3.2, the optimality condition is not directly
expressed by H(s) and Hr(s), but by G(s) and Gr(s), defined by (3.5)
and (3.7).  The reviewer cannot find a physical interpretation of G(s)
and Gr(s).  The authors can comment on the interpretation.\\
\serkan{\textsf{AR}:  We thank the referee for the insightful comment. We have addressed this issue on the revised version.}

\item In Numerical Simulation, the authors can add some time responses of
the original and reduced-order models.	For example, showing impulse
responses of the models enables readers to understand "practical"
importance of the proposed method.\\
\serkan{\textsf{AR}:  What model should we use for this?}

\item  There are some typos in the manuscript.  For example, the statement
"G(s) and Gr(s) be as defined in (3.5) and (3.5)" in Lemma 3.3 should
be corrected. \\
\serkan{\textsf{AR}:  We thank the referee for pointing out this issue. We have addressed it in the revised version.}

\end{itemize}

\end{document}

### 2019-03-23 FH Optimality Conditions Paper.tex
copying the files from google drive to github repo

\documentclass[preprint]{elsarticle}
\usepackage{graphics, euscript, float,blkarray, listings}
\usepackage {graphicx}
\usepackage[top = 1 in, bottom = 1 in, right = 1 in, left = 1 in]{geometry}
\usepackage{tabularx, booktabs, array, float, framed}
\usepackage{color} %red, green, blue, yellow, cyan, magenta, black, white
\usepackage{appendix, setspace}
\usepackage{mathtools}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\definecolor{mygreen}{RGB}{28,172,0} % color values Red, Green, Blue
\definecolor{mylilas}{RGB}{170,55,241}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{mathtools,thmtools,etoolbox}
\usepackage{algorithm,algorithmicx,algpseudocode}
\usepackage{booktabs,subcaption}
\usepackage{tikz,pgfplots}%\pgfplotsset{compat=1.14}
\pgfplotsset{compat=newest}
\usepackage{lineno}
\usepackage{todonotes}
\usepackage[colorlinks,citecolor=black]{hyperref}
\usepackage{cleveref}
\usepackage[scientific-notation=true]{siunitx}
\usepackage{multicol}
\usepackage[font={large}]{caption}
\usepackage{subcaption}
\usepackage{natbib}

%%%%%%%%%%%%%%%%%%%%%%%
%% `Elsevier LaTeX' style

% Theorem definitions
\newtheorem{theorem}{Theorem}[section]
\crefname{theorem}{theorem}{theorems}
\Crefname{theorem}{Theorem}{Theorems}

\newtheorem{prop}[theorem]{Proposition}
\crefname{prop}{proposition}{propositions}
\Crefname{prop}{Proposition}{Propositions}

\newtheorem{lemma}[theorem]{Lemma}
\crefname{lemma}{lemma}{lemmas}
\Crefname{lemma}{Lemma}{Lemmas}

\newtheorem{corollary}[theorem]{Corollary}
\crefname{corollary}{corollary}{corollaries}
\Crefname{corollary}{Corollary}{Corollaries}

%\theoremstyle{remark}
\theoremstyle{definition}
\newtheorem{remark}[theorem]{Remark}
\AtEndEnvironment{remark}{\hfill\ensuremath{\diamondsuit}}
\crefname{remark}{remark}{remarks}
\Crefname{remark}{Remark}{Remarks}

\theoremstyle{definition}
\newtheorem{defn}[theorem]{Definition}
\crefname{defn}{definition}{definitions}
\Crefname{defn}{Definition}{Definitions}

\newtheorem{example}[theorem]{Example}
\AtEndEnvironment{example}{\hfill\ensuremath{\bigcirc}}
\crefname{example}{example}{example}
\Crefname{example}{Example}{Examples}

% End theorem definitions

%% Some bold vectors
\newcommand{\bfell}{{\boldsymbol{\ell}}}
\newcommand{\bfphi}{{\boldsymbol{\phi}}}
\newcommand{\bflambda}{{\boldsymbol{\lambda}}}
\newcommand{\bfu}{{\boldsymbol{u}}}
\newcommand{\bfb}{{\boldsymbol{b}}}
\newcommand{\bfc}{{\boldsymbol{c}}}
\newcommand{\bfp}{{\boldsymbol{p}}}
\newcommand{\bfx}{{\boldsymbol{x}}}
\newcommand{\bfy}{{\boldsymbol{y}}}
\newcommand{\bfr}{{\boldsymbol{r}}}
\newcommand{\bff}{{\boldsymbol{f}}}
\newcommand{\bfe}{{\boldsymbol{e}}}
\newcommand{\bfg}{{\boldsymbol{g}}}
\newcommand{\bfv}{{\boldsymbol{v}}}
\newcommand{\bfw}{{\boldsymbol{w}}}
\newcommand{\bfV}{{\boldsymbol{V}}}
\newcommand{\bfW}{{\boldsymbol{W}}}
\newcommand{\bfF}{{\boldsymbol{F}}}
\newcommand{\bfG}{{\boldsymbol{G}}}
\newcommand{\bfpsi}{{\boldsymbol{\psi}}}
\newcommand{\bfpi}{{\boldsymbol{\pi}}}
\newcommand{\bfbeta}{{\boldsymbol{\beta}}}

%% Makros for figure legends
\newcommand{\original}{Original}
\newcommand{\loewner}{Loewner}
\newcommand{\additional}{Additional points}
\newcommand{\hermite}{Hermite}

%%%% Makros for the paper %%%%%%%%%
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\spann}{span}
\def\numData{n}
\def\leftDir{{\boldsymbol{\ell}}}
\def\LeftDir{\mathcal{L}}
\def\leftPoint{\mu}
\def\LeftPoint{\mathcal{M}}
\def\leftData{{\boldsymbol{f}}}
\def\LeftData{\mathcal{F}}
\def\rightDir{{\boldsymbol{\it r}}}
\def\RightDir{\mathcal{R}}
\def\rightPoint{\sigma}
\def\RightPoint{\mathcal{S}}
\def\rightData{{\boldsymbol{g}}}
\def\RightData{\mathcal{G}}
\def\bitangentialData{\theta}
% Realization and Reduced Matrices
\def\numFunctions{K}
\def\Hred{\smash{\widetilde{H}}}
\def\Ered{\smash{\widetilde{E}}}
\def\Ared {\smash{\widetilde{A}}}
\def\Bred{\smash{\widetilde{B}}}
\def\Cred{\smash{\widetilde{C}}}
\def\Kred{\smash{\widetilde{\mathcal{K}}}}
\def\xred{\widetilde{\boldsymbol{x}}}
\def\yred{\widetilde{\boldsymbol{y}}}
% Full model
\def\dimFOM{N}

\newcommand{\e}[1]{\mathrm{e}^{#1}}
\newcommand{\goyal}{GR}

%%%% END Macros for the paper %%%%%%%%%

\def\mathcal{\EuScript}

\numberwithin{equation}{section}
\newtheorem{Theorem}{Theorem}[section]
\newtheorem{Lemma}[Theorem]{Lemma}
\newtheorem{Definition}[Theorem]{Definition}
\newtheorem{Proposition}[Theorem]{Proposition}
\newcommand{\FH}{FHIRKA}
\newcommand{\R}{\mathbb{R}}
\newcommand{\bh}{\hat{B}}
\newcommand{\hc}{\hat{C}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\A}{\mathbf{A}}
\newcommand{\bpi}{\mathbf{\Pi}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\E}{\mathbf{E}}
\newcommand{\B}{\mathbf{B}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\G}{\mathbf{G}}
\newcommand{\C}{\mathbf{C}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\U}{\mathbf{U}}
\newcommand{\K}{\mathbf{K}}
\newcommand{\T}{\mathbf{T}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\f}{\mathbf{f}}
\newcommand{\uu}{\mathbf{u}}
\newcommand{\HH}{\mathbf{H}}
\newcommand{\PP}{\mathbf{P}}
\newcommand{\Q}{\mathbf{Q}}
\newcommand{\J}{\mathcal{J}}
\newcommand{\h}{\mathbf{h}}
\newcommand{\g}{\mathbf{g}}
\newcommand{\V}{\mathbf{V}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\tf}{t_f}
\newcommand{\rr}{\mathbf{r}}
\newcommand{\lb}{\mathbf{l}}
\newcommand{\rc}{\boldsymbol{r}}
\newcommand{\lc}{\boldsymbol{\ell}}
\newcommand{\cc}{\boldsymbol{c}}
\newcommand{\bb}{\boldsymbol{b}}
\newcommand{\ch}{\mathcal{H}}
\newcommand{\ca}{\mathcal{A}}
\newcommand{\cp}{\mathcal{P}}
\newcommand{\cl}{\mathcal{L}}
\newcommand{\cu}{\mathcal{U}}
\newcommand{\cq}{\mathcal{Q}}
\newcommand{\co}{\mathcal{O}}
\newcommand{\Si}{\mathbf{\Sigma}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcolumntype{x}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
\newcommand{\pder}[2][]{\dfrac{\partial#1}{\partial#2}}
\newcommand{\cbfF}{\mbox{\boldmath${\EuScript{F}}$} }
\newcommand{\cbfH}{\mbox{\boldmath${\EuScript{H}}$} }
\newcommand{\cbfK}{\mbox{\boldmath${\EuScript{K}}$} }
\newcommand{\cbfBsm}{\mbox{{\scriptsize \boldmath${\EuScript{B}}$}}}
\newcommand{\cbfCsm}{\mbox{{\scriptsize \boldmath${\EuScript{C}}$}}}
\newcommand{\cbfD}{\mbox{\boldmath${\EuScript{D}}$}}
\newcommand{\cbfDsm}{\mbox{{\scriptsize \boldmath${\EuScript{D}}$}}}
\newcommand{\cbfW}{\mbox{\boldmath${\EuScript{W}}$} }
\newcommand{\cbfZ}{\mbox{\boldmath${\EuScript{Z}}$} }
\newcommand{\cbfI}{\mbox{${\EuScript{I}}$} }
\newcommand{\cbfA}{\mbox{${\EuScript{A}}$} }
\newcommand{\cbfB}{\mbox{${\EuScript{B}}$}}
\newcommand{\cbfC}{\mbox{${\EuScript{C}}$}}
\newcommand{\cbfG}{\mbox{${\EuScript{G}}$} }
\newcommand{\cbfV}{\mbox{${\EuScript{V}}$} }

\biboptions{numbers,sort&compress}

\begin{document}

\begin{frontmatter}
\title{ {$\ch_2(\tf)$ Optimality Conditions for a Finite-time Horizon\tnoteref{mytitlenote}}}
\author{Klajdi Sinani$^*$}
\ead{klajdi@vt.edu}
\cortext[cor1]{Corresponding author}

\author{Serkan~Gugercin}
\ead{gugercin@vt.edu}

\tnotetext[mytitlenote]{This work was funded by the U.S. National Science Foundation under grants DMS-1522616 and  DMS-1720257.}

\address{Department of Mathematics, Virginia Polytechnic Institute and State University, Blacksburg, VA 24061, USA}

\begin{abstract} 
In this paper we establish the interpolatory model reduction framework for optimal approximation of MIMO dynamical systems with respect to the $\ch_2$ norm over a finite-time horizon, denoted as the $\ch_2(\tf)$ norm.
Using the underlying inner product space, we derive the interpolatory first-order necessary optimality conditions for approximation in the $\ch_2(\tf)$ norm. Then,  we  develop an algorithm, which yields a locally optimal reduced model that satisfies the established interpolation-based optimality conditions. We test the algorithm on various numerical examples to illustrate its performance.
\end{abstract}

\begin{keyword}
 time-limited model reduction \sep interpolation \sep unstable system \sep $\ch_2$-optimality \sep linear systems 
\end{keyword}

\end{frontmatter}
%Introduction
\section{Introduction}
Simulation, design, and control of dynamical systems play an important role in numerous scientific and industrial tasks such as signal propagation in the nervous system, heat dissipation, prediction of major weather events. The need for  detailed  models due to the increasing demand for greater resolution leads to  
large-scale dynamical systems, posing tremendous computational difficulties when applied in numerical simulations. In order to overcome these challenges, we perform  model reduction where we replace the large-scale dynamics with high-fidelity reduced representations.

Consider the linear time-invariant dynamical system:
\begin{equation}  \label{fom1}
\begin{array}{rcl}
\dot{\x}(t) & =& \A\x(t) +\B\uu(t)\\
\y(t)&=&\C\x(t) 
\end{array}
~~\mbox{with}~~ \x(0) = \mathbf{0},
\quad \Longleftrightarrow \quad
\y(t)=\int_{0}^{t}\h (t-\tau)\uu(\tau) d\tau,
\end{equation}
\noindent
where $\A \in \R^{n\times n}$, $\B\in \R^{n\times m}$, and $\C\in \R^{p\times n}$ are constant matrices; 
the variable $\x(t)\in \R^n$ denotes the internal variables, $\uu(t)\in \R^m$ denotes the control inputs, and $\y(t)\in \R^p$ denotes the outputs; and $\h(t)=\C e^{\A t}\B$ is the impulse response of the full model.   The length of the internal variable $\x(t)$, i.e., $n$, is called the order of the full model that we would like to reduce.  Model
reduction achieves this by replacing the original model with a lower dimensional one:

\begin{equation}  \label{rom1}
\begin{array}{rcl}
\dot{\x}_r(t) & =& \A_r\x_r(t) +\B_r\uu(t)\\
\y_r(t)&=&\C_r\x_r(t) 
\end{array}
~~\mbox{with}~~ \x_r(0) = \mathbf{0},
\quad \Longleftrightarrow \quad
\y_r(t)=\int_{0}^{t}\h_r (t-\tau)\uu(\tau) d\tau,
\end{equation}
where as in \eqref{fom1}, $\h_r(t)=\C_r e^{\A_r t}\B_r$ is the impulse response of the reduced model, and $\A_r \in \R^{r\times r}$, $\B_r\in \R^{r\times m}$, and $\C_r\in \R^{p\times r}$ with $r\ll n$. The goal is that the output of the reduced model, $\y_r(t)$, approximates the true output, $\y(t)$, of the original system accurately in an appropriate norm.

For the linear dynamical systems we consider here, a plethora of methods exists
for producing high-fidelity/optimal reduced models, such as  balanced truncation \cite{MulR76, Moo81} and its variants, optimal Hankel norm approximation \cite{Glo84}, and the Iterative Rational Krylov Algorithm (IRKA) \cite{GugBA08} and its variants; see \cite{Ant05,BauBF14} for further references.
These methods usually focus on constructing high-quality reduced models  over an infinite time horizon. However, in various settings, we might either have access to simulations over a finite horizon or can only simulate the system under investigation for a finite horizon such as in the case of unstable dynamical systems. Therefore, in those situations we are interested in the behavior of the dynamical system over a finite time interval $[0,t_f]$ where $t_f < \infty$, and we need the reduced model to be accurate only in the interval of interest.

Time-limited balanced truncation \cite{GawJ90, GugA03, RedK17, Kur17} and Proper Orthogonal Decomposition (POD)   \cite{HolLB96}  are two common frameworks to create reduced models on a finite horizon. 
For time-limited balanced truncation, \cite{GugA03} establishes an upper bound for the $\ch_\infty$ error between the full and reduced models,  \cite{RedK17} provides an $\ch_2$ error bound. In this paper,  we explore $\ch_2(\tf)$ optimality for model reduction over a finite time horizon. 
  We establish $\ch_2(\tf)$ optimality conditions over a finite time horizon and introduce an algorithm which yields better approximations of the large-scale dynamical systems compared to other model reduction methods.

The optimality requires a parametrization of the reduced model. We will work with the time-domain representation of the dynamical system to derive the optimality condition. Specifically, we represent the impulse response of the reduced dynamical system using the modal decomposition, i.e.,
 \begin{equation}\label{kernel}
\begin{aligned}
\h_r(t)=\C_r e^{\A_r t}\B_r=\sum_{i=1}^r e^{\lambda_it}\lc_i\rc_i^T.
\end{aligned}
\end{equation}
where $\lambda_i$'s are the eigenvalues of $\A_r$, and $\lc_i \in \CC^{p\times 1}, \rc_i\in \CC^{m\times1}$. 
In other words, the impulse response is expressed as a sum of $r$ rank-1 $p\times m$  matrices.
To simplify the presentation,  we assume  that $\lambda_i$'s, the reduced order poles,  are simple.  The representation \eqref{kernel} 
is nothing but a state-space transformation on $\h_r(t)=\C_r e^{\A_r t}\B_r$ using the eigenvectors of $\A_r$. 
Using the parametrization of the reduced model in \eqref{kernel}, we derive interpolatory optimality conditions in the $\ch_2(\tf)$ norm and implement a model reduction algorithm that satisfies these optimality conditions.
 The advantages of time-limited model reduction under the $\ch_2(\tf)$ norm are two fold: first, one can obtain a better approximation in the time interval of interest, and, second, one can obtain a locally optimal reduced order model in the case of unstable systems as well.

The rest of the paper is organized as follows: In Section  \ref{sec:intro} we briefly review optimal  $\ch_2$ model reduction in the infinite horizon case. The main results, including the new optimality conditions for finite horizon, are established in Section \ref{sec:main} followed by numerical examples in Section \ref{sec:num}. The papers ends with conclusions and future work in Section \ref{sec:conc}.

%Error Measures
\section{$\ch_2$-Optimal Model Reduction: The Infinite Horizon Case} \label{sec:intro}
Model reduction with respect to the $\ch_2$ norm in the infinite horizon case has been studied extensively; see, for examples, \cite{BarCO91, BryC90, FulO90, MeiL67, Hal92, HylB85, SpaMM92, YanL99, LepMPV91, GugBA08,AniBGA13,Wil70,vuillemin2014poles,CasL18,panzer2013,gerstner2007hom,VanGA08, breiten2013near} and the references therein. 
In this section, we briefly recall these results as they will help to highlight the similarities to and differences from the finite-horizon case that we are interested in.
\subsection{$\ch_2$ Norm and $\ch_2$ Error Measure}
The error analysis for model reduction of linear dynamical systems can be conducted either in the frequency domain or in the time domain. Therefore, we define the $\ch_2$ norm in each domain. 
\begin{Definition}
Let $\h(t)$ and $\g(t)$ be the impulse responses of two asymptotically stable\footnote{We will call
$\h(t) = \C e^{\A t} \B$ asymptotically stable if all the eigenvalues of $\A$ have negative real parts. We will call 
$\h(t)$ stable when $\A$ has some semi-simple eigenvalues on the imaginary axis in addition to those with negative real parts. Otherwise, we call  $\h(t)$ unstable.}
linear dynamical systems with real state-space realizations.
The $\ch_2$ inner product $\langle \cdot, \cdot \rangle_{\ch_2}$ and the $\ch_2$ norm $\| \cdot \|_{\ch_2}$  are 
\begin{align*}
\langle \h, \g \rangle_{\ch_2} =\int_0^\infty \mathsf{Tr}\left((\h(t))^T \g(t)\right) dt \quad \mbox{and}\quad \norm{\h}_{\ch_2}= \sqrt{ \int_0^\infty \norm{\h}_F^2 dt },
\end{align*} 
respectively, where $ \mathsf{Tr}(\cdot)$ denotes the trace and $\norm{\cdot}_F$ denotes the Frobenius norm of a matrix.
\end{Definition}
To define the $\ch_2$ norm in the frequency domain, let $\Y(s)$, $\U(s)$, and $\HH(s)$ denote the Laplace transforms of 
 the output $\y(t)$, the input $\uu(t)$, and the impulse response $\h(t) = \C e^{\A t}\B$
in \eqref{fom1}. Then, taking the Laplace transform of the convolution integral in  \eqref{fom1}, we obtain

\begin{align*}
\Y(s)= \HH(s) \U(s),  \quad \mbox{where} \quad \HH(s) &=\C(s\I-\A)^{-1}\B
\end{align*}
is called the transfer function of \eqref{fom1}. Let $\{\rho_1,\rho_2,\ldots,\rho_n\}$ denote the eigenvalues of $\A$, assumed simple. Then, similar to the parametrization of the reduced model $\h_r(t)$ in \eqref{kernel}, the full-model impulse response can be equivalently written as 
\begin{equation} \label{eqn:hfull}
\h(t) = \sum_{i=1}^n e^{\rho_i t} \cc_i \bb_i^T~~\mbox{with~the~transfer~function}~~
\HH(s) =\sum_{i=1}^n \frac{\cc_i \bb_i^T}{s-\rho_i},
\end{equation}
where $\cc_i \in \CC^{p}$ and $\bb_i \in \CC^m$, for $i=1,\ldots,n$.
This is called the pole-residue form where $\rho_i$'s are the poles of the (rational) transfer  function 
$\HH(s)$ with the corresponding rank-$1$ residues $\cc_i \bb_i^T$.
\begin{Definition}
Let $\HH(s)$ and $\G(s)$ denote the transfer functions  of two asymptotically stable  dynamical systems with real state-space realizations.
The $\ch_2$ inner product $\langle \cdot, \cdot \rangle_{\ch_2}$ 
and the $\ch_2$ norm $\| \cdot \|_{\ch_2}$ are
\begin{align*}
\langle \G, \HH \rangle_{\ch_2}:=\dfrac{1}{2\pi}\int_{-\infty}^\infty \mathsf{Tr}(\overline{\G}(-\imath\omega)\HH^T(\imath\omega)) d\omega \qquad
\mbox{and} \qquad
\norm{ \HH}_{\ch_2}:=\sqrt{\dfrac{1}{2\pi} \int_{-\infty}^\infty \norm{\HH(i\omega)}_F^2 d\omega},
\end{align*}
respectively.
\end{Definition}
Similar to $\HH(s)$, let $\HH_r(s)=\C_r(s\I_r-\A_r)^{-1}\B_r$ denote the transfer function of the reduced model. Then, the relevance and importance of the $\ch_2$ norm in the model reduction become clear by noting that 
\begin{equation}
\begin{aligned}
\norm{\y-\y_r}_{L_\infty}\leq\norm{\HH-\HH_r}_{\ch_2}\norm{\uu}_{L_2},
\end{aligned}
\end{equation}
i.e., the $L_\infty$ norm of the output error $\y(t)-\y_r(t)$ due to model reduction is bounded by the 
the $\ch_2$ norm of error transfer function relative to the $L_2$ norm of the input $\uu(t)$;
see, e.g.,  \cite{AntBG10b}, for a proof.
Therefore, to guarantee that the reduced model output $\y_r(t)$ is close to the original one $\y(t)$, one might look for a reduced model that minimizes the $\ch_2$ error norm.

%Interpolatory Model Reduction
\subsection{Interpolatory Conditions for Optimal $\ch_2$ Model Reduction}
Given a reduced order $r$, the goal is to construct a reduced order model whose transfer function $\HH_r(s)$ minimizes the $\ch_2$ error norm $\norm{\HH - \HH_r}_{\ch_2}$.  Since this is a non-convex optimization problem, the usual, numerically feasible, approach  is to find a reduced model that satisfies the necessary conditions for $\ch_2$ optimality. These conditions can be formulated in terms of Sylvester equations 
\cite{Wil70,HylB85} or interpolation \cite{MeiL67, GugBA08}. These two frameworks are equivalent \cite{GugBA08}. In this paper, we will focus on the interpolation framework.
\begin{theorem} \label{h2thm}%{\text{\cite {MeiL67, GugBA08}}}
Let 
$$
\h_r(t)=\sum_{k=1}^r e^{\lambda_kt}\lc_k\rc_k^T \qquad \Longleftrightarrow \qquad \HH_r(s) = \sum_{k=1}^r \frac{\lc_k\rc_k^T}{s-\lambda_k}
$$
be the best $r^{th}$ order approximation of an asymptotically stable linear dynamical system  $\HH(s)$ with respect to the $\ch_2$ norm. Then, for $k=1, 2, ... , r$,
\begin{align} \label{eqn:h2cond}
\lc_k^T\HH(-\lambda_k)=\lc_k^T\HH_r(-\lambda_k),~~
\HH(-\lambda_k)\rc_k=\HH_r(-\lambda_k)\rc_k,~~\mbox{and},~~
\lc_k^T\HH'(-\lambda_k)\rc_k=\lc_k^T\HH_r'(-\lambda_k)\rc_k.
\end{align}
\end{theorem}
{\color{red}For more details on theorem \ref{h2thm}, see {\cite {MeiL67, GugBA08}}.}
This result states that the transfer function of the optimal $\ch_2$ approximation to $\HH(s)$ is a (tangential) Hermite interpolant where the interpolation points are the mirror images of the reduced-order poles $\{\lambda_k\}$, and the tangental directions are given by the corresponding residues $\{\lc_k\rc_k^T\}$. {\color{red} Throughout this paper, we denote the derivative of $\HH(s)$ by $\HH'(s)$.} Since the optimality conditions depend on the reduced-model to be computed, the solution requires a nonlinear iteration. The Iterative Rational Krylov Algorithm (IRKA)  \cite{GugBA08} and its variants such as \cite{beattie2012realization,gerstner2007hom,CasL18,beattie2009trust} use these interpolation based optimality conditions to produce   an interpolatory, locally $\ch_2$ optimal reduced model. The next section will extend this framework to the finite-time interval case.

%FH model reduction
\section{$\ch_2(\tf)$ Optimal Model Reduction on a Finite Horizon}  \label{sec:main} 
If we are interested  in the behavior of a system only in a finite time interval (or access to a numerical simulation only on a finite time interval), we can obtain better reduced models by defining the measure of approximation quality over the interval of interest, instead of  focussing on the infinite time horizon. Also, in case of an unstable system, since the impulse grows unbounded as $t \to \infty$, the regular $\ch_2$ measure cannot be applied. Therefore, we will focus on model reduction on a finite time horizon.

We would like to point out that for an unstable dynamical system, if the matrix $\A$ does not have purely imaginary eigenvalues, one can work with the
 $\mathcal{L}_2$ norm by decomposing it into a stable and anti-stable system, and then obtain an interpolatory reduced model based on this measure.  However, this solution requires destroying the causality of the underlying dynamics \cite{magruder2010rational}. This is not the framework we are interested in here and we will work with a finite-time interval.

%Error measures in a finite-time horizon
\subsection{$\ch_2(t_f)$ Norm on a Finite-time Horizon}
It is immediately clear from the time-domain definition of the (infinite-horizon) $\ch_2$ norm how to define the finite-horizon version:
\begin{Definition}\label{innerh2}
Let $\h(t)$ and $\g(t)$ denote the impulse responses of two dynamical systems with real state-space realizations.  
For a finite-time horizon 
$[0,t_f]$, the $\ch_2(\tf)$ inner product $\langle \cdot, \cdot \rangle_{\ch_2(\tf)}$ 
and $\ch_2(\tf)$ norm $\norm{\cdot}_{\ch_2(\tf)}$ are defined as  
\begin{align*}
\langle \h, \g \rangle_{\ch_2(\tf)} =\int_0^{\tf} \mathsf{Tr}( (\h(t))^T \g(t)) dt \quad \mbox{and}\quad
\norm{\h}_{\ch_2(\tf)} =\sqrt{\int_0^{\tf} \norm{\h(t)}_F^2 dt},
\end{align*} 
respectively.
\end{Definition}

%Finite horizon pole-residue optimality conditons
\subsection{Finite Horizon Interpolation-based Conditions $\ch_2(\tf)$ Optimal Model Reduction}
The problem we are interested in is as follows: Given a dynamical system with impulse response $\h(t)$ (or equivalently with transfer function $\HH(s)$) and a reduced order $r$, find the reduced model with the impulse response 
\begin{equation}\label{kernel2}
\begin{aligned}
\h_r(t)=\sum_{i=1}^r e^{\lambda_it}\lc_i\rc_i^T.
\end{aligned}
\end{equation}
such that $\norm{\h - \h_r}_{\ch_2(\tf)}$ is minimized. As in the regular $\ch_2$ case, this is a non-convex optimization problem and we will focus on local minimizers. Using 
the parametrization \eqref{kernel2}, we we will derive interpolation-based necessary conditions for optimality.
The main result is given by Theorem \ref{mh2opt}. However, we  need many supplementary results, Lemmas 
\ref{minnerexp}-\ref{mgexp}, to reach this final conclusion.

It is immediately clear that, the $\ch_2(t_f)$-error satisfies 
\begin{equation}\label{merexp}
\begin{aligned}
\norm{\h-\h_r}_{\ch_2({\tf})}^2=\norm{\h}_{\ch_2({\tf})}^2-2 \langle\h, \h_r\rangle_{\ch_2({\tf})}+\norm{\h_r}_{\ch_2({\tf})}^2,
\end{aligned}
\end{equation}
where the inner product  $\langle\h, \h_r\rangle_{\ch_2({\tf})}$ is real since $\h(t)$ and $\h_r(t)$ are real. Finding the first-order necessary conditions for optimal ${\ch_2({\tf})}$ model reduction will require computing the gradient of the error expression \eqref{merexp} with respect to the optimization variables. 
Since the reduced model, as described by  the impulse response in $\h_r(t)$, is parametrized by the 
reduced order poles $\{\lambda_i\}$, and the residue directions $\{\lc_i\}$ and $\{\rc_i\}$, we will be computing the gradient of the error with respect to these variables. Since the first term in the error \eqref{merexp}, i.e., $\norm{\h}_{\ch_2({\tf})}$, is a constant, we will be focusing on the remaining two terms only. First, in the next lemma, we will formulate these two last terms in terms of $\{\lambda_i\}$, 
$\{\lc_i\}$ and $\{\rc_i\}$.
\begin{Lemma}\label{minnerexp}

Let  $\h (t)= \sum_{j=1}^n e^{\rho_jt} \cc_j \bb_j^T $ and $\h_r(t) = \sum_{i=1}^r e^{\lambda_it}\lc_i  \rc_i^T $ be, respectively, the impulse responses of the full and reduced models  as described in \eqref{eqn:hfull} and \eqref{kernel2}.
Then,
\begin{align} \label{eqn:hhr}
\langle\h, \h_r\rangle_{\ch_2({\tf})}=\sum_{j=1}^n\sum_{i=1}^r\lc_{i}^T\cc_j\bb_j^T\rc_{i}\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j}.
\end{align}
and
\begin{align} \label{eqn:hrhr}
\norm{\h_r}_{\ch_2({\tf})}^2= \sum_{i=1}^r\sum_{j=1}^r\lc_{i}^T\lc_{j}\rc_{j}^T\rc_{i}\dfrac{e^{(\lambda_i+\lambda_j){\tf}}-1}{\lambda_i+\lambda_j}.
\end{align}
\end{Lemma}
\begin{proof}
The both results  follow from the definition of the $\ch_2(\tf)$ inner product.  First consider
\begin{align*}
\langle\h, \h_r\rangle_{\ch_2({\tf})}&=\mathsf{Tr}\left(\int_0^{\tf}\h_r(t)^T\h(t)\, dt\right).
\end{align*}
Plug $\h (t)= \sum_{j=1}^n \cc_j \bb_j^T e^{\rho_jt}$ and $\h_r(t) = \sum_{i=1}^r \lc_i  \rc_i^T e^{\lambda_it}$
into this formula to obtain
\begin{align*}
\langle\h, \h_r\rangle_{\ch_2({\tf})}&=\mathsf{Tr} \left(\int_0^{\tf}\sum_{i=1}^r(\lc_{i}\rc_{i}^Te^{\lambda_it})^T\sum_{j=1}^n\cc_j\bb_j^Te^{\rho_jt}dt\right) = \mathsf{Tr} \left(\sum_{i=1}^r 
\sum_{j=1}^n 
\rc_{i}\lc_{i}^T\cc_j \bb_j^T \int_0^{\tf}e^{(\lambda_i + \rho_j )t} dt\right)
\end{align*}
Computing the integral and using the fact that $\mathsf{Tr}(\A_1 \A_2) = \mathsf{Tr}(\A_2 \A_1)$  for two matrices $\A_1$ and $\A_2$ of  appropriate sizes,   we obtain
\begin{align*}
\langle\h, \h_r\rangle_{\ch_2({\tf})}&=\mathsf{Tr}\left(\sum_{j=1}^n\sum_{i=1}^r\rc_{i}\lc_{i}^T\cc_j\bb_j^T\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j}\right)
=\sum_{j=1}^n\sum_{i=1}^r\lc_{i}^T\cc_j\bb_j^T\rc_{i}\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j},
\end{align*}
which proves \eqref{eqn:hhr}. Then, \eqref{eqn:hrhr} follows directly by replacing $\h(t)$ with $\h_r(t)$ in this derivation. 
\end{proof}
For infinite time horizon, Theorem \ref{h2thm} tells us that a locally $\ch_2$ optimal reduced transfer function is a tangential Hermite interpolant of the original transfer function at the mirror images of the reduced poles. We will show that in the finite horizon case, even though Hermite tangential interpolation is still the necessary conditions for optimality,  what is being interpolated and what the interpolant is are different.

\begin{theorem}\label{mh2opt}
Let $\HH(s)= \C(s\I-\A)^{-1} \B$, with $\A \in \R^{n \times n}$, $\B \in \R^{n \times m}$, and $\C \in \R^{p \times n}$,   be the transfer function of the full-order model with  the pole-residue representation 
$
\HH(s) =  \sum_{i=1}^n \frac{\cc_i \bb_i^T}{s-\rho_i}
$
as in \eqref{eqn:hfull}, where $\cc_i \in \CC^{p}$, $\bb_i \in \CC^m$, and $\rho_i \in \CC$ for $i=1,\ldots,n$. 
For a finite-time horizon $[0,t_f]$, define 
\begin{equation} \label{eqn:defG}
\G(s)=-e^{-s\tf}\C(s\I-\A)^{-1}e^{\A \tf}\B+\HH(s).
\end{equation}
Let 
\begin{equation}
\HH_r(s)= \C_r(s\I_r-\A_r)^{-1} \B_r = \sum_{i=1}^r \frac{\lc_i \rc_i^T}{s-\lambda_i} 
\end{equation}
be the transfer function of  the best $r^{th}$ order approximation of $\HH(s)$ with respect to the $\ch_2(\tf)$ norm
where $\A_r \in \R^{r \times r}$, $\B_r \in \R^{r \times m}$, $\C \in \R^{p \times r}$,  $\lc_i \in \CC^{p}$, $\rc_i \in \CC^m$, and $\lambda_i \in \CC$ for $i=1,\ldots,r$.
Define 
\begin{equation} \label{eqn:defGr}
\G_{r}(s)=-e^{-s\tf}\C(s\I_r-\A_r)^{-1}e^{\A_r \tf}\B_r+\HH_r(s).
 \end{equation}
 Then,  for $k=1, 2, ... , r$,
\begin{align}
\lc_k^T\G(-\lambda_k)&=\lc_k^T\G_r(-\lambda_k),  \label{eqn:cond1} \\
\G(-\lambda_k)\rc_k&=\G_r(-\lambda_k)\rc_k,~~\mbox{and}  \label{eqn:cond2} \\
\lc_k^T\G'(-\lambda_k)\rc_k&=\lc_k^T\G_r'(-\lambda_k)\rc_k. \label{eqn:cond3}
\end{align}
\end{theorem}

%Remark 3.2
\begin{remark}
Before we prove Theorem \ref{mh2opt}, we briefly comment on its implications. In the infinite-horizon case, if $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
the interpolant is $\G_r(s)$, which is a function of $\HH_r(s)$, and the interpolated function is $\G(s)$, which is a function of $\HH(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$ in this case. However, the optimal interpolation points  and directions are still determined by the reduced model $\HH_r(s)$. The algorithmic implications of this result will be revisited in Section \ref{sec:discuss}.
\end{remark}
The following lemma will be used in the proof of  Theorem \ref{mh2opt}.
\begin{Lemma}\label{mgexp}
Let $\G(s)$ and $\G_r(s)$ be as defined in \eqref{eqn:defG} and \eqref{eqn:defGr}, respectively. Then,
\begin{align}
\G(-\lambda_k)&=\phantom{-}\sum_{j=1}^n\cc_j\bb_j^T\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j},
\label{eqn:Gatlk}\\
 \G'(-\lambda_k)& =-\sum_{j=1}^n\cc_i\bb_i^T\dfrac {{\tf} (\lambda_k+\rho_j)e^{(\lambda_k+\rho_j){\tf}}+e^{(\lambda_k+\rho_j){\tf}}+1}{(\lambda_k+\rho_j)^2}, \label{eqn:Gpatlk}  \\
 \G_r(-\lambda_k)&=\phantom{-}\sum_{j=1}^n\lc_j\rc_j^T\dfrac{e^{(\lambda_k+\lambda_j){\tf}}-1}{\lambda_k+\lambda_j}, ~~\mbox{and} 
\label{eqn:Gratlk}\\
 \G_r'(-\lambda_k)& =-\sum_{j=1}^n\lc_j\rc_j^T\dfrac {{\tf} (\lambda_k+\lambda_j)e^{(\lambda_k+\lambda_i){\tf}}+e^{(\lambda_k+\lambda_i){\tf}}+1}{(\lambda_k+\lambda_j)^2}. \label{eqn:Grpatlk} 
\end{align}
\end{Lemma}
\begin{proof}
Recall the definition of $\G(s) = -e^{-s\tf}\C(s\I-\A)^{-1}e^{\A \tf}\B+\HH(s)$. Note that we assume that the eigenvalues of $\A$ are simple. Therefore, $e^{\A \tf}$ is also diagonalizable by the eigenvectors of $\A$. Using this fact and the pole-zero residue decomposition of   $\HH(s) = \C(s\I-\A)^{-1} \B = \sum_{j=1}^n \frac{\cc_j \bb_j^T}{s-\rho_j}$, we obtain 
\begin{align}
\G(s) &= -e^{-s\tf}\C(s\I-\A)^{-1}e^{\A \tf}\B+\HH(s)  \nonumber \\ 
&=-e^{-s{\tf}}\sum_{j=1}^n\cc_j\bb_j^T\dfrac{e^{\rho_j\tf}}{s-\rho_j}+\sum_{j=1}^n\cc_j\bb_j^T\dfrac{1}{s-\rho_j}
=\sum_{j=1}^n\cc_j\bb_j^T\dfrac{-e^{(-s + \rho_j){\tf}}}{s-\rho_j}+\sum_{j=1}^n\cc_j\bb_j^T\dfrac{1}{s-\rho_j} \nonumber\\
&=\sum_{j=1}^n\cc_j\bb_j^T\dfrac{e^{(-s+\rho_j)\tf}-1}{-s+\rho_j}. \label{eq:Gs}
\end{align}
Thus,
\begin{align*}
\G(-\lambda_k)=\sum_{j=1}^n\cc_j\bb_j^T\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j},
\end{align*}
which proves \eqref{eqn:Gatlk}. To prove \eqref{eqn:Gpatlk}, we first differentiate \eqref{eq:Gs} with respect to $s$ to obtain
 \begin{align*}
 \G'(s) =\sum_{j=1}^n\cc_j\bb_j^T\dfrac {{\tf} (s-\rho_j)e^{(-s+\rho_j){\tf}}+e^{(-s+\rho_j){\tf}}-1}{(s-\rho_j)^2}
 \end{align*}
 Plugging in $s= -\lambda_k$ in this last expression yields the desired result \eqref{eqn:Gpatlk}. The proofs of
 \eqref{eqn:Gpatlk} and \eqref{eqn:Grpatlk} follow analogously. 
\end{proof}

%%%Start proof here
\emph{{Proof of Theorem \ref{mh2opt}.}}
 Let $\J$ denote the $\ch_2({\tf})$ error norm square, i.e., 
$$ \J=\norm{\h -\h_r}^2_{\ch_2({\tf})} =\norm{\h}_{\ch_2({\tf})}^2-2 \langle\h, \h_r\rangle_{\ch_2({\tf})}+\norm{\h_r}_{\ch_2({\tf})}^2.$$
The expressions for $ \langle\h, \h_r\rangle_{\ch_2({\tf})}$ and $\norm{\h_r}_{\ch_2({\tf})}^2$ in terms of the optimization variables $\{\lambda_k\}$, $\{\rc_k\}$, and $\{\lc_k\}$, for $k=1,2,\ldots,r$, were already derived in Lemma \eqref{minnerexp}. To make the gradient computations with respect to the $k$th parameter more clear, we seperate the $k$th term from these expressions. For example, we write $\langle\h, \h_r\rangle_{\ch_2({\tf})}$ in \eqref{eqn:hhr} as
\begin{equation}
\begin{aligned}
\langle\h, \h_r\rangle_{\ch_2({\tf})}=\sum_{j=1}^n\lc_{k}^T\cc_j\bb_j^T\rc_{k}\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j}
+\sum_{j=1}^n\sum_{\substack{i=1\\i\neq k}}^r\lc_{i}^T\cc_j\bb_j^T\rc_{i}\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j}.
\end{aligned}
\end{equation}
Following the same procedure for $\norm{\h_r}_{\ch_2({\tf})}^2$, we obtain
\begin{equation} \label{merror}
\begin{aligned}
\J= \int_0^{\tf}&\h(t)^T\h(t) dt-2 \left(\sum_{j=1}^n\lc_{k}^T\cc_j\bb_j^T\rc_{k}\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j}
+\sum_{j=1}^n\sum_{\substack{i=1\\i\neq k}}^r\lc_{i}^T\cc_j\bb_j^T\rc_{i}\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j}\right)\\
&+ \lc_{k}^T\lc_{k}\rc_{k}^T\rc_{k}\dfrac{e^{(2\lambda_k){\tf}}-1}{2\lambda_k}
+\sum_{\substack{i=1 \\i \neq k}}^r\lc_{i}^T\lc_{k}\rc_{k}^T\rc_{i}\dfrac{e^{(\lambda_i+\lambda_k){\tf}}-1}{\lambda_i+\lambda_k}\\
&\qquad+\sum_{\substack{j=1\\j\neq k}}^r\lc_{k}^T\lc_{j}\rc_{j}^T\rc_{k}\dfrac{e^{(\lambda_k+\lambda_j){\tf}}-1}{\lambda_i+\lambda_j}+\sum_{\substack{j=1\\j\neq k}}^r\sum_{\substack{i=1\\i\neq k}}^r\lc_{i}^T\lc_{j}\rc_{j}^T\rc_{i}\dfrac{e^{(\lambda_i+\lambda_j){\tf}}-1}{\lambda_i+\lambda_j}.
\end{aligned}
\end{equation}
To compute the gradient of the cost function $\J$ we perturb the cost functional with respect to the residue directions, i.e., $\lc_k\to \lc_k+\Delta\lc_k$ and $\rc_k\to \rc_k+\Delta\rc_k$: 
\begin{align*}
\widetilde{\J}_k&= \int_0^{\tf}\h(t)^T\h(t) dt-2\bigg( \sum_{j=1}^n(\lc_{k}+\Delta\lc_{k})^T\cc_j\bb_j^T(\rc_{k}+\Delta\rc_{k})\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j}\\
&+\sum_{j=1}^n\sum_{\substack{i=1\\i\neq k}}^r\lc_{i}^T\cc_j\bb_j^T\rc_{i}\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j}\bigg)+ (\lc_{k}+\Delta\lc_{k})^T(\lc_{k}+\Delta\lc_{k})(\rc_{k}+\Delta\rc_{k})^T(\rc_{k}+\Delta\rc_{k})\dfrac{e^{(2\lambda_k){\tf}}-1}{2\lambda_k}\\
&+\sum_{\substack{i=1 \\i \neq k}}^r\lc_{i}^T(\lc_{k}+\Delta\lc_{k})(\rc_{k}+\Delta\rc_{k})^T\rc_{i}\dfrac{e^{(\lambda_i+\lambda_k){\tf}}-1}{\lambda_i+\lambda_k}+                             
\sum_{\substack{j=1\\j\neq k}}^r(\lc_{k}+\Delta\lc_{k})^T\lc_{j}\rc_{j}^T(\rc_{k}+\Delta\rc_{k})\dfrac{e^{(\lambda_k+\lambda_j){\tf}}-1}{\lambda_i+\lambda_j}\\
&+\sum_{\substack{j=1\\j\neq k}}^r\sum_{\substack{i=1\\i\neq k}}^r\lc_{i}^T\lc_{j}\rc_{j}^T\rc_{i}\dfrac{e^{(\lambda_i+\lambda_j){\tf}}-1}{\lambda_i+\lambda_j}.
\end{align*}
Then, collecting  the terms that are multiplied by $\Delta\lc_{k}$ and $\Delta\rc_{k}$, we obtain 
\begin{align*}
\nabla_{\rc_{k} }\J &=-2\lc_k^T\sum_{j=1}^n\cc_j\bb_j^T\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j}+2\lc_k^T\sum_{j=1}^n\lc_j\rc_j^T\dfrac{e^{(\lambda_k+\lambda_j){\tf}}-1}{\lambda_k+\lambda_j},~~ \mbox{and}\\
\nabla_{\lc_{k} }\J&=-2\left (\sum_{j=1}^n\cc_j\bb_j^T\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j}\right)\rc_k+2\left(\sum_{j=1}^n\lc_j\rc_j^T\dfrac{e^{(\lambda_k+\lambda_j){\tf}}-1}{\lambda_k+\lambda_j}\right)\rc_k.\\
\end{align*}
Setting $\nabla_{\rc_{k} }\J=0   \text{ and }\nabla_{\lc_{k} }\J=0 $, and using Lemma \ref{mgexp}, mainly
\eqref{eqn:Gatlk} and \eqref{eqn:Gratlk}, yield 
\begin{align*}
\lc_k^T\G(-\lambda_k)=\lc_k^T\G_r(-\lambda_k) \quad \mbox{and}\quad
\G(-\lambda_k)\rc_k=\G_r(-\lambda_k)\rc_k,
\end{align*}
which proves \eqref{eqn:cond1} and \eqref{eqn:cond2}. To prove \eqref{eqn:cond3},  we differentiate $\J$ in \eqref{merror}  with respect to the $k$-th pole $\lambda_k$: 
After some tedious manipulations, we obtain 
\begin{equation}\label{eqn:l1} 
\begin{aligned}
\pder[\J]{\lambda_k} =-2\lc_k^T &\left(\sum_{j=1}^n \cc_j\bb_j^T \dfrac{{\tf}(\lambda_k+\rho_j)e^{(\rho_j+\lambda_k){\tf}}-e^{(\rho_j+\lambda_k){\tf}}+1}{(\lambda_k+\rho_j)^2}\right)\rc_k  \\
&+2\lc_k^T \left(\sum^r_{i=1} \lc_i\rc_i^T\dfrac{{\tf}(\lambda_i+\lambda_j\k)e^{(\lambda_i+\lambda_k){\tf}}-e^{(\lambda_i+\lambda_k){\tf}}+1}{(\lambda_i+\lambda_k)^2}\right)\rc_k.
\end{aligned}
\end{equation}
Lemma \eqref{mgexp}, specifically \eqref{eqn:Gpatlk} and \eqref{eqn:Grpatlk}, show that the expressions in the parentheses in \eqref{eqn:l1}  are, respectively, $-\G'(-\lambda_k)$ and $-\G_r'(-\lambda_k)$. Thus,
 setting $\pder[\J]{\lambda_k}= 0$ yield 
\begin{align*}
\lc_k^T\G'(-\lambda_k)\rc_k&=\lc_k^T\G_r'(-\lambda_k)\rc_k,
\end{align*}
which completes the proof.
\begin{flushright} $\Box$  \end{flushright}
{\color{red} We note that the interval of interest is problem dependent and the choice of the interval i.e. the choice of $t_f$ may depend on the model, however, these optimality conditions hold for any choice of $t_f>0$.}
\begin{remark}
{\color{red} In the infinite-horizon case, if $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
the interpolant is $\G_r(s)$, which is a function of $\HH_r(s)$, and the interpolated function is $\G(s)$, which is a function of $\HH(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$ in this case. However, the optimal interpolation points  and directions are still determined by the reduced model $\HH_r(s)$. The algorithmic implications of this result will be revisited in Section \ref{sec:discuss}.}
\end{remark}

%Proposition 
{\color{red}
The following proposition sheds some light on the physical interpretation of 
\begin{Proposition}
Let  $\G(s)=-e^{-s{\tf}}\C(s\I-\A)^{-1}e^{\A {\tf}}\B+\HH(s)$ where $\HH(s)=\C(s\I-\A)^{-1}\B$  is the transfer function of \eqref{fom1}. Then, $\norm{\G}_{\ch_2({\tf})}=\norm{\HH}_{\ch_2({\tf})}$
\end{Proposition}
\begin{proof}
Note 
\begin{align*}
\G(s)&=-e^{-s{\tf}}\C(s\I-\A)^{-1}e^{\A {\tf}}\B+\HH(s)\\
&=-e^{-s{\tf}}\sum_{i=1}^n \dfrac{\phi_i e^{\lambda_i {\tf}}}{s-\lambda_i}+\HH(s)
\end{align*}
where $\lambda_i$-s and $\phi_i$-s are the poles and the residues of the system, respectively. 
Consider the inverse Laplace transform of $\G(s)$:
\begin{align*}
\mathcal{L}^{-1}\{\G(s)\}&=\h(t)+\sum_{i=1}^n u_{\tf}(t)\phi_i e^{\lambda_i {\tf}}e^{\lambda_i (t-{\tf})}\\
&=\h(t)+\sum_{i=1}^n u_{\tf}(t)\phi_i e^{\lambda_i t}
\end{align*}
where $u_{\tf}(t)$ is the unit step function
\begin{align*}
u_{\tf}(t)=
 \begin{cases} 
      1 &    t\leq {\tf} \\    
      0 &    t > {\tf} 
   \end{cases}.
\end{align*}
Thus, 
\begin{align*}
\norm{\G}_{\ch_2({\tf})}&=\int_0^{\tf}\h(t)+\sum_{i=1}^n u_{\tf}(t)\phi_i e^{\lambda_i t} dt\\
&=\int_0^{\tf}\h(t)dt\\
&=\norm{\HH}_{\ch_2({\tf})}
\end{align*}
\end{proof}
This proposition and its proof suggest that we are matching the models up to $t_f$. For $t>t_f$, we treat the model as insignificant, and the impulse response vanishes after $t_f$.
}

\subsection{Implication of the interpolatory $\ch_2({\tf})$ optimality conditions}
\label{sec:discuss}
Theorem \ref{mh2opt} extends the interpolatory infinite-horizon $\ch_2$ optimality conditions \eqref{eqn:h2cond} to the finite-horizon case. Note that in the case of asymptotically stable dynamical systems, if we let $t_f \to \infty$, we recover the infinite-horizon conditions  \eqref{eqn:h2cond}.

 The major difference from the regular $\ch_2$ problem is that  optimality no longer requires that the reduced model $\HH_r(s)$ tangentially interpolate the full model $\HH(s)$. Instead, the auxiliary reduced-order function $\G_r(s)$ should be a tangential Hermite interpolant to the auxiliary  full-order  function $\G(s)$. However, the optimal interpolation points and the optimal tangential directions still result from the pole-residue representation of the reduced-order transfer function $\HH_r(s)$. This situation is  similar to the interpolatory optimality conditions for the  frequency-weighted $\ch_2$-optimal model reduction problem in which  one tries to minimize a weighted $\ch_2$ norm in the frequency domain, i.e.,
find $\HH_r(s)$ that minimizes $ \norm{ \mathbf{W} ( \HH - \HH_r) }_{\ch_2}$ where $\mathbf{W}(s)$ represents a weighting function in the frequency domain. As \cite{breiten2013near} showed, the optimality  
in the frequency-weighted $\ch_2$-norm requires that a function of $\HH_r(s)$ tangentially interpolate
a function of  $\HH(s)$. Despite this  conceptual similarity, the resulting interpolation conditions are drastically different from what we obtained here as one would expect due to the different measures. For details, we refer the reader to  \cite{breiten2013near}.

As we pointed out in Section \ref{sec:intro}, in addition to the interpolatory framework, one can represent the $\ch_2$ optimality conditions in terms of  Sylvester equations, leading to a projection framework for the reduced model. This means that given the full-model $\HH(s) = \C(s\I - \A)^{-1} \B$, one constructs two bases $\V,\W \in \R^{n\times r}$ with $\V^T\W = \I_r$ such that the reduced-order quantities are obtained via projection, i.e.,
\begin{equation} \label{eqn:proj}
\A_r = \W^T \A \V,~~\B_r = \W^T \B, \quad \mbox{and}\quad \C_r = \C \V.
\end{equation}
In the infinite-horizon case, \cite{Wil70} showed that the optimal $\ch_2$ reduced model is indeed guaranteed to be obtained via projection. Recently, Goyal and Redmann  \cite{GoyR17} have established the Sylvester-equation based optimality conditions for the time-limited $\ch_2$ model reduction problem; i.e., they extended the Wilson framework to the time-limited (finite-horizon) $\ch_2$ problem. Furthermore, they have developed a projection-based IRKA-type numerical algorithm to construct the reduced models. However, as the authors point out in \cite{GoyR17}, even though their algorithm yields high-fidelity reduced models in terms of the $\ch_2(t_f)$ measure, the resulting reduced model  satisfies the optimality conditions only approximately. This is not surprising in light of the optimality conditions we derived here. Since the optimality requires that $\G_r(s)$ should interpolate $\G(s)$ (as opposed to $\HH_r(s)$ interpolating $\HH(s)$), unlike in the infinite-horizon case, the reduced model in the finite-horizon case is \emph{not} necessarily given by a projection  as in \eqref{eqn:proj}. Therefore, a projection-based approach would satisfy the optimality conditions only approximately. This was also the case in  \cite{breiten2013near} where even though a projection-based IRKA-like algorithm produced high-fidelity reduced models in the weighted norm,  it satisfied the optimality conditions approximately.

The advantage of the interpolation framework and the parametrization \eqref{kernel} we consider here is that
we do not require the reduced-model to be obtained via projection. By treating the poles and residues in \eqref{kernel} as the parameters and directly working with them, we can obtain a reduced model to satisfy the optimality conditions exactly. Even though the main focus of this paper is the theoretical interpolatory framework  and a robust numerical algorithm will be  fully considered in a future work, in the next section we will  discuss a basic numerical framework one can develop using the interpolatory conditions.

\begin{remark}
The finite-horizon approximation problem for  \emph{discrete-time dynamical systems} has been considered 
in \cite{MelVG14}. The derivation in \cite{MelVG14}, however, allows the reduced-model quantities to vary at every time- step, thus using a time-varying reduced model as opposed to the time-invariant formulation considered here and in  \cite{GoyR17}. Allowing time-varying quantities drastically simplifies the gradient computations, leading to a recurrence relations for optimality. Therefore, the model reduction problem for finite-horizon $\ch_2$ approximation for \emph{time-invariant} discrete-time dynamical systems is still an open question.
\end{remark}
\section{Numerical considerations}  \label{sec:num}
In this section, we briefly discuss a numerical framework to construct a reduced model that satisfies the optimality conditions \eqref{eqn:cond1}-\eqref{eqn:cond3}. To make the presentation and discussion concise,  we will focus on the single-input/single-output (SISO) version only. The complete numerical framework for the general case  with further details on the underlying optimization schemes will be discussed in a separate work.

\subsection{A descent-type algorithm for the single-input/single-output case}
Let $\HH(s)$ and $\HH_r(s)$ be SISO full- and reduced-model transfer functions, respectively, i.e.,
\begin{equation} \label{eqn:sisoh}
\HH(s) =  \bfc^T(s\I-\A)^{-1}\bfb = \sum_{i=1}^n \frac{\psi_i}{s-\rho_i}\qquad\mbox{and}\qquad
\HH_r(s) =  \bfc_r^T(s\I_r-\A_r)^{-1}\bfb_r = \sum_{i=1}^r \frac{\phi_i}{s-\lambda_i},
\end{equation}
where $\A \in \R^{n\times n}$, $\bfb,\bfc \in \R^n$, $\A_r \in \R^{r\times r}$, and 
$\bfb_r,\bfc_r \in \R^r$. Note that the residues $\psi_i$ and $\phi_i$ are scalar valued.
The following result, which is an immediate consequence of Theorem \ref{mh2opt}, summarizes the optimality conditions for SISO systems.
\begin{corollary}\label{h2opt}
Given the SISO transfer functions $\HH(s)$ and $\HH_r(s)$ as defined in \eqref{eqn:sisoh}, define
$$
\G(s)=-e^{-s{\tf}}\bfc^T(s\I-\A)^{-1}e^{\A {\tf}}\bfb+\HH(s)\quad\mbox{and}\quad\G_r(s)=-e^{-s{\tf}}\bfc^T_r(s\I_r-\A_r)^{-1}e^{\A_r {\tf}}\bfb_r+\HH_r(s).
$$
If $\HH_r$ is the best $r^{th}$ order approximation of $\HH$ with respect to the $\ch_2({\tf})$ norm, then
\begin{align} \label{def:GGrsiso}
\G(-\lambda_k)=\G_r(-\lambda_k),\qquad \mbox{and}\qquad
\G'(-\lambda_k)=\G_r'(-\lambda_k)
\end{align}
where  $\lambda_k$ for $k=1, 2, ... , r$ are the poles of the reduced system $\HH_r(s)$ as given in
 \eqref{eqn:sisoh}.
\end{corollary}
As stated before, the $\ch_2(t_f)$ minimization problem is a non-convex optimization problem and 
Corollary \ref{h2opt} gives the necessary conditions for optimality when both  poles and residues are treated as variables. However, if the poles are fixed, we can establish the necessary and sufficient optimality conditions for the residues and find the global minimizer,  the optimal residues, by solving a linear system.

\begin{corollary}\label{optres}
Let $\HH(s)$ and $\HH_r(s)$ be as given in \eqref{eqn:sisoh}, and $\G(s)$ and $\G_r(s)$ as
in \eqref{def:GGrsiso}.  Assume  the reduced  poles   $\{\lambda_i\}_{i=1}^r$ are fixed. Then,
$\HH_r(s)$ is the best $r^{th}$ order approximation of $\HH(s)$ with respect to the $\ch_2({\tf})$ norm if and only if 
\begin{align*}
\M \bfphi= \mathbf{z}, \quad \mbox{or~equivalently},\quad \G(-\lambda_k)=\G_r(-\lambda_k),\quad\mbox{for}\quad k = 1,2,\ldots, r, 
\end{align*}
where  $\bfphi = [\phi_1~\phi_2~\cdots~\phi_r]^T \in \CC^r$ is the vector of  residues; 
$\mathbf{z} \in \CC^r$ is the vector with  entries 
$$
\mathbf{z}_j=e^{\lambda_j{\tf}}\bfc^T(-\lambda_j\I-\A)^{-1}e^{\A {\tf}}\bfb-\HH(-\lambda_j), \quad \mbox{for}\quad
i=1,2,\ldots,r ;
$$
and 
$\M \in \CC^{r\times r}$  is the  matrix with  entries
$$
\M_{i,j}=\dfrac{e^{(\lambda_i+\lambda_j){\tf}}-1}{\lambda_i+\lambda_j}, \quad \mbox{for}\quad
i,j=1,2,\ldots,r.$$
\end{corollary}
\begin{proof}
Using the SISO counterparts for \eqref{eqn:hrhr} and \eqref{eqn:hhr} and applying some algebraic manipulation, the cost functional $\J$ can be written as
\begin{align*}
\J&= \norm{\h}_{\ch_2(t_f)}^2  - 2\bfphi^{T}\mathbf{w} + \bfphi^{T}\M\bfphi,
\end{align*}
where $\mathbf{w} \in \CC^{r\times 1}$ has the entries $\mathbf{w}_i=\sum_{k=1}^n \psi_k \dfrac{e^{(\rho_k+\lambda_i){\tf}}-1}{\lambda_i+\rho_k}$ for $i=1,2,\ldots,r$. Note that 
$\M$ is positive definite, $\phi^{T}\M\phi=\norm{\h_r}_{\ch_2({\tf})}^2 > 0$, and the cost function is quadratic in $\bfphi$. Thus the (global) minimizer is  obtained by solving $\M \bfphi= \mathbf{z}$, which corresponds to rewriting  $\G(-\lambda_k)=\G_r(-\lambda_k)$ for $k = 1,2,\ldots, r$ in a compact way. 
\end{proof}
The result is analogous to the regular infinite-horizon $\ch_2$ problem where the Lagrange optimality becomes necessary and sufficient once the poles are fixed \cite{gaier1987lectures,beattie2012realization}. What is important here is that once the reduced poles are fixed, the best residues can be computed directly by solving an $r\times r$ linear system $\M \bfphi= \mathbf{z}$.  This is the property that we will exploit in the numerical scheme next.

\subsubsection{FHIRKA: A numerical algorithm for $\ch_2(t_f)$ model reduction}
Here we describe a numerical algorithm which produces a reduced model that satisfies the necessary $\ch_2({\tf})$ optimality conditions upon convergence. Let $\bflambda \in \C^r$ denote the vector of reduced poles. Thus,
the error $\J$ is a function of $\bflambda$ and $\bfphi$. Since we explicitly know the gradients of the cost function
with respect to $\bflambda$ and $\bfphi$ (and indeed we can compute the Hessians as well), one can (locally) minimize $\J$ using well established optimization tools. However, as   Corollary \ref{optres} shows,  we can easily compute the globally optimal $\bfphi$ for fixed  $\bflambda$. Therefore, 
we will treat the reduced  poles $\bflambda$ as the optimization parameter, and once $\bflambda$ are updated at the $kth$ step of an optimization algorithm, we find/update the  corresponding optimal residues $\bfphi$ based on Corollary \ref{optres}, and then repeat the process. Similar strategies have been successfully employed in the regular 
$\ch_2$ optimal approximation problem as well; see \cite{beattie2012realization,beattie2009trust}. In summary,
we use a quasi-Newton type optimization as $\bflambda$ being the parameter and in each optimization  step,
we update the residues,  $\bfphi$, by solving the $r\times r$ linear system $\M \bfphi = \mathbf{z}$  as in Corollary \ref{optres}.  Since we are enforcing interpolation at every step of the algorithm, yet tackling the model reduction problem over a finite horizon, we name this algorithm Finite Horizon IRKA, denoted by \FH.  Unlike regular IRKA,  \FH \ is a descent algorithm, thus indeed mimics \cite{beattie2009trust} more closely. Upon convergence, the locally optimal reduced model satisfies the first-order necessary conditions of 
Corollary \ref{def:GGrsiso}.

\subsection{Numerical Results}  
In this section we compare the proposed algorithm FHIRKA with  Proper Orthogonal Decomposition (POD), Time-Limited Balanced Truncation (TLBT), and the recently introduced $\ch_2(t_f)$-based algorithm  by Goyal and Redmann (\goyal) \cite{GoyR17}, as briefly discussed in Section \ref{sec:discuss}.

We use three models: a heat model of order $n=197$ \cite{Niconet}, a model of  the International Space Station 1R Module (ISS 1R) of order $n=270$ \cite{gugercin2001approximation}, and a toy unstable model of order $n=402$. 
The ISS 1R model has $3$-inputs and $3$-outputs. We  focus on the SISO subsystem from the first-input to the first-output.
We have created the unstable system such that it has $400$ stable poles and $2$ unstable poles (positive real part).

For all three models, we choose $t_f = 1$, first reduce the original model using POD, \goyal \  or TLBT, and then use the resulting reduced model  to initialize \FH. Thus, we are trying to investigate how these different initializations affect the final reduced model via  \FH \ and how much improvement one might expect.
The results are shown in Figures \ref{fig:hm}--\ref{fig:uns},  where we show the $\ch_2(\tf)$ approximation error for different values of $r$, the order of the reduced model.
  All three initializations are used for the heat model (Figure \ref{fig:hm}) where the order is reduced from $r=2$ to $r=10$ with increments of one. For some $r$ values, certain initializations are excluded (e.g., the \goyal \ initialization for $r=6$) since the algorithm either did not converge or produced  poor approximations. However, this happened rarely. For the ISS model (Figure \ref{fig:iss}), we use  TLBT and \goyal \ initializations since  POD produced very poor approximations and thus is excluded. In this case, we reduce the order from $r=2$ to $r=14$ with increments of $2$. For the unstable model (Figure \ref{fig:uns}), we use POD and \goyal \ initializations; for this model our implementation of TLBT produced  poor results and is avoided. In this case,  we reduce the order from $r=2$ to $r=12$ with increments of $2$.

 The first observation is that,  since  \FH \ is a descent-method and drives the initialization to a local minimizer,  it improves the accuracy of  the reduced model for all three initializations as expected. The improvements could be dramatic. For example, 
 \FH \ is able to outperform POD as much as by an order of magnitude, see, for example,
 Figure \ref{fig:hm}, the  $r=4$ and $r=5$ cases. 
 While \FH \ improves TLBT and \goyal\ initialization as well, the improvements for the heat model are not as significant. However, for the ISS model,
\FH \ is able to improve the TLBT performance as much as $50\%$; see, e.g., Figure \ref{fig:iss}, the $r=8$ case. 
The best improvement of the \goyal \ initialization has occurred for the unstable model. For example, for $r=8$, for the unstable model, \FH\ improved the reduced model by more than $40 \%$. Gains were significantly better for POD, especially for larger $r$ values. 
 \begin{figure}[H] 
 \centering
 \begin{subfigure}{.5\textwidth}
 \centering
 \includegraphics [scale=0.4]{allDataHeatF} 
 \caption{\FH \ and other algorithms for a heat model\label{fig:hm}}
 \end{subfigure}%
 \begin{subfigure}{.5\textwidth}
 \centering
 \includegraphics [scale=0.4]{ErrorVSrISSF}
 \caption{\FH \ and other algorithms for the ISS model \label{fig:iss}}
 \end{subfigure}
  \begin{subfigure}{.5\textwidth}
 \centering
   \includegraphics [scale=0.4]{ErrorVSrUNSF}
      \caption{\FH \ and other algorithms for the unstable model \label{fig:uns}}
 \end{subfigure}
 \caption{Numerical results for three  models \label{mainfig}}
 \end{figure}
  Overall,  as expected, \FH \ yields a better approximation compared to the other algorithms for each model. 
 We find that \goyal\  provided the best initialization for \FH. This is not surprising since \goyal\ produces a reduced-model that approximately satisfies the $\ch_2(t_f)$ optimality conditions.

 As we stated above, the numerical issues will be fully investigated  in a future work where we will develop a robust interpolatory $\ch_2(t_f)$-descent algorithm for MIMO systems. We will not only study better initialization techniques in terms of speed and accuracy, but also make the algorithm numerically more efficient by using approximation  techniques for the matrix exponential $e^{\A t_f}$ appearing in the $\ch_2(t_f)$ setting. 
We will also investigate the MIMO version of Corollary \ref{optres}. In the MIMO case, even for fixed poles, one cannot simply find the globally optimal residue directions by solving a linear system, since the problem is no longer quadratic in these variables. In the regular $\ch_2$ case, finding the optimal residue directions for given poles required solving a nonlinear least-squares problem \cite{beattie2012realization}. We anticipate a similar formulation here and will investigate the corresponding numerical implications.

\section{Conclusions and Future Work} \label{sec:conc}
We  established interpolatory $\ch_2({\tf})$-optimality conditions for model reduction of MIMO dynamical systems over a finite horizon. Even though the optimal interpolation points and tangential directions are still determined by the reduced model, we showed that unlike the regular $\ch_2$ problem, a modified reduced-transfer function should interpolate a modified full-order transfer function. For the special case of SISO models, we have studied a numerical algorithm and illustrated that it performs effectively.

As in the regular $\ch_2$-problem, establishing equivalency between the Sylvester-equation based $\ch_2(t_f)$-optimality conditions of \cite{GoyR17} and the interpolation-based conditions we developed here will be an interesting direction to pursue. Furthermore, extensions to bilinear and quadratic-bilinear problems will be crucial.

%\section*{References}
\bibliographystyle{elsarticle-num}

\bibliography{bibliography}
\end{document}

### 2019-03-23 arxivVersion.tex
copying the arxiv version

--- 2019-03-23 FH Optimality Conditions Paper.tex
@@
 \cite{Wil70,HylB85} or interpolation \cite{MeiL67, GugBA08}. These two frameworks are equivalent \cite{GugBA08}. In this paper, we will focus on the interpolation framework.
-\begin{theorem} \label{h2thm}%{\text{\cite {MeiL67, GugBA08}}}
+\begin{theorem} \label{h2thm}
 Let 
@@
 \end{theorem}
-{\color{red}For more details on theorem \ref{h2thm}, see {\cite {MeiL67, GugBA08}}.}
 This result states that the transfer function of the optimal $\ch_2$ approximation to $\HH(s)$ is a (tangential) Hermite interpolant where the interpolation points are the mirror images of the reduced-order poles $\{\lambda_k\}$, and the tangental directions are given by the corresponding residues $\{\lc_k\rc_k^T\}$.
-{\color{red} Throughout this paper, we denote the derivative of $\HH(s)$ by $\HH'(s)$.} Since the optimality conditions depend on the reduced-model to be computed, the solution requires a nonlinear iteration.
+Since the optimality conditions depend on the reduced-model to be computed, the solution requires a nonlinear iteration.
 The Iterative Rational Krylov Algorithm (IRKA)  \cite{GugBA08} and its variants such as \cite{beattie2012realization,gerstner2007hom,CasL18,beattie2009trust} use these interpolation based optimality conditions to produce   an interpolatory, locally $\ch_2$ optimal reduced model.
@@
 \end{theorem}
-
-%Remark 3.2
 \begin{remark}
@@
 \begin{Lemma}\label{mgexp}
-Let $\G(s)$ and $\G_r(s)$ be as defined in \eqref{eqn:defG} and \eqref{eqn:defGr}, respectively. Then,
+Let $\G(s)$ and $\G_r(s)$ be as defined in \eqref{eqn:defG} and \eqref{eqn:defG}, respectively. Then,
 \begin{align}
@@
 After some tedious manipulations, we obtain 
-\begin{equation}\label{eqn:l1} 
-\begin{aligned}
-\pder[\J]{\lambda_k} =-2\lc_k^T &\left(\sum_{j=1}^n \cc_j\bb_j^T \dfrac{{\tf}(\lambda_k+\rho_j)e^{(\rho_j+\lambda_k){\tf}}-e^{(\rho_j+\lambda_k){\tf}}+1}{(\lambda_k+\rho_j)^2}\right)\rc_k  \\
-&+2\lc_k^T \left(\sum^r_{i=1} \lc_i\rc_i^T\dfrac{{\tf}(\lambda_i+\lambda_j\k)e^{(\lambda_i+\lambda_k){\tf}}-e^{(\lambda_i+\lambda_k){\tf}}+1}{(\lambda_i+\lambda_k)^2}\right)\rc_k.
-\end{aligned}
-\end{equation}
-Lemma \eqref{mgexp}, specifically \eqref{eqn:Gpatlk} and \eqref{eqn:Grpatlk}, show that the expressions in the parentheses in \eqref{eqn:l1}  are, respectively, $-\G'(-\lambda_k)$ and $-\G_r'(-\lambda_k)$.
+\begin{align}
+\pder[\J]{\lambda_k} =-2\lc_k^T &\left(\sum_{j=1}^n \cc_j\bb_j^T \dfrac{{\tf}(\lambda_k+\rho_j)e^{(\rho_j+\lambda_k){\tf}}-e^{(\rho_j+\lambda_k){\tf}}+1}{(\lambda_k+\rho_j)^2}\right)\rc_k \label{eqn:l1} \\
+&+2\lc_k^T \left(\sum^r_{i=1} \lc_i\rc_i^T\dfrac{{\tf}(\lambda_i+\lambda_j\k)e^{(\lambda_i+\lambda_k){\tf}}-e^{(\lambda_i+\lambda_k){\tf}}+1}{(\lambda_i+\lambda_k)^2}\right)\rc_k.\label{eqn:l2} 
+\end{align}
+Lemma \eqref{mgexp}, specifically \eqref{eqn:Gpatlk} and \eqref{eqn:Grpatlk}, show that the expressions in the parentheses in \eqref{eqn:l1} and \eqref{eqn:l1} are, respectively, $-\G'(-\lambda_k)$ and $-\G_r'(-\lambda_k)$.
 Thus,
@@
 \begin{flushright} $\Box$  \end{flushright}
-{\color{red} We note that the interval of interest is problem dependent and the choice of the interval i.e.
-the choice of $t_f$ may depend on the model, however, these optimality conditions hold for any choice of $t_f>0$.}
-\begin{remark}
-{\color{red} In the infinite-horizon case, if $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
-the interpolant is $\G_r(s)$, which is a function of $\HH_r(s)$, and the interpolated function is $\G(s)$, which is a function of $\HH(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$ in this case.
-However, the optimal interpolation points  and directions are still determined by the reduced model $\HH_r(s)$.
-The algorithmic implications of this result will be revisited in Section \ref{sec:discuss}.}
-\end{remark}
-
-%Proposition 
-{\color{red}
-The following proposition sheds some light on the physical interpretation of 
-\begin{Proposition}
-Let  $\G(s)=-e^{-s{\tf}}\C(s\I-\A)^{-1}e^{\A {\tf}}\B+\HH(s)$ where $\HH(s)=\C(s\I-\A)^{-1}\B$  is the transfer function of \eqref{fom1}. Then, $\norm{\G}_{\ch_2({\tf})}=\norm{\HH}_{\ch_2({\tf})}$
-\end{Proposition}
-\begin{proof}
-Note 
-\begin{align*}
-\G(s)&=-e^{-s{\tf}}\C(s\I-\A)^{-1}e^{\A {\tf}}\B+\HH(s)\\
-&=-e^{-s{\tf}}\sum_{i=1}^n \dfrac{\phi_i e^{\lambda_i {\tf}}}{s-\lambda_i}+\HH(s)
-\end{align*}
-where $\lambda_i$-s and $\phi_i$-s are the poles and the residues of the system, respectively. 
-Consider the inverse Laplace transform of $\G(s)$:
-\begin{align*}
-\mathcal{L}^{-1}\{\G(s)\}&=\h(t)+\sum_{i=1}^n u_{\tf}(t)\phi_i e^{\lambda_i {\tf}}e^{\lambda_i (t-{\tf})}\\
-&=\h(t)+\sum_{i=1}^n u_{\tf}(t)\phi_i e^{\lambda_i t}
-\end{align*}
-where $u_{\tf}(t)$ is the unit step function
-\begin{align*}
-u_{\tf}(t)=
- \begin{cases} 
-      1 &    t\leq {\tf} \\    
-      0 &    t > {\tf} 
-   \end{cases}.
-\end{align*}
-Thus, 
-\begin{align*}
-\norm{\G}_{\ch_2({\tf})}&=\int_0^{\tf}\h(t)+\sum_{i=1}^n u_{\tf}(t)\phi_i e^{\lambda_i t} dt\\
-&=\int_0^{\tf}\h(t)dt\\
-&=\norm{\HH}_{\ch_2({\tf})}
-\end{align*}
-\end{proof}
-This proposition and its proof suggest that we are matching the models up to $t_f$. For $t>t_f$, we treat the model as insignificant, and the impulse response vanishes after $t_f$.
-}
 

### 2019-03-26 FH Optimality Conditions Paper.tex
physical interpretation

@@
-\documentclass[preprint]{elsarticle}
+%\documentclass[preprint]{elsarticle}
+\documentclass[twocolumn]{autart}  
 \usepackage{graphics, euscript, float,blkarray, listings}
 \usepackage {graphicx}
-\usepackage[top = 1 in, bottom = 1 in, right = 1 in, left = 1 in]{geometry}
+%\usepackage[top = 1 in, bottom = 1 in, right = 1 in, left = 1 in]{geometry}
 \usepackage{tabularx, booktabs, array, float, framed}
@@
 \definecolor{mylilas}{RGB}{170,55,241}
-\usepackage{amsmath,amsfonts,amssymb,amsthm}
+\usepackage{amsmath,amsfonts,amssymb}
 \usepackage{mathtools,thmtools,etoolbox}
@@
 \usepackage[colorlinks,citecolor=black]{hyperref}
-\usepackage{cleveref}
+%\usepackage{cleveref}
 \usepackage[scientific-notation=true]{siunitx}
@@
 \usepackage{subcaption}
-\usepackage{natbib}
+%\usepackage{natbib}
 
@@
 \newtheorem{theorem}{Theorem}[section]
-\crefname{theorem}{theorem}{theorems}
-\Crefname{theorem}{Theorem}{Theorems}
+%\crefname{theorem}{theorem}{theorems}
+%\Crefname{theorem}{Theorem}{Theorems}
 
-\newtheorem{prop}[theorem]{Proposition}
-\crefname{prop}{proposition}{propositions}
-\Crefname{prop}{Proposition}{Propositions}
+%\newtheorem{prop}[theorem]{Proposition}
+%\crefname{prop}{proposition}{propositions}
+%\Crefname{prop}{Proposition}{Propositions}
 
 \newtheorem{lemma}[theorem]{Lemma}
-\crefname{lemma}{lemma}{lemmas}
-\Crefname{lemma}{Lemma}{Lemmas}
+%\crefname{lemma}{lemma}{lemmas}
+%\Crefname{lemma}{Lemma}{Lemmas}
 
 \newtheorem{corollary}[theorem]{Corollary}
-\crefname{corollary}{corollary}{corollaries}
-\Crefname{corollary}{Corollary}{Corollaries}
+%\crefname{corollary}{corollary}{corollaries}
+%\Crefname{corollary}{Corollary}{Corollaries}
 
-%\theoremstyle{remark}
+\theoremstyle{remark}
 \theoremstyle{definition}
@@
 \AtEndEnvironment{remark}{\hfill\ensuremath{\diamondsuit}}
-\crefname{remark}{remark}{remarks}
-\Crefname{remark}{Remark}{Remarks}
-
-\theoremstyle{definition}
-\newtheorem{defn}[theorem]{Definition}
-\crefname{defn}{definition}{definitions}
-\Crefname{defn}{Definition}{Definitions}
+%\crefname{remark}{remark}{remarks}
+%\Crefname{remark}{Remark}{Remarks}
+%
+%\theoremstyle{definition}
+%\newtheorem{defn}[theorem]{Definition}
+%\crefname{defn}{definition}{definitions}
+%\Crefname{defn}{Definition}{Definitions}
 
@@
 \AtEndEnvironment{example}{\hfill\ensuremath{\bigcirc}}
-\crefname{example}{example}{example}
-\Crefname{example}{Example}{Examples}
+%\crefname{example}{example}{example}
+%\Crefname{example}{Example}{Examples}
 
@@
 
-\newcommand{\e}[1]{\mathrm{e}^{#1}}
+%\newcommand{\e}[1]{\mathrm{e}^{#1}}
 \newcommand{\goyal}{GR}
@@
 
-\biboptions{numbers,sort&compress}
+%\biboptions{numbers,sort&compress}
 
@@
 \begin{frontmatter}
-\title{ {$\ch_2(\tf)$ Optimality Conditions for a Finite-time Horizon\tnoteref{mytitlenote}}}
-\author{Klajdi Sinani$^*$}
-\ead{klajdi@vt.edu}
-\cortext[cor1]{Corresponding author}
+\title{ {$\ch_2(\tf)$ Optimality Conditions for a Finite-time Horizon\thanksref{mytitlenote}} }
+\author{Klajdi Sinani\thanksref{cor}} \ead{klajdi@vt.edu},
+\author{Serkan~Gugercin} \ead{gugercin@vt.edu}
 
-\author{Serkan~Gugercin}
-\ead{gugercin@vt.edu}
-
-\tnotetext[mytitlenote]{This work was funded by the U.S. National Science Foundation under grants DMS-1522616 and  DMS-1720257.}
-
+\thanks[mytitlenote]{This work was funded by the U.S. National Science Foundation under grants DMS-1522616 and  DMS-1720257.}
+\thanks[cor]{Corresponding author }
 \address{Department of Mathematics, Virginia Polytechnic Institute and State University, Blacksburg, VA 24061, USA}
@@
 \end{array}
-~~\mbox{with}~~ \x(0) = \mathbf{0},
-\quad \Longleftrightarrow \quad
+%~~\mbox{with}~~ \x(0) = \mathbf{0},
+\; \Leftrightarrow \;
 \y(t)=\int_{0}^{t}\h (t-\tau)\uu(\tau) d\tau,
@@
 \noindent
-where $\A \in \R^{n\times n}$, $\B\in \R^{n\times m}$, and $\C\in \R^{p\times n}$ are constant matrices; 
+where $  \x(0) = \mathbf{0}, \A \in \R^{n\times n}$, $\B\in \R^{n\times m}$, and $\C\in \R^{p\times n}$ are constant matrices; 
 the variable $\x(t)\in \R^n$ denotes the internal variables, $\uu(t)\in \R^m$ denotes the control inputs, and $\y(t)\in \R^p$ denotes the outputs; and $\h(t)=\C e^{\A t}\B$ is the impulse response of the full model.
@@
 reduction achieves this by replacing the original model with a lower dimensional one:
-
 \begin{equation}  \label{rom1}
@@
 \end{array}
-~~\mbox{with}~~ \x_r(0) = \mathbf{0},
-\quad \Longleftrightarrow \quad
+%~~\mbox{with}~~ \x_r(0) = \mathbf{0},
+\; \Leftrightarrow \;
 \y_r(t)=\int_{0}^{t}\h_r (t-\tau)\uu(\tau) d\tau,
 \end{equation}
-where as in \eqref{fom1}, $\h_r(t)=\C_r e^{\A_r t}\B_r$ is the impulse response of the reduced model, and $\A_r \in \R^{r\times r}$, $\B_r\in \R^{r\times m}$, and $\C_r\in \R^{p\times r}$ with $r\ll n$.
+where as in \eqref{fom1}, $\x_r(0) = \mathbf{0}, \h_r(t)=\C_r e^{\A_r t}\B_r$ is the impulse response of the reduced model, and $\A_r \in \R^{r\times r}$, $\B_r\in \R^{r\times m}$, and $\C_r\in \R^{p\times r}$ with $r\ll n$.
 The goal is that the output of the reduced model, $\y_r(t)$, approximates the true output, $\y(t)$, of the original system accurately in an appropriate norm.
@@
 \section{$\ch_2$-Optimal Model Reduction: The Infinite Horizon Case} \label{sec:intro}
-Model reduction with respect to the $\ch_2$ norm in the infinite horizon case has been studied extensively; see, for examples, \cite{BarCO91, BryC90, FulO90, MeiL67, Hal92, HylB85, SpaMM92, YanL99, LepMPV91, GugBA08,AniBGA13,Wil70,vuillemin2014poles,CasL18,panzer2013,gerstner2007hom,VanGA08, breiten2013near} and the references therein.
+Model reduction with respect to the $\ch_2$ norm in the infinite horizon case has been studied extensively; see, for examples, \cite{BarCO91, BryC90 , FulO90, MeiL67, Hal92, HylB85, SpaMM92, YanL99, LepMPV91, GugBA08,AniBGA13,Wil70, vuillemin2014poles,CasL18, panzer2013, gerstner2007hom, VanGA08, breiten2013near} and the references therein.
 In this section, we briefly recall these results as they will help to highlight the similarities to and differences from the finite-horizon case that we are interested in.
@@
 \end{Lemma}
-\begin{proof}
+\begin{pf}
 The both results  follow from the definition of the $\ch_2(\tf)$ inner product.  First consider
@@
 which proves \eqref{eqn:hhr}. Then, \eqref{eqn:hrhr} follows directly by replacing $\h(t)$ with $\h_r(t)$ in this derivation. 
-\end{proof}
+\end{pf}
 For infinite time horizon, Theorem \ref{h2thm} tells us that a locally $\ch_2$ optimal reduced transfer function is a tangential Hermite interpolant of the original transfer function at the mirror images of the reduced poles.
@@
 \end{Lemma}
-\begin{proof}
+\begin{pf}
 Recall the definition of $\G(s) = -e^{-s\tf}\C(s\I-\A)^{-1}e^{\A \tf}\B+\HH(s)$.
@@
  \eqref{eqn:Gpatlk} and \eqref{eqn:Grpatlk} follow analogously. 
-\end{proof}
+\end{pf}
 
@@
 the choice of $t_f$ may depend on the model, however, these optimality conditions hold for any choice of $t_f>0$.}
+
+%Remark
 \begin{remark}
 {\color{red} In the infinite-horizon case, if  $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
-the interpolant is $\G_r(s)$, which is a function of $\HH_r(s)$, and the interpolated function is $\G(s)$, which is a function of $\HH(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$ in this case.
-However, the optimal interpolation points  and directions are still determined by the reduced model $\HH_r(s)$.
-The algorithmic implications of this result will be revisited in Section \ref{sec:discuss}.}
-\end{remark}
-
-%Proposition 
-{\color{red}
-The following proposition sheds some light on the physical interpretation of 
-\begin{Proposition}
-Let  $\G(s)=-e^{-s{\tf}}\C(s\I-\A)^{-1}e^{\A {\tf}}\B+\HH(s)$ where $\HH(s)=\C(s\I-\A)^{-1}\B$  is the transfer function of \eqref{fom1}. Then, $\norm{\G}_{\ch_2({\tf})}=\norm{\HH}_{\ch_2({\tf})}$
-\end{Proposition}
-\begin{proof}
-Note 
-\begin{align*}
-\G(s)&=-e^{-s{\tf}}\C(s\I-\A)^{-1}e^{\A {\tf}}\B+\HH(s)\\
-&=-e^{-s{\tf}}\sum_{i=1}^n \dfrac{\phi_i e^{\lambda_i {\tf}}}{s-\lambda_i}+\HH(s)
-\end{align*}
-where $\lambda_i$-s and $\phi_i$-s are the poles and the residues of the system, respectively. 
-Consider the inverse Laplace transform of $\G(s)$:
+the interpolant is $\G_r(s)$, which is a function of $\HH_r(s)$, and the interpolated function is $\G(s)$, which is a function of $\HH(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$.
+The interpolation of $\G(s)$ guarantees that we are approximating the model on the finite horizon.
+For instance, consider the inverse Laplace transform of $\G(s)$:
 \begin{align*}
@@
 \end{align*}
-Thus, 
-\begin{align*}
-\norm{\G}_{\ch_2({\tf})}&=\int_0^{\tf}\h(t)+\sum_{i=1}^n u_{\tf}(t)\phi_i e^{\lambda_i t} dt\\
-&=\int_0^{\tf}\h(t)dt\\
-&=\norm{\HH}_{\ch_2({\tf})}
-\end{align*}
-\end{proof}
-This proposition and its proof suggest that we are matching the models up to $t_f$. For $t>t_f$, we treat the model as insignificant, and the impulse response vanishes after $t_f$.
-}
+In other words, if we were to use an $\ch_2$ optimal algorithm on $\G(s)$, we would be approximating a finite horizon model by an infinite horizon one.
+For this reason, the interpolation of $\G(s)$ by $\G_r(s)$ is necessary.
+However, the optimal interpolation points  and directions are still determined by the reduced model $\HH_r(s)$.
+The algorithmic implications of this result will be revisited in Section \ref{sec:discuss}.}
+\end{remark}
+
+%%remark
+%\begin{remark}
+%{\color{red} In the infinite-horizon case, if $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
+%the interpolant is $\G_r(s)$, which is a function of $\HH_r(s)$, and the interpolated function is $\G(s)$, which is a function of $\HH(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$ in this case.
+However, the optimal interpolation points  and directions are still determined by the reduced model $\HH_r(s)$.
+The algorithmic implications of this result will be revisited in Section \ref{sec:discuss}.}
+%\end{remark}
+%
+%
+%%Proposition 
+%{\color{red}
+%The following proposition sheds some light on the physical interpretation of 
+%\begin{Proposition}
+%Let  $\G(s)=-e^{-s{\tf}}\C(s\I-\A)^{-1}e^{\A {\tf}}\B+\HH(s)$ where $\HH(s)=\C(s\I-\A)^{-1}\B$  is the transfer function of \eqref{fom1}. Then, $\norm{\G}_{\ch_2({\tf})}=\norm{\HH}_{\ch_2({\tf})}$
+%\end{Proposition}
+%\begin{pf}
+%Note 
+%\begin{align*}
+%\G(s)&=-e^{-s{\tf}}\C(s\I-\A)^{-1}e^{\A {\tf}}\B+\HH(s)\\
+%&=-e^{-s{\tf}}\sum_{i=1}^n \dfrac{\phi_i e^{\lambda_i {\tf}}}{s-\lambda_i}+\HH(s)
+%\end{align*}
+%where $\lambda_i$-s and $\phi_i$-s are the poles and the residues of the system, respectively. 
+%Consider the inverse Laplace transform of $\G(s)$:
+%\begin{align*}
+%\mathcal{L}^{-1}\{\G(s)\}&=\h(t)+\sum_{i=1}^n u_{\tf}(t)\phi_i e^{\lambda_i {\tf}}e^{\lambda_i (t-{\tf})}\\
+%&=\h(t)+\sum_{i=1}^n u_{\tf}(t)\phi_i e^{\lambda_i t}
+%\end{align*}
+%where $u_{\tf}(t)$ is the unit step function
+%\begin{align*}
+%u_{\tf}(t)=
+% \begin{cases} 
+%      1 &    t\leq {\tf} \\    
+%      0 &    t > {\tf} 
+%   \end{cases}.
+%\end{align*}
+%Thus, 
+%\begin{align*}
+%\norm{\G}_{\ch_2({\tf})}&=\int_0^{\tf}\h(t)+\sum_{i=1}^n u_{\tf}(t)\phi_i e^{\lambda_i t} dt\\
+%&=\int_0^{\tf}\h(t)dt\\
+%&=\norm{\HH}_{\ch_2({\tf})}
+%\end{align*}
+%\end{pf}
+%This proposition and its proof suggest that we are matching the models up to $t_f$. For $t>t_f$, we treat the model as insignificant, and the impulse response vanishes after $t_f$.
+%}
+%
 
@@
 \end{corollary}
-\begin{proof}
+\begin{pf}
 Using the SISO counterparts for \eqref{eqn:hrhr} and \eqref{eqn:hhr} and applying some algebraic manipulation, the cost functional $\J$ can be written as
@@
 Thus the (global) minimizer is  obtained by solving $\M \bfphi= \mathbf{z}$, which corresponds to rewriting  $\G(-\lambda_k)=\G_r(-\lambda_k)$ for $k = 1,2,\ldots, r$ in a compact way.
-\end{proof}
+\end{pf}
 The result is analogous to the regular infinite-horizon $\ch_2$ problem where the Lagrange optimality becomes necessary and sufficient once the poles are fixed \cite{gaier1987lectures,beattie2012realization}.
@@
 %\section*{References}
-\bibliographystyle{elsarticle-num}
+%\bibliographystyle{elsarticle-num}
+\bibliographystyle{plain}
 

### 2019-03-26 autosam/autosam.tex
physical interpretation

% autosam.tex
% Annotated sample file for the preparation of LaTeX files
% for the final versions of papers submitted to or accepted for 
% publication in AUTOMATICA.

% See also the Information for Authors.

% Make sure that the zip file that you send contains all the 
% files, including the files for the figures and the bib file.

% Output produced with the elsart style file does not imitate the
% AUTOMATICA style. The style file is generic for all Elsevier
% journals and the output is laid out for easy copy editing. The
% final document is produced from the source file in the
% AUTOMATICA style at Elsevier.

% You may use the style file autart.cls to obtain a two-column 
% document (see below) that more or less imitates the printed 
% Automatica style. This may helpful to improve the formatting 
% of the equations, tables and figures, and also serves to check 
% whether the paper satisfies the length requirements.

% Please note: Authors must not create their own macros.

% For further information regarding the preparation of LaTeX files 
% for Elsevier, please refer to the "Full Instructions to Authors" 
% from Elsevier's anonymous ftp server on ftp.elsevier.nl in the
% directory pub/styles, or from the internet (CTAN sites) on
% ftp.shsu.edu, ftp.dante.de and ftp.tex.ac.uk in the directory
% tex-archive/macros/latex/contrib/supported/elsevier.

%\documentclass{elsart}               % The use of LaTeX2e is preferred.

\documentclass[twocolumn]{autart}    % Enable this line and disable the 
                                     % preceding line to obtain a two-column 
                                     % document whose style resembles the
                                     % printed Automatica style.

\usepackage{graphicx}          % Include this line if your 
                               % document contains figures,
%\usepackage[dvips]{epsfig}    % or this line, depending on which
                               % you prefer.

\begin{document}

\begin{frontmatter}
%\runtitle{Insert a suggested running title}  % Running title for regular 
                                              % papers but only if the title  
                                              % is over 5 words. Running title 
                                              % is not shown in output.

\title{In Catilinam IV\thanksref{footnoteinfo}} % Title, preferably not more 
                                                % than 10 words.

\thanks[footnoteinfo]{Corresponding author M.~T.~Cicero. Tel. +XXXIX-VI-mmmxxi. 
Fax +XXXIX-VI-mmmxxv.}

\author[Paestum]{Marcus Tullius Cicero}\ead{cicero@senate.ir},    % Add the 
\author[Rome]{Julius Caesar}\ead{julius@caesar.ir},               % e-mail address 
\author[Baiae]{Publius Maro Vergilius}\ead{vergilius@culture.ir}  % (ead) as shown

\address[Paestum]{Buckingham Palace, Paestum}  % Please supply                                              
\address[Rome]{Senate House, Rome}             % full addresses
\address[Baiae]{The White House, Baiae}        % here.

\begin{keyword}                           % Five to ten keywords,  
Cicero; Catiline; orations.               % chosen from the IFAC 
\end{keyword}                             % keyword list or with the 
                                          % help of the Automatica 
                                          % keyword wizard

\begin{abstract}                          % Abstract of not more than 200 words.
Cum M.~Cicero consul Nonis Decembribus senatum in aede Iovis 
Statoris consuleret, quid de iis coniurationis Catilinae sociis 
fieri placeret, qui in custodiam traditi essent, factum est, ut 
duae potissimum sententiae proponerentur, una D.~Silani consulis 
designati, qui morte multandos illos censebat, altera C.~Caesaris, 
qui illos publicatis bonis per municipia Italiae distribuendos 
ac vinculis sempiternis tenendos existimabat.
\end{abstract}

\end{frontmatter}

\section{Introduction}
Video, patres conscripti, in me omnium vestrum ora atque oculos esse 
conversos, video vos non solunn de vestro ac rei publicae, verum 
etiam, si id depulsum sit, de meo periculo esse sollicitos. Est mihi 
iucunda in malis et grata in dolore vestra erga me voluntas, sed eam, 
per deos inmortales, deponite atque obliti salutis meae de vobis ac 
de vestris liberis cogitate. Mihi si haec condicio consulatus data 
est, ut omnis acerbitates, onunis dolores cruciatusque perferrem, 
feram non solum fortiter, verum etiam lubenter, dum modo meis 
laboribus vobis populoque Romano dignitas salusque pariatur.

\begin{figure}
\begin{center}
\includegraphics[height=4cm]{jcaesar.eps}    % The printed column  
\caption{Gaius Julius Caesar, 100--44 B.C.}  % width is 8.4 cm.
\label{fig1}                                 % Size the figures 
\end{center}                                 % accordingly.
\end{figure}

% OR

%\begin{figure}
%\begin{center}
%\epsfig{file=jcaesar,width=7cm}
%\caption{Gaius Julius Caesar, 100--44 B.C.}
%\label{fig1}
%\end{center}
%\end{figure}

\subsection{A subsection}
Marcus Tullius Cicero, 106--43 B.C. was a Roman statesman, orator, 
and philosopher.  A major figure in the last years of the Republic, 
he is best known for his orations against Catiline\footnote{
This footnote should be very brief.}
and for his mastery of Latin prose \cite{Heritage:92}. He was a 
contemporary of Julius Caesar (Fig.~\ref{fig1}).

\section{The argument}
Some words might be appropriate describing equation~(\ref{e1}), if 
we had but time and space enough.
\begin{equation} \label{e1}
{{\partial F}\over {\partial t}} =
D{{\partial^2 F}\over {\partial x^2}}.
\end{equation}
See \cite{Abl:56}, \cite{AbTaRu:54}, \cite{Keo:58} and 
\cite{Pow:85}.
This equation goes far beyond the celebrated theorem ascribed to the great
Pythagoras by his followers.
\begin{thm}
The square of the length of the hypotenuse of a right triangle equals the sum of the squares 
of the lengths of the other two sides.
\end{thm}
\section{Epilogue}
A word or two to conclude, and this even includes some inline 
maths:  $R(x,t)\sim t^{-\beta}g(x/t^\alpha)\exp(-|x|/t^\alpha)$.

\begin{ack}                               % Place acknowledgements
Partially supported by the Roman Senate.  % here.
\end{ack}

\bibliographystyle{plain}        % Include this if you use bibtex 
\bibliography{autosam}           % and a bib file to produce the 
                                 % bibliography (preferred). The
                                 % correct style is generated by
                                 % Elsevier at the time of printing.

%\begin{thebibliography}{99}     % Otherwise use the  
                                 % thebibliography environment.
                                 % Insert the full references here.
                                 % See a recent issue of Automatica 
                                 % for the style.
%  \bibitem[Heritage, 1992]{Heritage:92}
%     (1992) {\it The American Heritage. 
%     Dictionary of the American Language.}
%     Houghton Mifflin Company.
%  \bibitem[Able, 1956]{Abl:56}
%     B.~C.~Able (1956). Nucleic acid content of macroscope. 
%     {\it Nature 2}, 7--9. 
%  \bibitem[Able {\em et al.}, 1954]{AbTaRu:54}   
%     B.~C. Able, R.~A. Tagg, and M.~Rush (1954).
%     Enzyme-catalyzed cellular transanimations.
%     In A.~F.~Round, editor, 
%     {\it Advances in Enzymology Vol. 2} (125--247). 
%     New York, Academic Press.
%  \bibitem[R.~Keohane, 1958]{Keo:58}
%     R.~Keohane (1958).
%     {\it Power and Interdependence: 
%     World Politics in Transition.}
%     Boston, Little, Brown \& Co.
%  \bibitem[Powers, 1985]{Pow:85}
%     T.~Powers (1985).
%     Is there a way out?
%     {\it Harpers, June 1985}, 35--47.

%\end{thebibliography}

\appendix
\section{A summary of Latin grammar}    % Each appendix must have a short title.
\section{Some Latin vocabulary}         % Sections and subsections are supported  
                                        % in the appendices.
\end{document}

### 2019-03-26 FH Optimality Conditions Paper.tex
most recent changes

@@
 \section{$\ch_2$-Optimal Model Reduction: The Infinite Horizon Case} \label{sec:intro}
-Model reduction with respect to the $\ch_2$ norm in the infinite horizon case has been studied extensively; see, for examples, \cite{BarCO91, BryC90 , FulO90, MeiL67, Hal92, HylB85, SpaMM92, YanL99, LepMPV91, GugBA08,AniBGA13,Wil70, vuillemin2014poles,CasL18, panzer2013, gerstner2007hom, VanGA08, breiten2013near} and the references therein.
+Model reduction with respect to the $\ch_2$ norm in the infinite horizon case has been studied extensively; see, for examples, \cite{BarCO91,BryC90, FulO90, MeiL67, Hal92, HylB85, SpaMM92, YanL99, LepMPV91, GugBA08,AniBGA13,Wil70, vuillemin2014poles,CasL18, panzer2013, gerstner2007hom, VanGA08, breiten2013near} and the references therein.
 In this section, we briefly recall these results as they will help to highlight the similarities to and differences from the finite-horizon case that we are interested in.
+
 \subsection{$\ch_2$ Norm and $\ch_2$ Error Measure}
@@
 \begin{align*}
-\langle \h, \g \rangle_{\ch_2} =\int_0^\infty \mathsf{Tr}\left((\h(t))^T \g(t)\right) dt \quad \mbox{and}\quad \norm{\h}_{\ch_2}= \sqrt{ \int_0^\infty \norm{\h}_F^2 dt },
+\langle \h, \g \rangle_{\ch_2} &=\int_0^\infty \mathsf{Tr}\left((\h(t))^T \g(t)\right) dt,\\
+% \quad \mbox{and}\quad 
+\norm{\h}_{\ch_2}&= \sqrt{ \int_0^\infty \norm{\h}_F^2 dt },
 \end{align*} 
@@
 
-\begin{align*}
-\Y(s)= \HH(s) \U(s),  \quad \mbox{where} \quad \HH(s) &=\C(s\I-\A)^{-1}\B
-\end{align*}
+$\Y(s)= \HH(s) \U(s),  \quad \mbox{where} \quad \HH(s) =\C(s\I-\A)^{-1}\B$
+
 is called the transfer function of \eqref{fom1}.
@@
 \begin{equation} \label{eqn:hfull}
-\h(t) = \sum_{i=1}^n e^{\rho_i t} \cc_i \bb_i^T~~\mbox{with~the~transfer~function}~~
+\h(t) = \sum_{i=1}^n e^{\rho_i t} \cc_i \bb_i^T~~\mbox{with}~~
 \HH(s) =\sum_{i=1}^n \frac{\cc_i \bb_i^T}{s-\rho_i},
@@
 \begin{align*}
-\langle \G, \HH \rangle_{\ch_2}:=\dfrac{1}{2\pi}\int_{-\infty}^\infty \mathsf{Tr}(\overline{\G}(-\imath\omega)\HH^T(\imath\omega)) d\omega \qquad
-\mbox{and} \qquad
-\norm{ \HH}_{\ch_2}:=\sqrt{\dfrac{1}{2\pi} \int_{-\infty}^\infty \norm{\HH(i\omega)}_F^2 d\omega},
+\langle \G, \HH \rangle_{\ch_2}&:=\dfrac{1}{2\pi}\int_{-\infty}^\infty \mathsf{Tr}(\overline{\G}(-\imath\omega)\HH^T(\imath\omega)) d\omega \\
+%\qquad \mbox{and} \qquad\\
+\norm{ \HH}_{\ch_2}&:=\sqrt{\dfrac{1}{2\pi} \int_{-\infty}^\infty \norm{\HH(i\omega)}_F^2 d\omega},
 \end{align*}
@@
 \end{Definition}
+
 Similar to $\HH(s)$, let $\HH_r(s)=\C_r(s\I_r-\A_r)^{-1}\B_r$ denote the transfer function of the reduced model.
@@
 In this case,  we reduce the order from $r=2$ to $r=12$ with increments of $2$.
-
  The first observation is that,  since  \FH \ is a descent-method and drives the initialization to a local minimizer,  it improves the accuracy of  the reduced model for all three initializations as expected.
@@
  \caption{\FH \ and other algorithms for a heat model\label{fig:hm}}
- \end{subfigure}%
+ \end{subfigure}
  \begin{subfigure}{.5\textwidth}

### 2019-04-03 FH Optimality Conditions Paper.tex
fixed equations to fit in double column format

@@
 \documentclass[twocolumn]{autart}  
-\usepackage{graphics, euscript, float,blkarray, listings}
-\usepackage {graphicx}
-%\usepackage[top = 1 in, bottom = 1 in, right = 1 in, left = 1 in]{geometry}
-\usepackage{tabularx, booktabs, array, float, framed}
-\usepackage{color} %red, green, blue, yellow, cyan, magenta, black, white
-\usepackage{appendix, setspace}
-\usepackage{mathtools}
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\definecolor{mygreen}{RGB}{28,172,0} % color values Red, Green, Blue
-\definecolor{mylilas}{RGB}{170,55,241}
-\usepackage{amsmath,amsfonts,amssymb}
-\usepackage{mathtools,thmtools,etoolbox}
-\usepackage{algorithm,algorithmicx,algpseudocode}
-\usepackage{booktabs,subcaption}
-\usepackage{tikz,pgfplots}%\pgfplotsset{compat=1.14}
-\pgfplotsset{compat=newest}
-\usepackage{lineno}
-\usepackage{todonotes}
-\usepackage[colorlinks,citecolor=black]{hyperref}
-%\usepackage{cleveref}
-\usepackage[scientific-notation=true]{siunitx}
-\usepackage{multicol}
-\usepackage[font={large}]{caption}
-\usepackage{subcaption}
-%\usepackage{natbib}
-
-%%%%%%%%%%%%%%%%%%%%%%%
-%% `Elsevier LaTeX' style
-
-% Theorem definitions
-\newtheorem{theorem}{Theorem}[section]
-%\crefname{theorem}{theorem}{theorems}
-%\Crefname{theorem}{Theorem}{Theorems}
-
-%\newtheorem{prop}[theorem]{Proposition}
-%\crefname{prop}{proposition}{propositions}
-%\Crefname{prop}{Proposition}{Propositions}
-
-\newtheorem{lemma}[theorem]{Lemma}
-%\crefname{lemma}{lemma}{lemmas}
-%\Crefname{lemma}{Lemma}{Lemmas}
-
-\newtheorem{corollary}[theorem]{Corollary}
-%\crefname{corollary}{corollary}{corollaries}
-%\Crefname{corollary}{Corollary}{Corollaries}
-
-\theoremstyle{remark}
-\theoremstyle{definition}
-\newtheorem{remark}[theorem]{Remark}
-\AtEndEnvironment{remark}{\hfill\ensuremath{\diamondsuit}}
-%\crefname{remark}{remark}{remarks}
-%\Crefname{remark}{Remark}{Remarks}
-%
-%\theoremstyle{definition}
-%\newtheorem{defn}[theorem]{Definition}
-%\crefname{defn}{definition}{definitions}
-%\Crefname{defn}{Definition}{Definitions}
-
-\newtheorem{example}[theorem]{Example}
-\AtEndEnvironment{example}{\hfill\ensuremath{\bigcirc}}
-%\crefname{example}{example}{example}
-%\Crefname{example}{Example}{Examples}
-
-% End theorem definitions
-
-%% Some bold vectors
-\newcommand{\bfell}{{\boldsymbol{\ell}}}
-\newcommand{\bfphi}{{\boldsymbol{\phi}}}
-\newcommand{\bflambda}{{\boldsymbol{\lambda}}}
-\newcommand{\bfu}{{\boldsymbol{u}}}
-\newcommand{\bfb}{{\boldsymbol{b}}}
-\newcommand{\bfc}{{\boldsymbol{c}}}
-\newcommand{\bfp}{{\boldsymbol{p}}}
-\newcommand{\bfx}{{\boldsymbol{x}}}
-\newcommand{\bfy}{{\boldsymbol{y}}}
-\newcommand{\bfr}{{\boldsymbol{r}}}
-\newcommand{\bff}{{\boldsymbol{f}}}
-\newcommand{\bfe}{{\boldsymbol{e}}}
-\newcommand{\bfg}{{\boldsymbol{g}}}
-\newcommand{\bfv}{{\boldsymbol{v}}}
-\newcommand{\bfw}{{\boldsymbol{w}}}
-\newcommand{\bfV}{{\boldsymbol{V}}}
-\newcommand{\bfW}{{\boldsymbol{W}}}
-\newcommand{\bfF}{{\boldsymbol{F}}}
-\newcommand{\bfG}{{\boldsymbol{G}}}
-\newcommand{\bfpsi}{{\boldsymbol{\psi}}}
-\newcommand{\bfpi}{{\boldsymbol{\pi}}}
-\newcommand{\bfbeta}{{\boldsymbol{\beta}}}
-
-%% Makros for figure legends
-\newcommand{\original}{Original}
-\newcommand{\loewner}{Loewner}
-\newcommand{\additional}{Additional points}
-\newcommand{\hermite}{Hermite}
-
-%%%% Makros for the paper %%%%%%%%%
-\DeclareMathOperator{\rank}{rank}
-\DeclareMathOperator{\diag}{diag}
-\DeclareMathOperator{\spann}{span}
-\def\numData{n}
-\def\leftDir{{\boldsymbol{\ell}}}
-\def\LeftDir{\mathcal{L}}
-\def\leftPoint{\mu}
-\def\LeftPoint{\mathcal{M}}
-\def\leftData{{\boldsymbol{f}}}
-\def\LeftData{\mathcal{F}}
-\def\rightDir{{\boldsymbol{\it r}}}
-\def\RightDir{\mathcal{R}}
-\def\rightPoint{\sigma}
-\def\RightPoint{\mathcal{S}}
-\def\rightData{{\boldsymbol{g}}}
-\def\RightData{\mathcal{G}}
-\def\bitangentialData{\theta}
-% Realization and Reduced Matrices
-\def\numFunctions{K}
-\def\Hred{\smash{\widetilde{H}}}
-\def\Ered{\smash{\widetilde{E}}}
-\def\Ared {\smash{\widetilde{A}}}
-\def\Bred{\smash{\widetilde{B}}}
-\def\Cred{\smash{\widetilde{C}}}
-\def\Kred{\smash{\widetilde{\mathcal{K}}}}
-\def\xred{\widetilde{\boldsymbol{x}}}
-\def\yred{\widetilde{\boldsymbol{y}}}
-% Full model
-\def\dimFOM{N}
-
-%\newcommand{\e}[1]{\mathrm{e}^{#1}}
-\newcommand{\goyal}{GR}
-
-%%%% END Macros for the paper %%%%%%%%%
-
-\def\mathcal{\EuScript}
-
-\numberwithin{equation}{section}
-\newtheorem{Theorem}{Theorem}[section]
-\newtheorem{Lemma}[Theorem]{Lemma}
-\newtheorem{Definition}[Theorem]{Definition}
-\newtheorem{Proposition}[Theorem]{Proposition}
-\newcommand{\FH}{FHIRKA}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\bh}{\hat{B}}
-\newcommand{\hc}{\hat{C}}
-\newcommand{\CC}{\mathbb{C}}
-\newcommand{\x}{\mathbf{x}}
-\newcommand{\A}{\mathbf{A}}
-\newcommand{\bpi}{\mathbf{\Pi}}
-\newcommand{\Y}{\mathbf{Y}}
-\newcommand{\E}{\mathbf{E}}
-\newcommand{\B}{\mathbf{B}}
-\newcommand{\I}{\mathbf{I}}
-\newcommand{\G}{\mathbf{G}}
-\newcommand{\C}{\mathbf{C}}
-\newcommand{\D}{\mathbf{D}}
-\newcommand{\M}{\mathbf{M}}
-\newcommand{\X}{\mathbf{X}}
-\newcommand{\U}{\mathbf{U}}
-\newcommand{\K}{\mathbf{K}}
-\newcommand{\T}{\mathbf{T}}
-\newcommand{\y}{\mathbf{y}}
-\newcommand{\f}{\mathbf{f}}
-\newcommand{\uu}{\mathbf{u}}
-\newcommand{\HH}{\mathbf{H}}
-\newcommand{\PP}{\mathbf{P}}
-\newcommand{\Q}{\mathbf{Q}}
-\newcommand{\J}{\mathcal{J}}
-\newcommand{\h}{\mathbf{h}}
-\newcommand{\g}{\mathbf{g}}
-\newcommand{\V}{\mathbf{V}}
-\newcommand{\W}{\mathbf{W}}
-\newcommand{\tf}{t_f}
-\newcommand{\rr}{\mathbf{r}}
-\newcommand{\lb}{\mathbf{l}}
-\newcommand{\rc}{\boldsymbol{r}}
-\newcommand{\lc}{\boldsymbol{\ell}}
-\newcommand{\cc}{\boldsymbol{c}}
-\newcommand{\bb}{\boldsymbol{b}}
-\newcommand{\ch}{\mathcal{H}}
-\newcommand{\ca}{\mathcal{A}}
-\newcommand{\cp}{\mathcal{P}}
-\newcommand{\cl}{\mathcal{L}}
-\newcommand{\cu}{\mathcal{U}}
-\newcommand{\cq}{\mathcal{Q}}
-\newcommand{\co}{\mathcal{O}}
-\newcommand{\Si}{\mathbf{\Sigma}}
-\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
-\newcolumntype{x}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
-\newcommand{\pder}[2][]{\dfrac{\partial#1}{\partial#2}}
-\newcommand{\cbfF}{\mbox{\boldmath${\EuScript{F}}$} }
-\newcommand{\cbfH}{\mbox{\boldmath${\EuScript{H}}$} }
-\newcommand{\cbfK}{\mbox{\boldmath${\EuScript{K}}$} }
-\newcommand{\cbfBsm}{\mbox{{\scriptsize \boldmath${\EuScript{B}}$}}}
-\newcommand{\cbfCsm}{\mbox{{\scriptsize \boldmath${\EuScript{C}}$}}}
-\newcommand{\cbfD}{\mbox{\boldmath${\EuScript{D}}$}}
-\newcommand{\cbfDsm}{\mbox{{\scriptsize \boldmath${\EuScript{D}}$}}}
-\newcommand{\cbfW}{\mbox{\boldmath${\EuScript{W}}$} }
-\newcommand{\cbfZ}{\mbox{\boldmath${\EuScript{Z}}$} }
-\newcommand{\cbfI}{\mbox{${\EuScript{I}}$} }
-\newcommand{\cbfA}{\mbox{${\EuScript{A}}$} }
-\newcommand{\cbfB}{\mbox{${\EuScript{B}}$}}
-\newcommand{\cbfC}{\mbox{${\EuScript{C}}$}}
-\newcommand{\cbfG}{\mbox{${\EuScript{G}}$} }
-\newcommand{\cbfV}{\mbox{${\EuScript{V}}$} }
-
-%\biboptions{numbers,sort&compress}
-
+\input{paperPreamble}
+\setlength{\parskip}{0pt}
+\setlength{\parsep}{0pt}
+\setlength{\headsep}{0pt}
+\setlength{\topskip}{0pt}
+\setlength{\topmargin}{0pt}
+\setlength{\topsep}{0pt}
+\setlength{\partopsep}{0pt}
 \begin{document}
@@
 \section{Introduction}
-Simulation, design, and control of dynamical systems play an important role in numerous scientific and industrial tasks such as signal propagation in the nervous system, heat dissipation, prediction of major weather events.
+Simulation, design, and control of dynamical systems play an important role in numerous scientific and industrial tasks such as signal propagation in the nervous system\cite{Steve Cox}, heat dissipation, prediction of major weather events \cite{BoyP11,Sha03,Ant05}.
 The need for  detailed  models due to the increasing demand for greater resolution leads to
@@
 \begin{equation}  \label{rom1}
-\begin{array}{rcl}
-\dot{\x}_r(t) & =& \A_r\x_r(t) +\B_r\uu(t)\\
-\y_r(t)&=&\C_r\x_r(t) 
-\end{array}
+%\begin{array}{rcl}
+\begin{aligned}
+\dot{\x}_r(t)&=\A_r\x_r(t)+\B_r\uu(t)\\
+\y_r(t)&=\C_r\x_r(t) 
+\end{aligned}
+%\end{array}
 %~~\mbox{with}~~ \x_r(0) = \mathbf{0},
 \; \Leftrightarrow \;
-\y_r(t)=\int_{0}^{t}\h_r (t-\tau)\uu(\tau) d\tau,
+\y_r(t)=\int_{0}^{t}\h_r (t-\tau)\uu(\tau)d\tau,
 \end{equation}
@@
 For the linear dynamical systems we consider here, a plethora of methods exists
-for producing high-fidelity/optimal reduced models, such as  balanced truncation \cite{MulR76, Moo81} and its variants, optimal Hankel norm approximation \cite{Glo84}, and the Iterative Rational Krylov Algorithm (IRKA) \cite{GugBA08} and its variants; see \cite{Ant05,BauBF14} for further references.
+for producing high-fidelity/optimal reduced models, such as  balanced truncation \cite{MulR76,Moo81} and its variants, optimal Hankel norm approximation \cite{Glo84}, and the Iterative Rational Krylov Algorithm (IRKA) \cite{GugBA08} and its variants; see \cite{Ant05,BauBF14} for further references.
 These methods usually focus on constructing high-quality reduced models  over an infinite time horizon.
@@
 
-Time-limited balanced truncation \cite{GawJ90, GugA03, RedK17, Kur17} and Proper Orthogonal Decomposition (POD)   \cite{HolLB96}  are two common frameworks to create reduced models on a finite horizon.
+Time-limited balanced truncation \cite{GawJ90,GugA03,RedK17,Kur17} and Proper Orthogonal Decomposition (POD)   \cite{HolLB96}  are two common frameworks to create reduced models on a finite horizon. 
 For time-limited balanced truncation, \cite{GugA03} establishes an upper bound for the $\ch_\infty$ error between the full and reduced models,  \cite{RedK17} provides an $\ch_2$ error bound.
@@
 \section{$\ch_2$-Optimal Model Reduction: The Infinite Horizon Case} \label{sec:intro}
-Model reduction with respect to the $\ch_2$ norm in the infinite horizon case has been studied extensively; see, for examples, \cite{BarCO91,BryC90, FulO90, MeiL67, Hal92, HylB85, SpaMM92, YanL99, LepMPV91, GugBA08,AniBGA13,Wil70, vuillemin2014poles,CasL18, panzer2013, gerstner2007hom, VanGA08, breiten2013near} and the references therein.
+Model reduction with respect to the $\ch_2$ norm in the infinite horizon case has been studied extensively; see, for examples, \cite{BarCO91,BryC90,FulO90,MeiL67,Hal92,HylB85,SpaMM92,YanL99,LepMPV91,GugBA08,AniBGA13,Wil70,vuillemin2014poles,CasL18,panzer2013,gerstner2007hom,VanGA08,breiten2013near} and the references therein.
 In this section, we briefly recall these results as they will help to highlight the similarities to and differences from the finite-horizon case that we are interested in.
@@
 Since this is a non-convex optimization problem, the usual, numerically feasible, approach  is to find a reduced model that satisfies the necessary conditions for $\ch_2$ optimality.
-These conditions can be formulated in terms of Sylvester equations
-\cite{Wil70,HylB85} or interpolation \cite{MeiL67, GugBA08}. These two frameworks are equivalent \cite{GugBA08}. In this paper, we will focus on the interpolation framework.
+These conditions can be formulated in terms of Sylvester equations \cite{Wil70,HylB85} or interpolation \cite{MeiL67,GugBA08}.
+These two frameworks are equivalent \cite{GugBA08}.
+In this paper, we will focus on the interpolation framework.
 \begin{theorem} \label{h2thm}%{\text{\cite {MeiL67, GugBA08}}}
@@
 be the best $r^{th}$ order approximation of an asymptotically stable linear dynamical system  $\HH(s)$ with respect to the $\ch_2$ norm. Then, for $k=1, 2, ... , r$,
-\begin{align} \label{eqn:h2cond}
-\lc_k^T\HH(-\lambda_k)=\lc_k^T\HH_r(-\lambda_k),~~
-\HH(-\lambda_k)\rc_k=\HH_r(-\lambda_k)\rc_k,~~\mbox{and},~~
-\lc_k^T\HH'(-\lambda_k)\rc_k=\lc_k^T\HH_r'(-\lambda_k)\rc_k.
-\end{align}
+\begin{equation} \label{eqn:h2cond}
+\begin{aligned}
+\lc_k^T\HH(-\lambda_k)&=\lc_k^T\HH_r(-\lambda_k),\\~~
+\HH(-\lambda_k)\rc_k&=\HH_r(-\lambda_k)\rc_k,\\
+\lc_k^T\HH'(-\lambda_k)\rc_k&=\lc_k^T\HH_r'(-\lambda_k)\rc_k.
+\end{aligned}
+\end{equation}
 \end{theorem}
-{\color{red}For more details on theorem \ref{h2thm}, see {\cite {MeiL67, GugBA08}}.}
+{\color{red}For more details on theorem \ref{h2thm}, see {\cite {GugBA08}}.}
 This result states that the transfer function of the optimal $\ch_2$ approximation to $\HH(s)$ is a (tangential) Hermite interpolant where the interpolation points are the mirror images of the reduced-order poles $\{\lambda_k\}$, and the tangental directions are given by the corresponding residues $\{\lc_k\rc_k^T\}$.
@@
 \begin{align*}
-\langle \h, \g \rangle_{\ch_2(\tf)} =\int_0^{\tf} \mathsf{Tr}( (\h(t))^T \g(t)) dt \quad \mbox{and}\quad
-\norm{\h}_{\ch_2(\tf)} =\sqrt{\int_0^{\tf} \norm{\h(t)}_F^2 dt},
+\langle \h, \g \rangle_{\ch_2(\tf)} &=\int_0^{\tf} \mathsf{Tr}( (\h(t))^T \g(t)) dt, \\%\quad \mbox{and}\quad
+\norm{\h}_{\ch_2(\tf)} &=\sqrt{\int_0^{\tf} \norm{\h(t)}_F^2 dt}.
 \end{align*}
-respectively.
+
 \end{Definition}
@@
 \begin{aligned}
+%\scalemath{0.9}
+{
 \norm{\h-\h_r}_{\ch_2({\tf})}^2=\norm{\h}_{\ch_2({\tf})}^2-2 \langle\h, \h_r\rangle_{\ch_2({\tf})}+\norm{\h_r}_{\ch_2({\tf})}^2,
+}
 \end{aligned}
@@
 \begin{align} \label{eqn:hhr}
-\langle\h, \h_r\rangle_{\ch_2({\tf})}=\sum_{j=1}^n\sum_{i=1}^r\lc_{i}^T\cc_j\bb_j^T\rc_{i}\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j}.
+\langle\h, \h_r\rangle_{\ch_2({\tf})}=\sum_{j=1}^n\sum_{i=1}^r\lc_{i}^T\cc_j\bb_j^T\rc_{i}\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j},
 \end{align}
@@
 \begin{align*}
-\langle\h, \h_r\rangle_{\ch_2({\tf})}&=\mathsf{Tr} \left(\int_0^{\tf}\sum_{i=1}^r(\lc_{i}\rc_{i}^Te^{\lambda_it})^T\sum_{j=1}^n\cc_j\bb_j^Te^{\rho_jt}dt\right) = \mathsf{Tr} \left(\sum_{i=1}^r 
+%\begin{dmath*}
+\langle\h, \h_r\rangle_{\ch_2({\tf})}&=\mathsf{Tr} \left(\int_0^{\tf}\sum_{i=1}^r(\lc_{i}\rc_{i}^Te^{\lambda_it})^T\sum_{j=1}^n\cc_j\bb_j^Te^{\rho_jt}dt\right)\\
+ &= \mathsf{Tr} \left(\sum_{i=1}^r 
 \sum_{j=1}^n 
 \rc_{i}\lc_{i}^T\cc_j \bb_j^T \int_0^{\tf}e^{(\lambda_i + \rho_j )t} dt\right)
+%\end{dmath*}
 \end{align*}
@@
 \begin{align*}
-\langle\h, \h_r\rangle_{\ch_2({\tf})}&=\mathsf{Tr}\left(\sum_{j=1}^n\sum_{i=1}^r\rc_{i}\lc_{i}^T\cc_j\bb_j^T\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j}\right)
-=\sum_{j=1}^n\sum_{i=1}^r\lc_{i}^T\cc_j\bb_j^T\rc_{i}\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j},
+\langle\h, \h_r\rangle_{\ch_2({\tf})}&=\mathsf{Tr}\left(\sum_{j=1}^n\sum_{i=1}^r\rc_{i}\lc_{i}^T\cc_j\bb_j^T\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j}\right)\\
+&=\sum_{j=1}^n\sum_{i=1}^r\lc_{i}^T\cc_j\bb_j^T\rc_{i}\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j},
 \end{align*}
@@
 
-%Remark 3.2
-\begin{remark}
-Before we prove Theorem \ref{mh2opt}, we briefly comment on its implications.
+%%Remark 3.2
+%\begin{remark}
+%Before we prove Theorem \ref{mh2opt}, we briefly comment on its implications.
 In the infinite-horizon case, if $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$.
 However, in the finite-horizon case,
-the interpolant is $\G_r(s)$, which is a function of $\HH_r(s)$, and the interpolated function is $\G(s)$, which is a function of $\HH(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$ in this case.
+%the interpolant is $\G_r(s)$, which is a function of $\HH_r(s)$, and the interpolated function is $\G(s)$, which is a function of $\HH(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$ in this case.
 However, the optimal interpolation points  and directions are still determined by the reduced model $\HH_r(s)$.
 The algorithmic implications of this result will be revisited in Section \ref{sec:discuss}.
-\end{remark}
-The following lemma will be used in the proof of  Theorem \ref{mh2opt}.
+%\end{remark}
+The next lemma will be used in proving Theorem \ref{mh2opt}.
 \begin{Lemma}\label{mgexp}
@@
 \label{eqn:Gatlk}\\
- \G'(-\lambda_k)& =-\sum_{j=1}^n\cc_i\bb_i^T\dfrac {{\tf} (\lambda_k+\rho_j)e^{(\lambda_k+\rho_j){\tf}}+e^{(\lambda_k+\rho_j){\tf}}+1}{(\lambda_k+\rho_j)^2}, \label{eqn:Gpatlk}  \\
+ \G'(-\lambda_k)& =-\sum_{j=1}^n\cc_i\bb_i^T\dfrac {({\tf} (\lambda_k+\rho_j)-1)e^{(\lambda_k+\rho_j){\tf}}+1}{(\lambda_k+\rho_j)^2}, \label{eqn:Gpatlk}  \\
  \G_r(-\lambda_k)&=\phantom{-}\sum_{j=1}^n\lc_j\rc_j^T\dfrac{e^{(\lambda_k+\lambda_j){\tf}}-1}{\lambda_k+\lambda_j}, ~~\mbox{and} 
 \label{eqn:Gratlk}\\
- \G_r'(-\lambda_k)& =-\sum_{j=1}^n\lc_j\rc_j^T\dfrac {{\tf} (\lambda_k+\lambda_j)e^{(\lambda_k+\lambda_i){\tf}}+e^{(\lambda_k+\lambda_i){\tf}}+1}{(\lambda_k+\lambda_j)^2}. \label{eqn:Grpatlk} 
+ \G_r'(-\lambda_k)& =-\sum_{j=1}^n\lc_j\rc_j^T\dfrac {({\tf} (\lambda_k+\lambda_j)-1)e^{(\lambda_k+\lambda_i){\tf}}+1}{(\lambda_k+\lambda_j)^2}. \label{eqn:Grpatlk} 
 \end{align}
@@
 \G(s) &= -e^{-s\tf}\C(s\I-\A)^{-1}e^{\A \tf}\B+\HH(s)  \nonumber \\ 
-&=-e^{-s{\tf}}\sum_{j=1}^n\cc_j\bb_j^T\dfrac{e^{\rho_j\tf}}{s-\rho_j}+\sum_{j=1}^n\cc_j\bb_j^T\dfrac{1}{s-\rho_j}
-=\sum_{j=1}^n\cc_j\bb_j^T\dfrac{-e^{(-s + \rho_j){\tf}}}{s-\rho_j}+\sum_{j=1}^n\cc_j\bb_j^T\dfrac{1}{s-\rho_j} \nonumber\\
+&=-e^{-s{\tf}}\sum_{j=1}^n\cc_j\bb_j^T\dfrac{e^{\rho_j\tf}}{s-\rho_j}+\sum_{j=1}^n\cc_j\bb_j^T\dfrac{1}{s-\rho_j}\\
+&=\sum_{j=1}^n\cc_j\bb_j^T\dfrac{-e^{(-s + \rho_j){\tf}}}{s-\rho_j}+\sum_{j=1}^n\cc_j\bb_j^T\dfrac{1}{s-\rho_j} \nonumber\\
 &=\sum_{j=1}^n\cc_j\bb_j^T\dfrac{e^{(-s+\rho_j)\tf}-1}{-s+\rho_j}. \label{eq:Gs}
@@
 Thus,
-\begin{align*}
-\G(-\lambda_k)=\sum_{j=1}^n\cc_j\bb_j^T\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j},
-\end{align*}
+$\G(-\lambda_k)=\sum_{j=1}^n\cc_j\bb_j^T\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j}$,
 which proves \eqref{eqn:Gatlk}. To prove \eqref{eqn:Gpatlk}, we first differentiate \eqref{eq:Gs} with respect to $s$ to obtain
@@
  \eqref{eqn:Gpatlk} and \eqref{eqn:Grpatlk} follow analogously. 
-\end{pf}
-
-%%%Start proof here
+\end{pf}%%%Start proof here
 \emph{{Proof of Theorem \ref{mh2opt}.}}
  Let $\J$ denote the $\ch_2({\tf})$ error norm square, i.e., 
-$$ \J=\norm{\h -\h_r}^2_{\ch_2({\tf})} =\norm{\h}_{\ch_2({\tf})}^2-2 \langle\h, \h_r\rangle_{\ch_2({\tf})}+\norm{\h_r}_{\ch_2({\tf})}^2.$$
+\begin{align*}
+ \J&=\norm{\h -\h_r}^2_{\ch_2({\tf})} \\
+&=\norm{\h}_{\ch_2({\tf})}^2-2 \langle\h, \h_r\rangle_{\ch_2({\tf})}+\norm{\h_r}_{\ch_2({\tf})}^2.
+\end{align*}
 The expressions for $ \langle\h, \h_r\rangle_{\ch_2({\tf})}$ and $\norm{\h_r}_{\ch_2({\tf})}^2$ in terms of the optimization variables $\{\lambda_k\}$, $\{\rc_k\}$, and $\{\lc_k\}$, for $k=1,2,\ldots,r$, were already derived in Lemma \eqref{minnerexp}.
@@
 For example, we write $\langle\h, \h_r\rangle_{\ch_2({\tf})}$ in \eqref{eqn:hhr} as
-\begin{equation}
-\begin{aligned}
+\begin{dmath*}
 \langle\h, \h_r\rangle_{\ch_2({\tf})}=\sum_{j=1}^n\lc_{k}^T\cc_j\bb_j^T\rc_{k}\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j}
 +\sum_{j=1}^n\sum_{\substack{i=1\\i\neq k}}^r\lc_{i}^T\cc_j\bb_j^T\rc_{i}\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j}.
-\end{aligned}
-\end{equation}
+\end{dmath*}
 Following the same procedure for $\norm{\h_r}_{\ch_2({\tf})}^2$, we obtain
-\begin{equation} \label{merror}
-\begin{aligned}
-\J= \int_0^{\tf}&\h(t)^T\h(t) dt-2 \left(\sum_{j=1}^n\lc_{k}^T\cc_j\bb_j^T\rc_{k}\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j}
-+\sum_{j=1}^n\sum_{\substack{i=1\\i\neq k}}^r\lc_{i}^T\cc_j\bb_j^T\rc_{i}\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j}\right)\\
-&+ \lc_{k}^T\lc_{k}\rc_{k}^T\rc_{k}\dfrac{e^{(2\lambda_k){\tf}}-1}{2\lambda_k}
-+\sum_{\substack{i=1 \\i \neq k}}^r\lc_{i}^T\lc_{k}\rc_{k}^T\rc_{i}\dfrac{e^{(\lambda_i+\lambda_k){\tf}}-1}{\lambda_i+\lambda_k}\\
-&\qquad+\sum_{\substack{j=1\\j\neq k}}^r\lc_{k}^T\lc_{j}\rc_{j}^T\rc_{k}\dfrac{e^{(\lambda_k+\lambda_j){\tf}}-1}{\lambda_i+\lambda_j}+\sum_{\substack{j=1\\j\neq k}}^r\sum_{\substack{i=1\\i\neq k}}^r\lc_{i}^T\lc_{j}\rc_{j}^T\rc_{i}\dfrac{e^{(\lambda_i+\lambda_j){\tf}}-1}{\lambda_i+\lambda_j}.
-\end{aligned}
-\end{equation}
+\begin{dmath} \label{merror}
+\J= \int_0^{\tf}\h(t)^T\h(t) dt-2 \left(\sum_{j=1}^n\lc_{k}^T\cc_j\bb_j^T\rc_{k}\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j}
++\sum_{j=1}^n\sum_{\substack{i=1\\i\neq k}}^r\lc_{i}^T\cc_j\bb_j^T\rc_{i}\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j}\right) %\\
++ \lc_{k}^T\lc_{k}\rc_{k}^T\rc_{k}\dfrac{e^{(2\lambda_k){\tf}}-1}{2\lambda_k}
++\sum_{\substack{i=1 \\i \neq k}}^r\lc_{i}^T\lc_{k}\rc_{k}^T\rc_{i}\dfrac{e^{(\lambda_i+\lambda_k){\tf}}-1}{\lambda_i+\lambda_k}%\\
++\sum_{\substack{j=1\\j\neq k}}^r\lc_{k}^T\lc_{j}\rc_{j}^T\rc_{k}\dfrac{e^{(\lambda_k+\lambda_j){\tf}}-1}{\lambda_i+\lambda_j}+\sum_{\substack{j=1\\j\neq k}}^r\sum_{\substack{i=1\\i\neq k}}^r\lc_{i}^T\lc_{j}\rc_{j}^T\rc_{i}\dfrac{e^{(\lambda_i+\lambda_j){\tf}}-1}{\lambda_i+\lambda_j}.
+\end{dmath}
 To compute the gradient of the cost function $\J$ we perturb the cost functional with respect to the residue directions, i.e., $\lc_k\to \lc_k+\Delta\lc_k$ and $\rc_k\to \rc_k+\Delta\rc_k$: 
-\begin{align*}
-\widetilde{\J}_k&= \int_0^{\tf}\h(t)^T\h(t) dt-2\bigg( \sum_{j=1}^n(\lc_{k}+\Delta\lc_{k})^T\cc_j\bb_j^T(\rc_{k}+\Delta\rc_{k})\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j}\\
-&+\sum_{j=1}^n\sum_{\substack{i=1\\i\neq k}}^r\lc_{i}^T\cc_j\bb_j^T\rc_{i}\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j}\bigg)+ (\lc_{k}+\Delta\lc_{k})^T(\lc_{k}+\Delta\lc_{k})(\rc_{k}+\Delta\rc_{k})^T(\rc_{k}+\Delta\rc_{k})\dfrac{e^{(2\lambda_k){\tf}}-1}{2\lambda_k}\\
-&+\sum_{\substack{i=1 \\i \neq k}}^r\lc_{i}^T(\lc_{k}+\Delta\lc_{k})(\rc_{k}+\Delta\rc_{k})^T\rc_{i}\dfrac{e^{(\lambda_i+\lambda_k){\tf}}-1}{\lambda_i+\lambda_k}+                             
-\sum_{\substack{j=1\\j\neq k}}^r(\lc_{k}+\Delta\lc_{k})^T\lc_{j}\rc_{j}^T(\rc_{k}+\Delta\rc_{k})\dfrac{e^{(\lambda_k+\lambda_j){\tf}}-1}{\lambda_i+\lambda_j}\\
-&+\sum_{\substack{j=1\\j\neq k}}^r\sum_{\substack{i=1\\i\neq k}}^r\lc_{i}^T\lc_{j}\rc_{j}^T\rc_{i}\dfrac{e^{(\lambda_i+\lambda_j){\tf}}-1}{\lambda_i+\lambda_j}.
-\end{align*}
+\begin{dmath*}
+\widetilde{\J}_k= \int_0^{\tf}\h(t)^T\h(t) dt-2\bigg( \sum_{j=1}^n(\lc_{k}+\Delta\lc_{k})^T\cc_j\bb_j^T(\rc_{k}+\Delta\rc_{k})\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j} 
++\sum_{j=1}^n\sum_{\substack{i=1\\i\neq k}}^r\lc_{i}^T\cc_j\bb_j^T\rc_{i}\dfrac{e^{(\lambda_i+\rho_j){\tf}}-1}{\lambda_i+\rho_j}\bigg)+ (\lc_{k}+\Delta\lc_{k})^T(\lc_{k}+\Delta\lc_{k})(\rc_{k}+\Delta\rc_{k})^T(\rc_{k}+\Delta\rc_{k})\dfrac{e^{(2\lambda_k){\tf}}-1}{2\lambda_k}
++\sum_{\substack{i=1 \\i \neq k}}^r\lc_{i}^T(\lc_{k}+\Delta\lc_{k})(\rc_{k}+\Delta\rc_{k})^T\rc_{i}\dfrac{e^{(\lambda_i+\lambda_k){\tf}}-1}{\lambda_i+\lambda_k}+                             
+\sum_{\substack{j=1\\j\neq k}}^r(\lc_{k}+\Delta\lc_{k})^T\lc_{j}\rc_{j}^T(\rc_{k}+\Delta\rc_{k})\dfrac{e^{(\lambda_k+\lambda_j){\tf}}-1}{\lambda_i+\lambda_j}
++\sum_{\substack{j=1\\j\neq k}}^r\sum_{\substack{i=1\\i\neq k}}^r\lc_{i}^T\lc_{j}\rc_{j}^T\rc_{i}\dfrac{e^{(\lambda_i+\lambda_j){\tf}}-1}{\lambda_i+\lambda_j}.
+\end{dmath*}
 Then, collecting  the terms that are multiplied by $\Delta\lc_{k}$ and $\Delta\rc_{k}$, we obtain 
-\begin{align*}
-\nabla_{\rc_{k} }\J &=-2\lc_k^T\sum_{j=1}^n\cc_j\bb_j^T\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j}+2\lc_k^T\sum_{j=1}^n\lc_j\rc_j^T\dfrac{e^{(\lambda_k+\lambda_j){\tf}}-1}{\lambda_k+\lambda_j},~~ \mbox{and}\\
-\nabla_{\lc_{k} }\J&=-2\left (\sum_{j=1}^n\cc_j\bb_j^T\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j}\right)\rc_k+2\left(\sum_{j=1}^n\lc_j\rc_j^T\dfrac{e^{(\lambda_k+\lambda_j){\tf}}-1}{\lambda_k+\lambda_j}\right)\rc_k.\\
-\end{align*}
+\begin{dmath*}
+\nabla_{\rc_{k} }\J =-2\lc_k^T\sum_{j=1}^n\cc_j\bb_j^T\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j}+2\lc_k^T\sum_{j=1}^n\lc_j\rc_j^T\dfrac{e^{(\lambda_k+\lambda_j){\tf}}-1}{\lambda_k+\lambda_j},
+\end{dmath*}
+\begin{dmath*}
+\nabla_{\lc_{k} }\J=-2\left (\sum_{j=1}^n\cc_j\bb_j^T\dfrac{e^{(\lambda_k+\rho_j){\tf}}-1}{\lambda_k+\rho_j}\right)\rc_k+2\left(\sum_{j=1}^n\lc_j\rc_j^T\dfrac{e^{(\lambda_k+\lambda_j){\tf}}-1}{\lambda_k+\lambda_j}\right)\rc_k.\\
+\end{dmath*}
 Setting $\nabla_{\rc_{k} }\J=0   \text{ and }\nabla_{\lc_{k} }\J=0 $, and using Lemma \ref{mgexp}, mainly
@@
 After some tedious manipulations, we obtain 
-\begin{equation}\label{eqn:l1} 
-\begin{aligned}
-\pder[\J]{\lambda_k} =-2\lc_k^T &\left(\sum_{j=1}^n \cc_j\bb_j^T \dfrac{{\tf}(\lambda_k+\rho_j)e^{(\rho_j+\lambda_k){\tf}}-e^{(\rho_j+\lambda_k){\tf}}+1}{(\lambda_k+\rho_j)^2}\right)\rc_k  \\
-&+2\lc_k^T \left(\sum^r_{i=1} \lc_i\rc_i^T\dfrac{{\tf}(\lambda_i+\lambda_j\k)e^{(\lambda_i+\lambda_k){\tf}}-e^{(\lambda_i+\lambda_k){\tf}}+1}{(\lambda_i+\lambda_k)^2}\right)\rc_k.
-\end{aligned}
-\end{equation}
+%\begin{equation}\label{eqn:l1} 
+%\begin{aligned}
+\begin{dmath}\label{eqn:l1} 
+\pder[\J]{\lambda_k} =-2\lc_k^T \left(\sum_{j=1}^n \cc_j\bb_j^T \dfrac{({\tf}(\lambda_k+\rho_j)-1)e^{(\rho_j+\lambda_k){\tf}}+1}{(\lambda_k+\rho_j)^2}\right)\rc_k  
++2\lc_k^T \left(\sum^r_{i=1} \lc_i\rc_i^T\dfrac{({\tf}(\lambda_i+\lambda_j\k)-1)e^{(\lambda_i+\lambda_k){\tf}}+1}{(\lambda_i+\lambda_k)^2}\right)\rc_k.
+\end{dmath}
+%\end{aligned}
+%\end{equation}
 Lemma \eqref{mgexp}, specifically \eqref{eqn:Gpatlk} and \eqref{eqn:Grpatlk}, show that the expressions in the parentheses in \eqref{eqn:l1}  are, respectively, $-\G'(-\lambda_k)$ and $-\G_r'(-\lambda_k)$.
-Thus,
- setting $\pder[\J]{\lambda_k}= 0$ yield 
+If $\pder[\J]{\lambda_k}= 0$, then
 \begin{align*}
@@
 The complete numerical framework for the general case  with further details on the underlying optimization schemes will be discussed in a separate work.
-
 \subsection{A descent-type algorithm for the single-input/single-output case}
@@
 \begin{equation} \label{eqn:sisoh}
-\HH(s) =  \bfc^T(s\I-\A)^{-1}\bfb = \sum_{i=1}^n \frac{\psi_i}{s-\rho_i}\qquad\mbox{and}\qquad
-\HH_r(s) =  \bfc_r^T(s\I_r-\A_r)^{-1}\bfb_r = \sum_{i=1}^r \frac{\phi_i}{s-\lambda_i},
+\begin{aligned}
+\HH(s) &=  \bfc^T(s\I-\A)^{-1}\bfb = \sum_{i=1}^n \frac{\psi_i}{s-\rho_i}\\
+\HH_r(s) &=  \bfc_r^T(s\I_r-\A_r)^{-1}\bfb_r = \sum_{i=1}^r \frac{\phi_i}{s-\lambda_i},
+\end{aligned}
 \end{equation}
@@
 Given the SISO transfer functions $\HH(s)$ and $\HH_r(s)$ as defined in \eqref{eqn:sisoh}, define
-$$
-\G(s)=-e^{-s{\tf}}\bfc^T(s\I-\A)^{-1}e^{\A {\tf}}\bfb+\HH(s)\quad\mbox{and}\quad\G_r(s)=-e^{-s{\tf}}\bfc^T_r(s\I_r-\A_r)^{-1}e^{\A_r {\tf}}\bfb_r+\HH_r(s).
-$$
+
+\begin{align*}
+\G(s)&=-e^{-s{\tf}}\bfc^T(s\I-\A)^{-1}e^{\A {\tf}}\bfb+\HH(s),\\
+\G_r(s)&=-e^{-s{\tf}}\bfc^T_r(s\I_r-\A_r)^{-1}e^{\A_r {\tf}}\bfb_r+\HH_r(s).
+\end{align*}
 If $\HH_r$ is the best $r^{th}$ order approximation of $\HH$ with respect to the $\ch_2({\tf})$ norm, then
 \begin{align} \label{def:GGrsiso}
-\G(-\lambda_k)=\G_r(-\lambda_k),\qquad \mbox{and}\qquad
+\G(-\lambda_k)=\G_r(-\lambda_k),\ \mbox{and}\
 \G'(-\lambda_k)=\G_r'(-\lambda_k)
@@
 in \eqref{def:GGrsiso}.  Assume  the reduced  poles   $\{\lambda_i\}_{i=1}^r$ are fixed. Then,
-$\HH_r(s)$ is the best $r^{th}$ order approximation of $\HH(s)$ with respect to the $\ch_2({\tf})$ norm if and only if 
-\begin{align*}
-\M \bfphi= \mathbf{z}, \quad \mbox{or~equivalently},\quad \G(-\lambda_k)=\G_r(-\lambda_k),\quad\mbox{for}\quad k = 1,2,\ldots, r, 
-\end{align*}
+$\HH_r(s)$ is the best $r^{th}$ order approximation of $\HH(s)$ with respect to the $\ch_2({\tf})$ norm if and only if  $\M \bfphi= \mathbf{z}$, \mbox{or~equivalently},$\G(-\lambda_k)=\G_r(-\lambda_k)$,\ \mbox{for}\ $k = 1,2,\ldots, r,$
 where  $\bfphi = [\phi_1~\phi_2~\cdots~\phi_r]^T \in \CC^r$ is the vector of  residues; 
@@
 $$
-\mathbf{z}_j=e^{\lambda_j{\tf}}\bfc^T(-\lambda_j\I-\A)^{-1}e^{\A {\tf}}\bfb-\HH(-\lambda_j), \quad \mbox{for}\quad
+\mathbf{z}_j=e^{\lambda_j{\tf}}\bfc^T(-\lambda_j\I-\A)^{-1}e^{\A {\tf}}\bfb-\HH(-\lambda_j), 
 i=1,2,\ldots,r ;
@@
 \end{align*}
-where $\mathbf{w} \in \CC^{r\times 1}$ has the entries $\mathbf{w}_i=\sum_{k=1}^n \psi_k \dfrac{e^{(\rho_k+\lambda_i){\tf}}-1}{\lambda_i+\rho_k}$ for $i=1,2,\ldots,r$. Note that 
+where $\mathbf{w} \in \CC^{r\times 1}$ has the entries
+ $$\mathbf{w}_i=\sum_{k=1}^n \psi_k \dfrac{e^{(\rho_k+\lambda_i){\tf}}-1}{\lambda_i+\rho_k}  \quad \mbox{for}\quad i=1,2,\ldots,r.$$
+Note that 
 $\M$ is positive definite, $\phi^{T}\M\phi=\norm{\h_r}_{\ch_2({\tf})}^2 > 0$, and the cost function is quadratic in $\bfphi$.
@@
 Gains were significantly better for POD, especially for larger $r$ values.
+% \begin{figure}[H] 
+% \centering
+% \begin{subfigure}{.5\textwidth}
+% \centering
+% \includegraphics [scale=0.4]{allDataHeatF} 
+% \caption{\FH \ and other algorithms for a heat model\label{fig:hm}}
+% \end{subfigure}
+% \begin{subfigure}{.5\textwidth}
+% \centering
+% \includegraphics [scale=0.4]{ErrorVSrISSF}
+% \caption{\FH \ and other algorithms for the ISS model \label{fig:iss}}
+% \end{subfigure}
+%  \begin{subfigure}{.5\textwidth}
+% \centering
+%   \includegraphics [scale=0.4]{ErrorVSrUNSF}
+%      \caption{\FH \ and other algorithms for the unstable model \label{fig:uns}}
+% \end{subfigure}
+% \caption{Numerical results for three  models \label{mainfig}}
+% \end{figure}
+
   \begin{figure}[H] 
- \centering
- \begin{subfigure}{.5\textwidth}
  \centering
@@
  \caption{\FH \ and other algorithms for a heat model\label{fig:hm}}
- \end{subfigure}
- \begin{subfigure}{.5\textwidth}
+ \end{figure}
+ \begin{figure}[H]
  \centering
@@
  \caption{\FH \ and other algorithms for the ISS model \label{fig:iss}}
- \end{subfigure}
-  \begin{subfigure}{.5\textwidth}
+ \end{figure}
+  \begin{figure}[H]
  \centering
@@
       \caption{\FH \ and other algorithms for the unstable model \label{fig:uns}}
- \end{subfigure}
- \caption{Numerical results for three  models \label{mainfig}}
  \end{figure}
+
   Overall,  as expected, \FH \ yields a better approximation compared to the other algorithms for each model. 

### 2019-04-09 FH Optimality Conditions Paper.tex
some further changes
subplots and more details on diff wrt lambda

@@
 \input{paperPreamble}
+
+%to avoid whitespace
 \setlength{\parskip}{0pt}
@@
 \setlength{\partopsep}{0pt}
+
 \begin{document}
@@
 
-It is immediately clear that, the $\ch_2(t_f)$-error satisfies 
+It is immediately clear that, the $\ch_2(t_f)$-error, denoted by $\J$, satisfies 
 \begin{equation}\label{merexp}
 \begin{aligned}
-%\scalemath{0.9}
-{
-\norm{\h-\h_r}_{\ch_2({\tf})}^2=\norm{\h}_{\ch_2({\tf})}^2-2 \langle\h, \h_r\rangle_{\ch_2({\tf})}+\norm{\h_r}_{\ch_2({\tf})}^2,
-}
+\J&=\norm{\h-\h_r}_{\ch_2({\tf})}^2\\
+&=\norm{\h}_{\ch_2({\tf})}^2-2 \langle\h, \h_r\rangle_{\ch_2({\tf})}+\norm{\h_r}_{\ch_2({\tf})}^2,
 \end{aligned}
@@
 \emph{{Proof of Theorem \ref{mh2opt}.}}
- Let $\J$ denote the $\ch_2({\tf})$ error norm square, i.e., 
+ As mentioned above, let $\J$ denote the $\ch_2({\tf})$ error norm square, i.e., 
 \begin{align*}
@@
 \begin{align*}
-\lc_k^T\G(-\lambda_k)=\lc_k^T\G_r(-\lambda_k) \quad \mbox{and}\quad
-\G(-\lambda_k)\rc_k=\G_r(-\lambda_k)\rc_k,
+\lc_k^T\G(-\lambda_k)&=\lc_k^T\G_r(-\lambda_k), \\ %\quad \mbox{and}\quad
+\G(-\lambda_k)\rc_k&=\G_r(-\lambda_k)\rc_k,
 \end{align*}
 which proves \eqref{eqn:cond1} and \eqref{eqn:cond2}. To prove \eqref{eqn:cond3},  we differentiate $\J$ in \eqref{merror}  with respect to the $k$-th pole $\lambda_k$: 
-After some tedious manipulations, we obtain 
 %\begin{equation}\label{eqn:l1} 
 %\begin{aligned}
-\begin{dmath}\label{eqn:l1} 
-\pder[\J]{\lambda_k} =-2\lc_k^T \left(\sum_{j=1}^n \cc_j\bb_j^T \dfrac{({\tf}(\lambda_k+\rho_j)-1)e^{(\rho_j+\lambda_k){\tf}}+1}{(\lambda_k+\rho_j)^2}\right)\rc_k  
-+2\lc_k^T \left(\sum^r_{i=1} \lc_i\rc_i^T\dfrac{({\tf}(\lambda_i+\lambda_j\k)-1)e^{(\lambda_i+\lambda_k){\tf}}+1}{(\lambda_i+\lambda_k)^2}\right)\rc_k.
+\begin{dmath}\label{eqn:diffl}
+\pder[\J]{\lambda_k}=-2\lc_k^T \left(\sum_{j=1}^n \cc_j\bb_j^T \dfrac{{\tf}(\lambda_k+\rho_j)e^{(\rho_j+\lambda_k){\tf}}} {(\lambda_k+\rho_j)^2}     -\dfrac       {e^{(\rho_j+\lambda_k){\tf}}-1}{(\lambda_k+\rho_j)^2}\right)\rc_k
++2\lc_k^T \left(\sum^r_{i=1} \lc_i\rc_i^T\dfrac{{\tf}(\lambda_i+\lambda_j\k)e^{(\lambda_i+\lambda_k){\tf}}}  {(\lambda_i+\lambda_k)^2}         -\dfrac{e^{(\lambda_i+\lambda_k){\tf}}-1}{(\lambda_i+\lambda_k)^2}\right)\rc_k.
 \end{dmath}
+
+We rewrite \eqref{eqn:diffl} to obtain
+
+\begin{dmath}
+\pder[\J]{\lambda_k} =-2K_1+2K_2
+\end{dmath}
+where
+\begin{dmath}\label{eqn:l11} 
+K_1=\lc_k^T \left(\sum_{j=1}^n \cc_j\bb_j^T \dfrac{({\tf}(\lambda_k+\rho_j)-1)e^{(\rho_j+\lambda_k){\tf}}+1}{(\lambda_k+\rho_j)^2}\right)\rc_k  
+\end{dmath}
+\begin{dmath}\label{eqn:l12} 
+K_2=\lc_k^T \left(\sum^r_{i=1} \lc_i\rc_i^T\dfrac{({\tf}(\lambda_i+\lambda_j\k)-1)e^{(\lambda_i+\lambda_k){\tf}}+1}{(\lambda_i+\lambda_k)^2}\right)\rc_k.
+\end{dmath}
+
+%\begin{dmath}\label{eqn:l1} 
+%\pder[\J]{\lambda_k} =-2\lc_k^T \left(\sum_{j=1}^n \cc_j\bb_j^T \dfrac{({\tf}(\lambda_k+\rho_j)-1)e^{(\rho_j+\lambda_k){\tf}}+1}{(\lambda_k+\rho_j)^2}\right)\rc_k  
+%+2\lc_k^T \left(\sum^r_{i=1} \lc_i\rc_i^T\dfrac{({\tf}(\lambda_i+\lambda_j\k)-1)e^{(\lambda_i+\lambda_k){\tf}}+1}{(\lambda_i+\lambda_k)^2}\right)\rc_k.
+%\end{dmath}
 %\end{aligned}
 %\end{equation}
-Lemma \eqref{mgexp}, specifically \eqref{eqn:Gpatlk} and \eqref{eqn:Grpatlk}, show that the expressions in the parentheses in \eqref{eqn:l1}  are, respectively, $-\G'(-\lambda_k)$ and $-\G_r'(-\lambda_k)$.
+Lemma \eqref{mgexp}, specifically \eqref{eqn:Gpatlk} and \eqref{eqn:Grpatlk}, show that the expressions in the parentheses in \eqref{eqn:l11}  and \eqref{eqn:l12}   are, respectively, $-\G'(-\lambda_k)$ and $-\G_r'(-\lambda_k)$.
 If $\pder[\J]{\lambda_k}= 0$, then
@@
 
+ %This is the original figure
+%  \begin{figure}[H] 
+% \centering
+% \includegraphics [scale=0.4]{allDataHeatF} 
+% \caption{\FH \ and other algorithms for a heat model\label{fig:hm}}
+% \end{figure}
+% \begin{figure}[H]
+% \centering
+%  \includegraphics [scale=0.4]{ErrorVSrISSF}
+% \caption{\FH \ and other algorithms for the ISS model \label{fig:iss}}
+% \end{figure}
+%  \begin{figure}[H]
+% \centering
+%   \includegraphics [scale=0.4]{ErrorVSrUNSF}
+%      \caption{\FH \ and other algorithms for the unstable model \label{fig:uns}}
+% \end{figure}
+%
+
  \begin{figure}[H] 
  \centering
- \includegraphics [scale=0.4]{allDataHeatF} 
+ \includegraphics [scale=0.4]{fig1} 
  \caption{\FH \ and other algorithms for a heat model\label{fig:hm}}
@@
  \centering
-  \includegraphics [scale=0.4]{ErrorVSrISSF}
- \caption{\FH \ and other algorithms for the ISS model \label{fig:iss}}
+  \includegraphics [scale=0.4]{fig2}
+ \caption{\FH \ and other algorithms for the heat model \label{fig:hm1}}
  \end{figure}
@@
  \centering
-   \includegraphics [scale=0.4]{ErrorVSrUNSF}
+   \includegraphics [scale=0.4]{fig3}
+      \caption{\FH \ and other algorithms for the ISS model \label{fig:iss}}
+ \end{figure}
+
+ \begin{figure}[H]
+ \centering
+   \includegraphics [scale=0.4]{fig4}
       \caption{\FH \ and other algorithms for the unstable model \label{fig:uns}}

### 2019-04-16 response.tex
some further changes
test

@@
 used to state the purpose of the paper.  \\
-\serkan{\textsf{AR}:  Should we move these lines to the introduction section?}
+\serkan{\textsf{AR}: We thank the referee for the comment.
+These 5 lines state the purpose of infinite horizon optimal model reduction.
+The purpose of the paper i.e.
+optimal model reduction on a finite horizon,  is stated in the introductory section.}
 
@@
 
-\serkan{\textsf{AR}:  Should we remove these lines completely, or argue that it makes the paper easier to read for someone who is not as familiar with the concepts in the paper? }
+\serkan{\textsf{AR}:  We thank the referee for the comment. We have removed these 5 lines and reorganized the beginning of section 3 }
 
@@
 
-\serkan{\textsf{AR}:  How should we organize (reorganize?) this section? }
+\serkan{\textsf{AR}:  We thank the referee for the comment and we have implemented this suggestion. }
 
@@
 allowing to obtain such result. \\
-\serkan{\textsf{AR}:  What details? Should I include the trivial differentiations?}
+\serkan{\textsf{AR}:  We thank the referee for the comment. We have added more details explaining the computation.}
 
@@
 \item 3.3 section can be a (concise) part of the introduction of the paper. \\
-\serkan{\textsf{AR}:  We thank the referee for the comment but we respectfully disagree???.}
+\serkan{\textsf{AR}:  We thank the referee for the comment.
+We believe that section 3.3 addresses the importance of the result.
+We believe that explaining the result after it is given is more appropriate and ti makes it more tangible for the reader.}
 
@@
 the reduced order model which is not the case in the current version.\\
-\serkan{\textsf{AR}:  We thank the referee for the comment but it is the case in the current version???.}
+\serkan{\textsf{AR}:  We thank the referee for the comment.
+We believe that the Corollaries 4.1 and 4.2, which follow directly from the main result, explain how we are using the optimality conditions to construct the algorithm.}
 
@@
 
-\serkan{\textsf{AR}:  How should we proceed here?.}
+\serkan{\textsf{AR}:  We thank the referee for the comment. We have addressed the issue by providing one figure for each comparison.}
 

### 2019-04-16 FH Optimality Conditions Paper.tex
test

@@
 \section{$\ch_2(\tf)$ Optimal Model Reduction on a Finite Horizon}  \label{sec:main} 
-If we are interested  in the behavior of a system only in a finite time interval (or access to a numerical simulation only on a finite time interval), we can obtain better reduced models by defining the measure of approximation quality over the interval of interest, instead of  focussing on the infinite time horizon.
-Also, in case of an unstable system, since the impulse grows unbounded as $t \to \infty$, the regular $\ch_2$ measure cannot be applied.
+%If we are interested  in the behavior of a system only in a finite time interval (or access to a numerical simulation only on a finite time interval), we can obtain better reduced models by defining the measure of approximation quality over the interval of interest, instead of  focussing on the infinite time horizon.
+%
+{\color{red}
+In this section, we discuss the main result of the paper i.e.
+the $\ch_2(\tf)$ local optimality conditions, their proof, and the importance of such conditions for model reduction of large-scale dynamical systems.}
+
+%Also, in case of an unstable system, since the impulse grows unbounded as $t \to \infty$, the regular $\ch_2$ measure cannot be applied.
 Therefore, we will focus on model reduction on a finite time horizon.
-
-We would like to point out that for an unstable dynamical system, if the matrix $\A$ does not have purely imaginary eigenvalues, one can work with the
- $\mathcal{L}_2$ norm by decomposing it into a stable and anti-stable system, and then obtain an interpolatory reduced model based on this measure.
-However, this solution requires destroying the causality of the underlying dynamics \cite{magruder2010rational}.
-This is not the framework we are interested in here and we will work with a finite-time interval.
 
@@
 \end{equation}
-such that $\norm{\h - \h_r}_{\ch_2(\tf)}$ is minimized. As in the regular $\ch_2$ case, this is a non-convex optimization problem and we will focus on local minimizers. Using 
-the parametrization \eqref{kernel2}, we we will derive interpolation-based necessary conditions for optimality.
+such that $\norm{\h - \h_r}_{\ch_2(\tf)}$ is minimized.
+As in the regular $\ch_2$ case, this is a non-convex optimization problem and we will focus on local minimizers.
+Using the parametrization \eqref{kernel2}, we we will derive interpolation-based necessary conditions for optimality.
 The main result is given by Theorem \ref{mh2opt}. However, we  need many supplementary results, Lemmas 
-\ref{minnerexp}-\ref{mgexp}, to reach this final conclusion.
-
-It is immediately clear that, the $\ch_2(t_f)$-error, denoted by $\J$, satisfies 
+\ref{minnerexp}-\ref{mgexp}, to reach this final conclusion. It is immediately clear that, the $\ch_2(t_f)$-error, denoted by $\J$, satisfies 
 \begin{equation}\label{merexp}
@@
 \end{align*}
-which proves \eqref{eqn:cond1} and \eqref{eqn:cond2}. To prove \eqref{eqn:cond3},  we differentiate $\J$ in \eqref{merror}  with respect to the $k$-th pole $\lambda_k$: 
+which proves \eqref{eqn:cond1} and \eqref{eqn:cond2}.
+To prove \eqref{eqn:cond3},  we differentiate $\J$ in \eqref{merror}  with respect to the $k$-th pole $\lambda_k$.
+Note that we have written $\J$ in such a way to isolate the terms that depend on $\lambda_k$ from the ones that do not.
+Thus, many of the terms in \eqref{merror} have a zero derivative and we obtain:
 %\begin{equation}\label{eqn:l1} 
@@
 \end{dmath}
-
+{\color{red}
+Note that the first term in \eqref{eqn:diffl} corresponds to the derivative of the second term in \eqref{merror} and the second term in \eqref{eqn:diffl} corresponds to the derivative of the last four terms in \eqref{merror}.}
 We rewrite \eqref{eqn:diffl} to obtain
@@
 The algorithmic implications of this result will be revisited in Section \ref{sec:discuss}.}
+\end{remark}
+
+\begin{remark}
+{\color {red}
+We would like to point out that for an unstable dynamical system, if the matrix $\A$ does not have purely imaginary eigenvalues, one can work with the
+ $\mathcal{L}_2$ norm by decomposing it into a stable and anti-stable system, and then obtain an interpolatory reduced model based on this measure.
+However, this solution requires destroying the causality of the underlying dynamics \cite{magruder2010rational}.
+This is not the framework we are interested in here and we will work with a finite-time interval.
+}
  \end{remark}
@@
 
- \begin{figure}[H] 
+ \begin{figure}%[H] 
  \centering
- \includegraphics [scale=0.4]{fig1} 
+ \includegraphics [scale=0.5]{fig1h} 
  \caption{\FH \ and other algorithms for a heat model\label{fig:hm}}
  \end{figure}
- \begin{figure}[H]
+ \begin{figure}%[H]
  \centering
-  \includegraphics [scale=0.4]{fig2}
- \caption{\FH \ and other algorithms for the heat model \label{fig:hm1}}
+  \includegraphics [scale=0.5]{fig2i}
+ \caption{\FH \ and other algorithms for the heat model \label{fig:iss}}
  \end{figure}
-  \begin{figure}[H]
+  \begin{figure}%[H]
  \centering
-   \includegraphics [scale=0.4]{fig3}
-      \caption{\FH \ and other algorithms for the ISS model \label{fig:iss}}
+   \includegraphics [scale=0.5]{fig3u}
+      \caption{\FH \ and other algorithms for the ISS model \label{fig:uns}}
  \end{figure}
 
- \begin{figure}[H]
- \centering
-   \includegraphics [scale=0.4]{fig4}
-      \caption{\FH \ and other algorithms for the unstable model \label{fig:uns}}
- \end{figure}
+% \begin{figure}[H]
+% \centering
+%   \includegraphics [scale=0.4]{fig4}
+%      \caption{\FH \ and other algorithms for the unstable model \label{fig:uns}}
+% \end{figure}
+%
 

### 2019-04-17 FH Optimality Conditions Paper.tex
Formating of main proof
added footnote

@@
 In this section, we discuss the main result of the paper i.e.
-the $\ch_2(\tf)$ local optimality conditions, their proof, and the importance of such conditions for model reduction of large-scale dynamical systems.}
+the $\ch_2(\tf)$ \footnote{Even though we are on the time domain we will still use the $\ch_2(\tf)$ notation.
+Since the time domain and frequency domain functions are equivalent,  this notation is appropriate.
+Furthermore, since the infinite horizon, and finite horizon case appear to be connected, the notation $\mathcal{H}_2(t_f)$ illustrates this connection.
+\cite{RedK17,GoyR17} use similar notation.} local optimality conditions, their proof, and the importance of such conditions for model reduction of large-scale dynamical systems.}
 
@@
 \end{pf}%%%Start proof here
-\emph{{Proof of Theorem \ref{mh2opt}.}}
+\textbf{{PROOF OF THEOREM \ref{mh2opt}.}}
  As mentioned above, let $\J$ denote the $\ch_2({\tf})$ error norm square, i.e., 

### 2019-05-01 response.tex
added impulse plots

@@
 \documentclass[11pt]{article}
+\newcommand{\FH}{FHIRKA}
 
@@
 Since the time domain and frequency domain functions are equivalent, we believe this notation is appropriate.
-Furthermore, since the infinite horizon, and finite horizon case appear to be connected, the notation $\mathcal{H}_2(t_f)$ illustrates this connection.}
+Furthermore, since the infinite horizon, and finite horizon case appear to be connected, the notation $\mathcal{H}_2(t_f)$ illustrates this connection.
+We have explained further in the paper our choice of notation.
+} 
 \end{itemize}
@@
 importance of the proposed method.\\
-\serkan{\textsf{AR}:  What model should we use for this?}
+\serkan{\textsf{AR}:  We thank the referee for the comment. We have added a graph showing the time responses for the error for different methods. Here are two additional graphs for two other models}
+ \begin{figure}[H]
+ \centering
+   \includegraphics [scale=0.5]{absHeatErrorYlim}
+      \caption{POD vs \FH for a heat model\label{fig:impulseHeat}}
+ \end{figure}
+ \begin{figure}[H]
+ \centering
+   \includegraphics [scale=0.5]{absUNSError}
+      \caption{POD vs \FH for an unstable model\ \label{fig:impulseUns}}
+ \end{figure}
 

### 2019-05-01 FH Optimality Conditions Paper.tex
added impulse plots

@@
 Gains were significantly better for POD, especially for larger $r$ values.
+{\color{red} Also, in Figure \ref{impulse}, we compare the impulse responses of the error systems of POD and \FH \ for the ISS model.
+For both methods, POD and \FH\  the reduced model was of order $r=14$.
+As we can see from the graph, \FH \ clearly outperforms POD on the time interval $[0, 1]$.
+
+}
+
 % \begin{figure}[H] 
@@
 
- \begin{figure}%[H] 
+ \begin{figure}[H] 
  \centering
- \includegraphics [scale=0.5]{fig1h} 
+ \includegraphics [scale=0.35]{fig1h} 
  \caption{\FH \ and other algorithms for a heat model\label{fig:hm}}
  \end{figure}
- \begin{figure}%[H]
+ \begin{figure}[H]
  \centering
-  \includegraphics [scale=0.5]{fig2i}
+  \includegraphics [scale=0.35]{fig2i}
  \caption{\FH \ and other algorithms for the heat model \label{fig:iss}}
  \end{figure}
-  \begin{figure}%[H]
+  \begin{figure}[H]
  \centering
-   \includegraphics [scale=0.5]{fig3u}
+   \includegraphics [scale=0.35]{fig3u}
       \caption{\FH \ and other algorithms for the ISS model \label{fig:uns}}
@@
 
-% \begin{figure}[H]
-% \centering
-%   \includegraphics [scale=0.4]{fig4}
-%      \caption{\FH \ and other algorithms for the unstable model \label{fig:uns}}
-% \end{figure}
-%
+ \begin{figure}[H]
+ \centering
+   \includegraphics [scale=0.35]{ISS1RErrorSmooth}
+      \caption{\FH \ and other algorithms for the unstable model \label{fig:impulse}}
+ \end{figure}
 
@@
 %\bibliographystyle{elsarticle-num}
+
 \bibliographystyle{plain}

### 2019-05-01 FH Optimality Conditions Paper.tex
SG changes

@@
 %Introduction
+%Introduction
 \section{Introduction}
-Simulation, design, and control of dynamical systems play an important role in numerous scientific and industrial tasks such as signal propagation in the nervous system\cite{Steve Cox}, heat dissipation, prediction of major weather events \cite{BoyP11,Sha03,Ant05}.
+%Simulation, design, and control of dynamical systems play an important role in numerous scientific and industrial tasks such as signal propagation in the nervous system\cite{kellems2009low}, heat dissipation, prediction of major weather events \cite{BoyP11,Sha03,Ant05}.
+Simulation, design, and control of dynamical systems play an important role in numerous scientific and industrial tasks such as \textcolor{red}{signal propagation in the nervous system\cite{kellems2009low};  the synthesis  of
+interconnect \cite{BonD07} and semiconductor devices \cite{Hess2014}; large-scale inverse problem,
+ \cite{Druskin2011solution,Lieberman2010,de2015nonlinear}; and prediction of major weather events \cite{BoyP11,Sha03,Ant05}.}
 The need for  detailed  models due to the increasing demand for greater resolution leads to  
@@
 Consider the linear time-invariant dynamical system:
+%\begin{equation}  \label{fom1}
+%\begin{array}{rcl}
+%\dot{\x}(t) & =& \A\x(t) +\B\uu(t)\\
+%\y(t)&=&\C\x(t) 
+%\end{array}
+%%~~\mbox{with}~~ \x(0) = \mathbf{0},
+%\; \Leftrightarrow \;
+%\y(t)=\int_{0}^{t}\h (t-\tau)\uu(\tau) d\tau,
+%\end{equation}
 \begin{equation}  \label{fom1}
-\begin{array}{rcl}
-\dot{\x}(t) & =& \A\x(t) +\B\uu(t)\\
-\y(t)&=&\C\x(t) 
-\end{array}
-%~~\mbox{with}~~ \x(0) = \mathbf{0},
-\; \Leftrightarrow \;
-\y(t)=\int_{0}^{t}\h (t-\tau)\uu(\tau) d\tau,
+\begin{aligned}
+\dot{\x}(t) & = \A\x(t) +\B\uu(t),~~~\x(0) = \mathbf{0},\\
+\y(t)&= \C\x(t)  = \int_{0}^{t}\h (t-\tau)\uu(\tau) d\tau,
+\end{aligned}
 \end{equation}
 \noindent
-where $  \x(0) = \mathbf{0}, \A \in \R^{n\times n}$, $\B\in \R^{n\times m}$, and $\C\in \R^{p\times n}$ are constant matrices; 
+where %$  \x(0) = \mathbf{0}, \A \in \R^{n\times n}$,
+$ \A \in \R^{n\times n}$,
+ $\B\in \R^{n\times m}$, and $\C\in \R^{p\times n}$ are constant matrices; 
 the variable $\x(t)\in \R^n$ denotes the internal variables, $\uu(t)\in \R^m$ denotes the control inputs, and $\y(t)\in \R^p$ denotes the outputs; and $\h(t)=\C e^{\A t}\B$ is the impulse response of the full model.
@@
 reduction achieves this by replacing the original model with a lower dimensional one:
+%\begin{equation}  \label{rom1}
+%%\begin{array}{rcl}
+%\begin{aligned}
+%\dot{\x}_r(t)&=\A_r\x_r(t)+\B_r\uu(t)\\
+%\y_r(t)&=\C_r\x_r(t) 
+%\end{aligned}
+%%\end{array}
+%%~~\mbox{with}~~ \x_r(0) = \mathbf{0},
+%\; \Leftrightarrow \;
+%\y_r(t)=\int_{0}^{t}\h_r (t-\tau)\uu(\tau)d\tau,
+%\end{equation}
 \begin{equation}  \label{rom1}
-%\begin{array}{rcl}
 \begin{aligned}
-\dot{\x}_r(t)&=\A_r\x_r(t)+\B_r\uu(t)\\
-\y_r(t)&=\C_r\x_r(t) 
+\dot{\x}_r(t)&=\A_r\x_r(t)+\B_r\uu(t),~~~\x_r(0) = \mathbf{0},\\
+\y_r(t)&=\C_r\x_r(t)  = \int_{0}^{t}\h_r (t-\tau)\uu(\tau)d\tau,
 \end{aligned}
-%\end{array}
-%~~\mbox{with}~~ \x_r(0) = \mathbf{0},
-\; \Leftrightarrow \;
-\y_r(t)=\int_{0}^{t}\h_r (t-\tau)\uu(\tau)d\tau,
 \end{equation}
-where as in \eqref{fom1}, $\x_r(0) = \mathbf{0}, \h_r(t)=\C_r e^{\A_r t}\B_r$ is the impulse response of the reduced model, and $\A_r \in \R^{r\times r}$, $\B_r\in \R^{r\times m}$, and $\C_r\in \R^{p\times r}$ with $r\ll n$.
+where as in \eqref{fom1}, %$\x_r(0) = \mathbf{0}, \h_r(t)=\C_r e^{\A_r t}\B_r$
+$\h_r(t)=\C_r e^{\A_r t}\B_r$  is the impulse response of the reduced model, and $\A_r \in \R^{r\times r}$, $\B_r\in \R^{r\times m}$, and $\C_r\in \R^{p\times r}$ with $r\ll n$.
 The goal is that the output of the reduced model, $\y_r(t)$, approximates the true output, $\y(t)$, of the original system accurately in an appropriate norm.
@@
 For time-limited balanced truncation, \cite{GugA03} establishes an upper bound for the $\ch_\infty$ error between the full and reduced models,  \cite{RedK17} provides an $\ch_2$ error bound.
-In this paper,  we explore $\ch_2(\tf)$ optimality for model reduction over a finite time horizon.
-  We establish $\ch_2(\tf)$ optimality conditions over a finite time horizon and introduce an algorithm which yields better approximations of the large-scale dynamical systems compared to other model reduction methods.
 
-The optimality requires a parametrization of the reduced model.
-We will work with the time-domain representation of the dynamical system to derive the optimality condition.
+%In this paper,  we explore $\ch_2(\tf)$ optimality for model reduction over a finite time horizon. 
+%  We establish $\ch_2(\tf)$ optimality conditions over a finite time horizon and introduce an algorithm which yields better approximations of the large-scale dynamical systems compared to other model reduction methods.
+
+\textcolor{red}{In this paper,  we explore optimal model reduction over a finite time horizon.
+We use a time-limited version of the $\ch_2$ norm, denoted by $\ch_2(\tf)$, to quantify the model reduction error.}
+ Optimality requires a parametrization of the reduced model.
+We will work with the time-domain representation of the dynamical system to derive the optimality conditions.
 Specifically, we represent the impulse response of the reduced dynamical system using the modal decomposition, i.e.,
@@
 Using the parametrization of the reduced model in \eqref{kernel}, we derive interpolatory optimality conditions in the $\ch_2(\tf)$ norm and implement a model reduction algorithm that satisfies these optimality conditions.
- The advantages of time-limited model reduction under the $\ch_2(\tf)$ norm are two fold: first, one can obtain a better approximation in the time interval of interest, and, second, one can obtain a locally optimal reduced order model in the case of unstable systems as well.
+
+\textcolor{red}{The advantage of the interpolation framework we will develop  is that 
+we do not require the reduced-model to be obtained via projection, as usually assumed in model reduction \cite{Ant05}.
+Indeed, one observation we will make is that unlike in the infinite-horizon $\ch_2$ approximation problem, the optimal reduced model in the finite-horizon norm is not necessarily given by a projection and thus a projection-based approach will not be able to satisfy the optimality conditions.
+Therefore,
+by treating the poles and residues in \eqref{kernel} as the parameters and directly working with them, we  obtain a reduced model to satisfy the optimality conditions exactly.}
+
+%The advantages of time-limited model reduction under the $\ch_2(\tf)$ norm are two fold: first, one can obtain a better approximation in the time interval of interest, and, second, one can obtain a locally optimal reduced order model in the case of unstable systems as well.
 
@@
 {\color{red}
-In this section, we discuss the main result of the paper i.e.
-the $\ch_2(\tf)$ \footnote{Even though we are on the time domain we will still use the $\ch_2(\tf)$ notation.
-Since the time domain and frequency domain functions are equivalent,  this notation is appropriate.
-Furthermore, since the infinite horizon, and finite horizon case appear to be connected, the notation $\mathcal{H}_2(t_f)$ illustrates this connection.
-\cite{RedK17,GoyR17} use similar notation.} local optimality conditions, their proof, and the importance of such conditions for model reduction of large-scale dynamical systems.}
+In this section, we present  the main theoretical results of the paper i.e., the interpolatory $\ch_2(\tf)$  optimality conditions, and discuss their implications.}
 
@@
 For a finite-time horizon 
-$[0,t_f]$, the $\ch_2(\tf)$ inner product $\langle \cdot, \cdot \rangle_{\ch_2(\tf)}$ 
+$[0,t_f]$, the $\ch_2(\tf)$\footnote{\textcolor{red}{Even though we are using an error measure specifically formulated in the time domain,   following \cite{RedK17,GoyR17} we choose to use  the $\ch_2(\tf)$ notation, despite the term $\ch_2$  is mostly associated with a measure defined in the frequency domain.
+Our main reason is to keep the connection to the infinite horizon problem where the measure in the frequency- and time-domains are equivalent.
+And more importantly, as in the regular $\ch_2$ case, the optimality conditions will still appear as interpolation conditions in the frequency domain.
+ %In the Since the time domain and frequency domain functions are equivalent,  this notation is appropriate.
+Furthermore, since the infinite horizon, and finite horizon case appear to be connected, the notation $\mathcal{H}_2(t_f)$ illustrates this connection.
+\cite{RedK17,GoyR17} use similar notation.
+ }}
+   inner product $\langle \cdot, \cdot \rangle_{\ch_2(\tf)}$ 
 and $\ch_2(\tf)$ norm $\norm{\cdot}_{\ch_2(\tf)}$ are defined as  

### 2019-05-01 FH Optimality Conditions Paper.tex
fixed bibliography

@@
 interconnect \cite{BonD07} and semiconductor devices \cite{Hess2014}; large-scale inverse problem,
- \cite{Druskin2011solution,Lieberman2010,de2015nonlinear}; and prediction of major weather events \cite{BoyP11,Sha03,Ant05}.}
+ \cite{Druskin2011solution,Lieberman2010,de2015nonlinear}; and prediction of major weather events \cite{Ant05}.}
 The need for  detailed  models due to the increasing demand for greater resolution leads to  

### 2019-05-03 FH Optimality Conditions Paper.tex
fixed figure reference
laplace transform

@@
 {\color{red} In the infinite-horizon case, if  $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
-the interpolant is $\G_r(s)$, which is a function of $\HH_r(s)$, and the interpolated function is $\G(s)$, which is a function of $\HH(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$.
+the interpolant is $\G_r(s)$, and the interpolated function is $\G(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$.
 The interpolation of $\G(s)$ guarantees that we are approximating the model on the finite horizon.
@@
 \begin{align*}
-\mathcal{L}^{-1}\{\G(s)\}&=\h(t)+\sum_{i=1}^n u_{\tf}(t)\phi_i e^{\lambda_i {\tf}}e^{\lambda_i (t-{\tf})}\\
-&=\h(t)+\sum_{i=1}^n u_{\tf}(t)\phi_i e^{\lambda_i t}
+\mathcal{L}^{-1}\{\G(s)\}&=u_{\tf}\h(t)
 \end{align*}
@@
 For this reason, the interpolation of $\G(s)$ by $\G_r(s)$ is necessary.
-However, the optimal interpolation points  and directions are still determined by the reduced model $\HH_r(s)$.
-The algorithmic implications of this result will be revisited in Section \ref{sec:discuss}.}
+The optimal interpolation points  and directions are still determined by the reduced model $\HH_r(s)$. The algorithmic implications of this result will be revisited in Section \ref{sec:discuss}.}
 \end{remark}
@@
 Gains were significantly better for POD, especially for larger $r$ values.
-{\color{red} Also, in Figure \ref{impulse}, we compare the impulse responses of the error systems of POD and \FH \ for the ISS model.
+{\color{red} Also, in Figure \ref{fig:impulse}, we compare the impulse responses of the error systems of POD and \FH \ for the ISS model.
 For both methods, POD and \FH\  the reduced model was of order $r=14$.

### 2019-05-13 response.tex
Latest SG Changes

--- 2019-04-16 response.tex
@@
 
+Also in the revised manuscripts, all the revisions  are highlighted by color blue for the referees to spot them easily.
+
 \section*{Referee \#1}%change this about referee #1
@@
 function, which might be inappropriate from a mathematical point of
-view?\\
-\serkan{\textsf{AR}: We thank the reviewer for this comment.
-Since the time domain and frequency domain functions are equivalent, we believe this notation is appropriate.
-Furthermore, since the infinite horizon, and finite horizon case appear to be connected, the notation $\mathcal{H}_2(t_f)$ illustrates this connection.}
+view?\\[1ex]
+\serkan{\textsf{AR}: We agree with the referee that the term $H_2$ is associated with measure in the frequency domain.
+To clarify this point.
+we have added a footnote, Footnote 4 at the bottom of page 3, explaining the usage.
+} 
 \end{itemize}
@@
 accurately even though valuable elements of this positioning are given
-in section 3.3.
-
-\serkan{\textsf{AR}:   }
+in section 3.3. \\ [1ex]
+\serkan{\textsf{AR}:  We thank the referee for pointing this out.
+We have added some of this discussion from Section 3.3 to Introduction now (the second to last paragraph in Introduction to help resolve this issue.}
 
 \item Some particular application domains are cited in the beginning without
-justification. \\
-\serkan{\textsf{AR}:   }  
-\item What about the choice of the finite horizon tf ?
+justification. \\[1ex]
+\serkan{\textsf{AR}:  We have added citations.}  
+\item What about the choice of the finite horizon tf ? \\ [1ex]
+\serkan{\textsf{AR}: The choice of the finite horizon is really problem and application dependent; however the interpolation conditions hold for any $t_f$.
+We added a sentence right before Remark 3.2 to state this fact.}
 
-\serkan{\textsf{AR}: The choice of the finite horizon depends on the problem. We added a clarification for this issue at the end of section 3.2}
-
-\item Justify the choice of simple poles for the reduced order model. 
-\serkan{\textsf{AR}:  The choice of simple poles was motivated by the Iterative Rational Krylov Algorithm (IRKA).
-Since for many applications, the poles of the model are simple, we believe this choice is justified.
+\item Justify the choice of simple poles for the reduced order model. \\ [1ex]
+\serkan{\textsf{AR}:  The choice of simple poles was motivated by the infinite horizon $\mathcal{H}_2$ optimal approximation problems and the corresponding interpolation, as in, for example, citations [3,6,7, 11,23] in the paper.
+Indeed, the interpolatory $\mathcal{H}_2$  approximation has been extended some specific nonlinear dynamical systems as well, such as
+\begin{itemize}
+\item   P. Benner and T. Breiten, Interpolation-based $\mathcal{H}_2$-model reduction of
+                  bilinear control systems, SIAM Journal on Matrix Analysis and Applications, 2012.
+\item  G. Flagg and S. Gugercin,  Multipoint Volterra Series Interpolation and $\mathcal{H}_2$ Optimal Model Reduction of Bilinear Systems, SIAM Journal on Matrix Analysis and Applications, 2015.
+\item P. Benner , P. Goyal, and S. Gugercin, $\mathcal{H}_2$-quasi-optimal model order reduction for quadratic-bilinear control systems, SIAM Journal on Matrix Analysis and Applications, 2018.
+\end{itemize}
+These nonlinear extensions also use the same assumptions.
+It is certainly theoretically possible that an optimal $\mathcal{H}_2$ reduced model can have repeated poles and interpolation conditions have been extended to this case as well:
+\begin{itemize}
+\item P. van Dooren, K. Gallivan, and  P.-A. Absil, 
+$\mathcal{H}_2$-Optimal Model Reduction with Higher-Order Poles,
+	SIAM Journal on Matrix Analysis and Applications, 2010.
+\end{itemize}
+However, this is situation is rarely observed in practice and we chose to use simple poles. Also, we note that the full model can have repeated poles. The assumption is only on the reduced model.
 }
 
-\item What is the reference of the existing result given by Theorem 2.1 ? \\
-\serkan{\textsf{AR}:  We have added the citations immediately after the theorem in order to clarify any confusion. }
+\item What is the reference of the existing result given by Theorem 2.1 ? \\ [1ex]
+\serkan{\textsf{AR}:  We have added the citations immediately after the theorem.}
 
@@
 derivative of H with respect to s. This can be stated at the end of the
-introduction section (with all other notations...)\\
+introduction section (with all other notations...)\\ [1ex]
 \serkan{\textsf{AR}:  We have added a clarification for this notation immediately after Theorem 2.1, which corresponds to the first occurrence of $\bf{H}'(s)$ .}
@@
 \item  Section 2.2 : the 5 introduction lines are very important and can be
-used to state the purpose of the paper.  \\
-\serkan{\textsf{AR}: We thank the referee for the comment.
-These 5 lines state the purpose of infinite horizon optimal model reduction.
-The purpose of the paper i.e.
-optimal model reduction on a finite horizon,  is stated in the introductory section.}
+used to state the purpose of the paper. \\ [1ex]
+\serkan{\textsf{AR}: These five lines state the purpose of infinite horizon optimal model reduction even though what we consider here, i.e., optimal model reduction on a finite horizon, has some similar difficulties.
+Following the referee's earlier suggestion (the first comment), we have added some more discussion about our approach to the introduction.}
 
 \item section 3 : the first 5 lines are a "useless repetition" and can be
-removed.
+removed. \\ [1ex]
+\serkan{\textsf{AR}:  We have removed these 5 lines and reorganized the beginning of section 3 }
 
-\serkan{\textsf{AR}:  We thank the referee for the comment. We have removed these 5 lines and reorganized the beginning of section 3 }
-
-\item Section 3 : lines 6 to 10 can be given as a Remark (important remark!)
-
-\serkan{\textsf{AR}:  We thank the referee for the comment and we have implemented this suggestion. }
+\item Section 3 : lines 6 to 10 can be given as a Remark (important remark!)  \\ [1ex]
+\serkan{\textsf{AR}:  We have followed  the referee's recommendation and list these lines as Remark 3.3 now. }
 
 \item Remark 3.2 can be given after the proof of the main result and can be
-a part of the discussion of the given result. \\
-\serkan{\textsf{AR}:  We have moved the remark after the proof. }
+a part of the discussion of the given result. \\[1ex]
+\serkan{\textsf{AR}:  We have moved the remark after the proof as the referee suggested. }
 
 \item Lemma 3.3 : "Let G(s) and Gr(s) be as defined in (3.5) and (3.7)"
-instead of "Let G(s) and Gr(s) be as defined in (3.5) and (3.5)"
-
+instead of "Let G(s) and Gr(s) be as defined in (3.5) and (3.5)" \\ [1ex]
 \serkan{\textsf{AR}:  We thank the referee for pointing out this issue. We have addressed it in the revised version.}
 
-\item Page 8 : (3.18) is one equation no need to (3.19).
+\item Page 8 : (3.18) is one equation no need to (3.19). \\ [1ex]
 \serkan{\textsf{AR}:  We thank the referee for pointing out this issue. We have addressed it in the revised version.}
@@
 \item The computation of (3.18) is not clear. One can give more details
-allowing to obtain such result. \\
-\serkan{\textsf{AR}:  We thank the referee for the comment. We have added more details explaining the computation.}
+allowing to obtain such result. \\ [1ex]
+\serkan{\textsf{AR}:   We have added more details after Equation (3.18) better explaining these computations.}
 
 \item Sentence after (3.19) : "...in the parentheses in (3.18) and (3.19)"
-instead of "...in the parentheses in (3.18) and (3.18)". \\
-
+instead of "...in the parentheses in (3.18) and (3.18)". \\[1ex]
 \serkan{\textsf{AR}:  We thank the referee for pointing out this issue. We have addressed it in the revised version.}
 
-\item 3.3 section can be a (concise) part of the introduction of the paper. \\
+\item 3.3 section can be a (concise) part of the introduction of the paper. \\[1ex]
 \serkan{\textsf{AR}:  We thank the referee for the comment.
@@
 the reduced order model which is not the case in the current version.\\
-\serkan{\textsf{AR}:  We thank the referee for the comment.
-We believe that the Corollaries 4.1 and 4.2, which follow directly from the main result, explain how we are using the optimality conditions to construct the algorithm.}
+\serkan{\textsf{AR}:  This comment was not very clear to us.
+Corollaries 4.1 and 4.2, which follow directly from the main result, explain how we are using the optimality conditions to construct the algorithm.
+We were not sure what else the referee would like to see explained.}
 
@@
 a clear figure (with only one comparison aim at each time). The given
-figures are almost indecipherable !\\
+figures are almost indecipherable !\\[1ex]
+\serkan{\textsf{AR}:  We thank the referee for pointing out the difficulties in reading the figures.
+We have not separated the figures into subplots, making the comparisons easier and hopefully making the figures more clear.
+}
 
-\serkan{\textsf{AR}:  We thank the referee for the comment. We have addressed the issue by providing one figure for each comparison.}
+\textcolor{blue}{
+We were not sure why the referee claimed that the proposed method outperformed the other methods.
+For every model and for every reduced order we have tried, FHIRKA has improved whatever initialization we have used.
+In some cases the improvements were dramatic in some cases less so.
+However regardless it has produced a lower error.
+We have excluded some $r$ values not because FHIRKA performed bad.
+It was simply because other methods has lead to very poor approximations that should not be used as an initialization.}
 
@@
 and Gr(s).  The authors can comment on the interpretation.\\
-\serkan{\textsf{AR}:  We thank the referee for the insightful comment. We have addressed this issue on the revised version.}
+\serkan{\textsf{AR}:  We thank the referee for this comment. We have addressed this issue on the revised version  and added Remark 3.2 to precisely address this issue.}
 
@@
 importance of the proposed method.\\
-\serkan{\textsf{AR}:  What model should we use for this?}
+\serkan{\textsf{AR}:  We have added a graph showing the time responses for the error for different methods; Figure 4 in the revised version.
+We have chosen to include the impulse error plot instead of the impulse responses to better illustrate the deviations.
+Due to the page limit stated by Automatica, we were able to include only one such figure.
+For a reference, we are including  two additional graphs here.}
+ \begin{figure}[H]
+ \centering
+   \includegraphics [scale=0.5]{absHeatErrorYlim}
+      \caption{POD vs \FH for the heat model\label{fig:impulseHeat}}
+ \end{figure}
+ \begin{figure}[H]
+ \centering
+   \includegraphics [scale=0.5]{absUNSError}
+      \caption{POD vs \FH for the unstable model\ \label{fig:impulseUns}}
+ \end{figure}
 

### 2019-05-13 FH Optimality Conditions Paper.tex
Latest SG Changes

@@
 %Simulation, design, and control of dynamical systems play an important role in numerous scientific and industrial tasks such as signal propagation in the nervous system\cite{kellems2009low}, heat dissipation, prediction of major weather events \cite{BoyP11,Sha03,Ant05}.
-Simulation, design, and control of dynamical systems play an important role in numerous scientific and industrial tasks such as \textcolor{red}{signal propagation in the nervous system\cite{kellems2009low};  the synthesis  of
+Simulation, design, and control of dynamical systems play an important role in numerous scientific and industrial tasks such as \textcolor{blue}{signal propagation in the nervous system\cite{kellems2009low};  the synthesis  of
 interconnect \cite{BonD07} and semiconductor devices \cite{Hess2014}; large-scale inverse problem,
@@
 
-\textcolor{red}{In this paper,  we explore optimal model reduction over a finite time horizon.
+\textcolor{blue}{In this paper,  we explore optimal model reduction over a finite time horizon.
 We use a time-limited version of the $\ch_2$ norm, denoted by $\ch_2(\tf)$, to quantify the model reduction error.}
@@
 
-\textcolor{red}{The advantage of the interpolation framework we will develop  is that 
+\textcolor{blue}{The advantage of the interpolation framework we will develop  is that 
 we do not require the reduced-model to be obtained via projection, as usually assumed in model reduction \cite{Ant05}.
@@
 \end{theorem}
-{\color{red}For more details on theorem \ref{h2thm}, see {\cite {GugBA08}}.}
+{\color{blue} where $\HH'(s)$ denotes the derivative of $\HH(s)$ with respect to $s$.}
+{\color{blue}For more details on Theorem \ref{h2thm}, see {\cite {GugBA08,AntBG10b}}.}
 This result states that the transfer function of the optimal $\ch_2$ approximation to $\HH(s)$ is a (tangential) Hermite interpolant where the interpolation points are the mirror images of the reduced-order poles $\{\lambda_k\}$, and the tangental directions are given by the corresponding residues $\{\lc_k\rc_k^T\}$.
-{\color{red} Throughout this paper, we denote the derivative of $\HH(s)$ by $\HH'(s)$.} Since the optimality conditions depend on the reduced-model to be computed, the solution requires a nonlinear iteration.
+Since the optimality conditions depend on the reduced-model to be computed, the solution requires a nonlinear iteration.
 The Iterative Rational Krylov Algorithm (IRKA)  \cite{GugBA08} and its variants such as \cite{beattie2012realization,gerstner2007hom,CasL18,beattie2009trust} use these interpolation based optimality conditions to produce   an interpolatory, locally $\ch_2$ optimal reduced model.
@@
 %
-{\color{red}
+{\color{blue}
 In this section, we present  the main theoretical results of the paper i.e., the interpolatory $\ch_2(\tf)$  optimality conditions, and discuss their implications.}
@@
 For a finite-time horizon 
-$[0,t_f]$, the $\ch_2(\tf)$\footnote{\textcolor{red}{Even though we are using an error measure specifically formulated in the time domain,   following \cite{RedK17,GoyR17} we choose to use  the $\ch_2(\tf)$ notation, despite the term $\ch_2$  is mostly associated with a measure defined in the frequency domain.
+$[0,t_f]$, the $\ch_2(\tf)$\footnote{\textcolor{blue}{Even though we are using an error measure specifically formulated in the time domain,   following \cite{RedK17,GoyR17} we choose to use  the $\ch_2(\tf)$ notation, despite the term $\ch_2$  is mostly associated with a measure defined in the frequency domain.
 Our main reason is to keep the connection to the infinite horizon problem where the measure in the frequency- and time-domains are equivalent.
@@
 \end{pf}%%%Start proof here
+
 \textbf{{PROOF OF THEOREM \ref{mh2opt}.}}
@@
 \end{dmath}
-{\color{red}
+{\color{blue}
 Note that the first term in \eqref{eqn:diffl} corresponds to the derivative of the second term in \eqref{merror} and the second term in \eqref{eqn:diffl} corresponds to the derivative of the last four terms in \eqref{merror}.}
@@
 \begin{flushright} $\Box$  \end{flushright}
-{\color{red} We note that the interval of interest is problem dependent and the choice of the interval i.e.
-the choice of $t_f$ may depend on the model, however, these optimality conditions hold for any choice of $t_f>0$.}
+{\color{blue} We note that the interval of interest is problem dependent and the choice of the interval, i.e., the choice of $t_f$, may depend on the model.
+However, these optimality conditions hold for any choice of $t_f>0$.}
 
@@
 \begin{remark}
-{\color{red} In the infinite-horizon case, if  $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
+{\color{blue} In the infinite-horizon case, if  $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
 the interpolant is $\G_r(s)$, and the interpolated function is $\G(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$.
-The interpolation of $\G(s)$ guarantees that we are approximating the model on the finite horizon.
-For instance, consider the inverse Laplace transform of $\G(s)$:
-\begin{align*}
-\mathcal{L}^{-1}\{\G(s)\}&=u_{\tf}\h(t)
-\end{align*}
-where $u_{\tf}(t)$ is the unit step function
-\begin{align*}
-u_{\tf}(t)=
- \begin{cases} 
-      1 &    t\leq {\tf} \\    
-      0 &    t > {\tf} 
-   \end{cases}.
-\end{align*}
-In other words, if we were to use an $\ch_2$ optimal algorithm on $\G(s)$, we would be approximating a finite horizon model by an infinite horizon one.
-For this reason, the interpolation of $\G(s)$ by $\G_r(s)$ is necessary.
-The optimal interpolation points  and directions are still determined by the reduced model $\HH_r(s)$. The algorithmic implications of this result will be revisited in Section \ref{sec:discuss}.}
+To give more intuition about these resulting interpolation conditions, considere the time-limited function $\g(t)$ such that
+$\g(t)=\h(t)$ when   $t\leq {\tf}$ and $\g(t)=0$ when when $t > {\tf}$.
+%\begin{align*}
+%\g(t)=
+% \begin{cases} 
+%      \h(t) &    t\leq {\tf}, \\    
+%      0 &    t > {\tf}.
+%   \end{cases}
+%\end{align*}
+A direct calculation shows that $\G(s)$ is the Laplace transformation of $\g(t)$. Similarly let $\g_r(t)$ 
+denote the time-limited version of $\h_r(t)$. Then its Laplace transformation is 
+ $\G_r(s)$.  Therefore, the optimality conditions  \eqref{eqn:cond1}--\eqref{eqn:cond3}
+ correspond to optimal interpolation of $\G(s)$ (Laplace transform of a time-limited function)
+ by $\G_r(s)$, (Laplace transform of another time limited function.) The fact that  
+ $\g(t)$ and   $\g_r(t)$ are both time-limited is the precise reason  why we cannot simply apply $\ch_2$ optimal reduction  to $\G(s)$.
+The method of \cite{beattie2012realization}, called TF-IRKA, does not require the original function to be a rational function.
+Thus, in principle we can use TF-IRKA to reduce $\G(s)$.
+However, the resulting reduced model is a rational function without any structure.
+In our case, the reduced model $\G_r(s)$ needs to retain the same structure as $\G(s)$ so that we can extract an $\HH_r(s)$.
+In other words, if we simply apply an $\ch_2$ optimal algorithm to $\G(s)$, we would be approximating a finite horizon model by an infinite horizon one and we cannot extract $\HH_r(s)$.
+Therefore, a new  algorithmic framework  is needed as we discuss in more detail in Section \ref{sec:discuss}.}
 \end{remark}
@@
 \begin{remark}
-{\color {red}
-We would like to point out that for an unstable dynamical system, if the matrix $\A$ does not have purely imaginary eigenvalues, one can work with the
+{\color{blue}
+For an unstable dynamical system without any purely imaginary eigenvalues, one can work with the
  $\mathcal{L}_2$ norm by decomposing it into a stable and anti-stable system, and then obtain an interpolatory reduced model based on this measure.
 However, this solution requires destroying the causality of the underlying dynamics \cite{magruder2010rational}.
-This is not the framework we are interested in here and we will work with a finite-time interval.
+This is not the framework we are interested in here and we  work with a finite-time interval.
 }
@@
 %\begin{remark}
-%{\color{red} In the infinite-horizon case, if $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
+%{\color{blue} In the infinite-horizon case, if $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
 %the interpolant is $\G_r(s)$, which is a function of $\HH_r(s)$, and the interpolated function is $\G(s)$, which is a function of $\HH(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$ in this case.
@@
 %%Proposition 
-%{\color{red}
+%{\color{blue}
 %The following proposition sheds some light on the physical interpretation of 
@@
 For some $r$ values, certain initializations are excluded (e.g., the \goyal \ initialization for $r=6$) since the algorithm either did not converge or produced  poor approximations.
-However, this happened rarely.
-For the ISS model (Figure \ref{fig:iss}), we use  TLBT and \goyal \ initializations since  POD produced very poor approximations and thus is excluded.
+However, this happened only rarely.
+
+   \begin{figure}[H] 
+ \centering
+ \includegraphics [scale=0.35]{fig1h} 
+ \caption{\FH \ and other algorithms for the heat model\label{fig:hm}}
+ \end{figure}
+
+   For the ISS model (Figure \ref{fig:iss}), we use  TLBT and \goyal \ initializations since  POD approximations was very poor and  is excluded.
 In this case, we reduce the order from $r=2$ to $r=14$ with increments of $2$.
@@
 Gains were significantly better for POD, especially for larger $r$ values.
-{\color{red} Also, in Figure \ref{fig:impulse}, we compare the impulse responses of the error systems of POD and \FH \ for the ISS model.
-For both methods, POD and \FH\  the reduced model was of order $r=14$.
-As we can see from the graph, \FH \ clearly outperforms POD on the time interval $[0, 1]$.
-
+{\color{blue} 
 }
@@
  \centering
- \includegraphics [scale=0.35]{fig1h} 
- \caption{\FH \ and other algorithms for a heat model\label{fig:hm}}
- \end{figure}
- \begin{figure}[H]
- \centering
   \includegraphics [scale=0.35]{fig2i}
@@
    \includegraphics [scale=0.35]{fig3u}
-      \caption{\FH \ and other algorithms for the ISS model \label{fig:uns}}
+      \caption{\FH \ and other algorithms for the unstable model \label{fig:uns}}
  \end{figure}
+
+\textcolor{blue}{Finally, in Figure \ref{fig:impulse}, we compare the error in the impulse responses due to POD and \FH \ for the ISS model.
+For both methods, POD and \FH\  the reduced model was of order $r=14$.
+As we can see from the graph, \FH \ clearly outperforms POD on the time interval $[0, 1]$.}
 
@@
    \includegraphics [scale=0.35]{ISS1RErrorSmooth}
-      \caption{\FH \ and other algorithms for the unstable model \label{fig:impulse}}
+      \caption{\FH \ and POD for the ISS 1R model \label{fig:impulse}}
  \end{figure}

### 2019-09-02 Final Submission/response.tex
final version
The folder named "Final Submission" contains the final version of the paper

--- 2019-05-13 response.tex
@@
 We thank the referees {for their detailed reading of the manuscript and their valuable comments}.
- We have incorporated the changes that the referees have suggested into the revised
+ We have incorporated all the changes that the referees have suggested into the revised
 manuscript.   We list below the referees' comments, followed by our responses, denoted by \serkan{\textsf{AR}}.
@@
 To clarify this point.
-we have added a footnote, Footnote 4 at the bottom of page 3, explaining the usage.
+we have added a footnote, namely Footnote 4, at the bottom of page 3, explaining the usage.
 } 
@@
 \serkan{\textsf{AR}:  We thank the referee for pointing this out.
-We have added some of this discussion from Section 3.3 to Introduction now (the second to last paragraph in Introduction to help resolve this issue.}
+We have added some of this discussion from Section 3.3 to Introduction now (the second to last paragraph in Introduction) to help resolve this issue.}
 
@@
 justification. \\[1ex]
-\serkan{\textsf{AR}:  We have added citations.}  
+\serkan{\textsf{AR}:  We have added citations as suggested.}  
 \item What about the choice of the finite horizon tf ? \\ [1ex]
-\serkan{\textsf{AR}: The choice of the finite horizon is really problem and application dependent; however the interpolation conditions hold for any $t_f$.
+\serkan{\textsf{AR}: The choice of the finite horizon is  problem and application dependent; however the interpolation conditions hold for any $t_f$.
 We added a sentence right before Remark 3.2 to state this fact.}
@@
 \item Justify the choice of simple poles for the reduced order model. \\ [1ex]
-\serkan{\textsf{AR}:  The choice of simple poles was motivated by the infinite horizon $\mathcal{H}_2$ optimal approximation problems and the corresponding interpolation, as in, for example, citations [3,6,7, 11,23] in the paper.
-Indeed, the interpolatory $\mathcal{H}_2$  approximation has been extended some specific nonlinear dynamical systems as well, such as
+\serkan{\textsf{AR}:  The choice of simple poles was motivated by the infinite horizon $\mathcal{H}_2$ optimal approximation problem and the corresponding interpolation conditions, as in, for example, citations [3,6,7, 11,23] in our manuscript.
+The interpolatory $\mathcal{H}_2$  approximation has been extended some specific nonlinear dynamical systems as well, such as
 \begin{itemize}
@@
 \end{itemize}
-These nonlinear extensions also use the same assumptions.
-It is certainly theoretically possible that an optimal $\mathcal{H}_2$ reduced model can have repeated poles and interpolation conditions have been extended to this case as well:
+These nonlinear extensions also use the same assumption.
+It is certainly theoretically possible that an optimal $\mathcal{H}_2$ reduced model can have repeated poles and interpolation conditions have been extended to this case:
 \begin{itemize}
@@
 \end{itemize}
-However, this is situation is rarely observed in practice and we chose to use simple poles. Also, we note that the full model can have repeated poles. The assumption is only on the reduced model.
+However, this situation is rarely observed in practice and for simplicity of the presentation and notation we chose to use simple poles.
+Also, we note that the full model can have repeated poles.
+The assumption is only for the reduced model.
 }
@@
 \item What is the reference of the existing result given by Theorem 2.1 ? \\ [1ex]
-\serkan{\textsf{AR}:  We have added the citations immediately after the theorem.}
+\serkan{\textsf{AR}:  We have added the citation immediately after the theorem.}
 
@@
 \serkan{\textsf{AR}: These five lines state the purpose of infinite horizon optimal model reduction even though what we consider here, i.e., optimal model reduction on a finite horizon, has some similar difficulties.
-Following the referee's earlier suggestion (the first comment), we have added some more discussion about our approach to the introduction.}
+Following the referee's earlier suggestion (the first comment), we have added some more discussion about our approach to  Introduction.}
 
@@
 removed. \\ [1ex]
-\serkan{\textsf{AR}:  We have removed these 5 lines and reorganized the beginning of section 3 }
+\serkan{\textsf{AR}:  We have removed these 5 lines and reorganized the beginning of Section 3 as suggested.}
 
@@
 \item 3.3 section can be a (concise) part of the introduction of the paper. \\[1ex]
-\serkan{\textsf{AR}:  We thank the referee for the comment.
-We believe that section 3.3 addresses the importance of the result.
-We believe that explaining the result after it is given is more appropriate and ti makes it more tangible for the reader.}
+\serkan{\textsf{AR}:  Following the referee's suggestion, we have added the basic points of Section 3.3 to Introduction.
+However, some of the discussions in  Section 3.3 depend on the main theoretical results of the paper and it is hard to discuss them in Introduction without these theoretical details.
+Therefore, as the referee suggested, our motivation to Introduction only highlight the main points of the analysis.}
 
@@
 \serkan{\textsf{AR}:  We thank the referee for pointing out the difficulties in reading the figures.
-We have not separated the figures into subplots, making the comparisons easier and hopefully making the figures more clear.
+We have now separated the figures into subplots, making the comparisons easier and hopefully making the figures more clear.
 }
@@
 \textcolor{blue}{
-We were not sure why the referee claimed that the proposed method outperformed the other methods.
-For every model and for every reduced order we have tried, FHIRKA has improved whatever initialization we have used.
+ For every model and for every reduced order we have tried, FHIRKA has improved whatever initialization method we have used.
+Even though we have not used the term ``supremacy" specifically, we believe it is fair to say that  FHIRKA outperformed other methods since this is what we have observed in every scenario we have tried.
 In some cases the improvements were dramatic in some cases less so.
@@
 We have excluded some $r$ values not because FHIRKA performed bad.
-It was simply because other methods has lead to very poor approximations that should not be used as an initialization.}
+It was simply because other methods has lead to such poor approximations that they should not be used as an initialization.}
 
@@
 and (3.7).  The reviewer cannot find a physical interpretation of G(s)
-and Gr(s).  The authors can comment on the interpretation.\\
-\serkan{\textsf{AR}:  We thank the referee for this comment. We have addressed this issue on the revised version  and added Remark 3.2 to precisely address this issue.}
+and Gr(s).  The authors can comment on the interpretation.\\[1ex]
+\serkan{\textsf{AR}:   We have addressed this issue on the revised version  and added Remark 3.2 to precisely address the referee's comment.}
 
@@
 responses of the models enables readers to understand "practical"
-importance of the proposed method.\\
+importance of the proposed method.\\[1ex]
 \serkan{\textsf{AR}:  We have added a graph showing the time responses for the error for different methods; Figure 4 in the revised version.
@@
    \includegraphics [scale=0.5]{absUNSError}
-      \caption{POD vs \FH for the unstable model\ \label{fig:impulseUns}}
+      \caption{POD vs\  \FH for the unstable model\ \label{fig:impulseUns}}
  \end{figure}
@@
 "G(s) and Gr(s) be as defined in (3.5) and (3.5)" in Lemma 3.3 should
-be corrected. \\
+be corrected. \\[1ex]
 \serkan{\textsf{AR}:  We thank the referee for pointing out this issue. We have addressed it in the revised version.}

### 2019-09-02 Final Submission/FH Optimality Conditions Paper.tex
final version
The folder named "Final Submission" contains the final version of the paper

--- 2019-05-13 FH Optimality Conditions Paper.tex
@@
 \end{frontmatter}
-%Introduction
-%Introduction
+
 \section{Introduction}
-%Simulation, design, and control of dynamical systems play an important role in numerous scientific and industrial tasks such as signal propagation in the nervous system\cite{kellems2009low}, heat dissipation, prediction of major weather events \cite{BoyP11,Sha03,Ant05}.
-Simulation, design, and control of dynamical systems play an important role in numerous scientific and industrial tasks such as \textcolor{blue}{signal propagation in the nervous system\cite{kellems2009low};  the synthesis  of
-interconnect \cite{BonD07} and semiconductor devices \cite{Hess2014}; large-scale inverse problem,
- \cite{Druskin2011solution,Lieberman2010,de2015nonlinear}; and prediction of major weather events \cite{Ant05}.}
+Simulation, design, and control of dynamical systems play an important role in numerous scientific and industrial tasks such as signal propagation in the nervous system\cite{kellems2009low};  the synthesis  of
+interconnect \cite{BonD07} and semiconductor devices \cite{Hess2014}; large-scale inverse problems,
+ \cite{Druskin2011solution,Lieberman2010,de2015nonlinear}; and prediction of major weather events \cite{Ant05}.
 The need for  detailed  models due to the increasing demand for greater resolution leads to  
@@
 Consider the linear time-invariant dynamical system:
-%\begin{equation}  \label{fom1}
-%\begin{array}{rcl}
-%\dot{\x}(t) & =& \A\x(t) +\B\uu(t)\\
-%\y(t)&=&\C\x(t) 
-%\end{array}
-%%~~\mbox{with}~~ \x(0) = \mathbf{0},
-%\; \Leftrightarrow \;
-%\y(t)=\int_{0}^{t}\h (t-\tau)\uu(\tau) d\tau,
-%\end{equation}
 \begin{equation}  \label{fom1}
@@
 \noindent
-where %$  \x(0) = \mathbf{0}, \A \in \R^{n\times n}$,
+where 
 $ \A \in \R^{n\times n}$,
@@
 reduction achieves this by replacing the original model with a lower dimensional one:
-%\begin{equation}  \label{rom1}
-%%\begin{array}{rcl}
-%\begin{aligned}
-%\dot{\x}_r(t)&=\A_r\x_r(t)+\B_r\uu(t)\\
-%\y_r(t)&=\C_r\x_r(t) 
-%\end{aligned}
-%%\end{array}
-%%~~\mbox{with}~~ \x_r(0) = \mathbf{0},
-%\; \Leftrightarrow \;
-%\y_r(t)=\int_{0}^{t}\h_r (t-\tau)\uu(\tau)d\tau,
-%\end{equation}
 \begin{equation}  \label{rom1}
@@
 \end{equation}
-where as in \eqref{fom1}, %$\x_r(0) = \mathbf{0}, \h_r(t)=\C_r e^{\A_r t}\B_r$
+where as in \eqref{fom1}, 
 $\h_r(t)=\C_r e^{\A_r t}\B_r$  is the impulse response of the reduced model, and $\A_r \in \R^{r\times r}$, $\B_r\in \R^{r\times m}$, and $\C_r\in \R^{p\times r}$ with $r\ll n$.
@@
 
-%In this paper,  we explore $\ch_2(\tf)$ optimality for model reduction over a finite time horizon. 
-%  We establish $\ch_2(\tf)$ optimality conditions over a finite time horizon and introduce an algorithm which yields better approximations of the large-scale dynamical systems compared to other model reduction methods.
-
-\textcolor{blue}{In this paper,  we explore optimal model reduction over a finite time horizon.
-We use a time-limited version of the $\ch_2$ norm, denoted by $\ch_2(\tf)$, to quantify the model reduction error.}
+In this paper,  we explore optimal model reduction over a finite time horizon. We use a time-limited version of the $\ch_2$ norm, denoted by $\ch_2(\tf)$, to quantify the model reduction error.
  Optimality requires a parametrization of the reduced model.
@@
 
-\textcolor{blue}{The advantage of the interpolation framework we will develop  is that 
+The advantage of the interpolation framework we will develop  is that 
 we do not require the reduced-model to be obtained via projection, as usually assumed in model reduction \cite{Ant05}.
-Indeed, one observation we will make is that unlike in the infinite-horizon $\ch_2$ approximation problem, the optimal reduced model in the finite-horizon norm is not necessarily given by a projection and thus a projection-based approach will not be able to satisfy the optimality conditions.
+Indeed, one observation we will make is that unlike in the infinite-horizon $\ch_2$ approximation problem, the optimal reduced model in the finite-horizon case is not necessarily given by a projection and thus a projection-based approach will not be able to satisfy the optimality conditions.
 Therefore,
-by treating the poles and residues in \eqref{kernel} as the parameters and directly working with them, we  obtain a reduced model to satisfy the optimality conditions exactly.}
-
-%The advantages of time-limited model reduction under the $\ch_2(\tf)$ norm are two fold: first, one can obtain a better approximation in the time interval of interest, and, second, one can obtain a locally optimal reduced order model in the case of unstable systems as well.
+by treating the poles and residues in \eqref{kernel} as the parameters and directly working with them, we  obtain a reduced model to satisfy the optimality conditions exactly.
 
@@
 \HH(-\lambda_k)\rc_k&=\HH_r(-\lambda_k)\rc_k,\\
-\lc_k^T\HH'(-\lambda_k)\rc_k&=\lc_k^T\HH_r'(-\lambda_k)\rc_k.
+\lc_k^T\HH'(-\lambda_k)\rc_k&=\lc_k^T\HH_r'(-\lambda_k)\rc_k,
 \end{aligned}
 \end{equation}
+
+where $\HH'(s)$ denotes the derivative of $\HH(s)$ with respect to $s$.
 \end{theorem}
-{\color{blue} where $\HH'(s)$ denotes the derivative of $\HH(s)$ with respect to $s$.}
-{\color{blue}For more details on Theorem \ref{h2thm}, see {\cite {GugBA08,AntBG10b}}.}
+For more details on Theorem \ref{h2thm}, see {\cite {GugBA08,AntBG10b}}.
 This result states that the transfer function of the optimal $\ch_2$ approximation to $\HH(s)$ is a (tangential) Hermite interpolant where the interpolation points are the mirror images of the reduced-order poles $\{\lambda_k\}$, and the tangental directions are given by the corresponding residues $\{\lc_k\rc_k^T\}$.
@@
 \section{$\ch_2(\tf)$ Optimal Model Reduction on a Finite Horizon}  \label{sec:main} 
-%If we are interested  in the behavior of a system only in a finite time interval (or access to a numerical simulation only on a finite time interval), we can obtain better reduced models by defining the measure of approximation quality over the interval of interest, instead of  focussing on the infinite time horizon.
-%
-{\color{blue}
-In this section, we present  the main theoretical results of the paper i.e., the interpolatory $\ch_2(\tf)$  optimality conditions, and discuss their implications.}
-
-%Also, in case of an unstable system, since the impulse grows unbounded as $t \to \infty$, the regular $\ch_2$ measure cannot be applied.
-Therefore, we will focus on model reduction on a finite time horizon.
+In this section, we present  the main theoretical results of the paper i.e., the interpolatory $\ch_2(\tf)$  optimality conditions, and discuss their implications.
 
@@
 For a finite-time horizon 
-$[0,t_f]$, the $\ch_2(\tf)$\footnote{\textcolor{blue}{Even though we are using an error measure specifically formulated in the time domain,   following \cite{RedK17,GoyR17} we choose to use  the $\ch_2(\tf)$ notation, despite the term $\ch_2$  is mostly associated with a measure defined in the frequency domain.
+$[0,t_f]$, the $\ch_2(\tf)$\footnote{Even though the term  $\ch_2$ is mostly associated with a measure in the frequency domain, following \cite{RedK17,GoyR17} we are using the notation $\ch_2(\tf)$ here as well to denote the error measure specifically formulated in the time domain.
 Our main reason is to keep the connection to the infinite horizon problem where the measure in the frequency- and time-domains are equivalent.
 And more importantly, as in the regular $\ch_2$ case, the optimality conditions will still appear as interpolation conditions in the frequency domain.
- %In the Since the time domain and frequency domain functions are equivalent,  this notation is appropriate.
-Furthermore, since the infinite horizon, and finite horizon case appear to be connected, the notation $\mathcal{H}_2(t_f)$ illustrates this connection.
-\cite{RedK17,GoyR17} use similar notation.
- }}
+}
    inner product $\langle \cdot, \cdot \rangle_{\ch_2(\tf)}$ 
@@
 
-%%Remark 3.2
-%\begin{remark}
-%Before we prove Theorem \ref{mh2opt}, we briefly comment on its implications.
-In the infinite-horizon case, if $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$.
-However, in the finite-horizon case,
-%the interpolant is $\G_r(s)$, which is a function of $\HH_r(s)$, and the interpolated function is $\G(s)$, which is a function of $\HH(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$ in this case.
-However, the optimal interpolation points  and directions are still determined by the reduced model $\HH_r(s)$.
-The algorithmic implications of this result will be revisited in Section \ref{sec:discuss}.
-%\end{remark}
 The next lemma will be used in proving Theorem \ref{mh2opt}.
@@
 \end{dmath}
-{\color{blue}
-Note that the first term in \eqref{eqn:diffl} corresponds to the derivative of the second term in \eqref{merror} and the second term in \eqref{eqn:diffl} corresponds to the derivative of the last four terms in \eqref{merror}.}
+Note that the first term in \eqref{eqn:diffl} corresponds to the derivative of the second term in \eqref{merror} and the second term in \eqref{eqn:diffl} corresponds to the derivative of the last four terms in \eqref{merror}.
 We rewrite \eqref{eqn:diffl} to obtain
@@
 
-%\begin{dmath}\label{eqn:l1} 
-%\pder[\J]{\lambda_k} =-2\lc_k^T \left(\sum_{j=1}^n \cc_j\bb_j^T \dfrac{({\tf}(\lambda_k+\rho_j)-1)e^{(\rho_j+\lambda_k){\tf}}+1}{(\lambda_k+\rho_j)^2}\right)\rc_k  
-%+2\lc_k^T \left(\sum^r_{i=1} \lc_i\rc_i^T\dfrac{({\tf}(\lambda_i+\lambda_j\k)-1)e^{(\lambda_i+\lambda_k){\tf}}+1}{(\lambda_i+\lambda_k)^2}\right)\rc_k.
-%\end{dmath}
-%\end{aligned}
-%\end{equation}
 Lemma \eqref{mgexp}, specifically \eqref{eqn:Gpatlk} and \eqref{eqn:Grpatlk}, show that the expressions in the parentheses in \eqref{eqn:l11}  and \eqref{eqn:l12}   are, respectively, $-\G'(-\lambda_k)$ and $-\G_r'(-\lambda_k)$.
@@
 \begin{flushright} $\Box$  \end{flushright}
-{\color{blue} We note that the interval of interest is problem dependent and the choice of the interval, i.e., the choice of $t_f$, may depend on the model.
-However, these optimality conditions hold for any choice of $t_f>0$.}
+We note that the interval of interest is problem dependent and the choice of the interval, i.e., the choice of $t_f$, may depend on the model.
+However, these optimality conditions hold for any choice of $t_f>0$.
 
@@
 \begin{remark}
-{\color{blue} In the infinite-horizon case, if  $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
+In the infinite-horizon case, if  $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
 the interpolant is $\G_r(s)$, and the interpolated function is $\G(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$.
-To give more intuition about these resulting interpolation conditions, considere the time-limited function $\g(t)$ such that
-$\g(t)=\h(t)$ when   $t\leq {\tf}$ and $\g(t)=0$ when when $t > {\tf}$.
-%\begin{align*}
-%\g(t)=
-% \begin{cases} 
-%      \h(t) &    t\leq {\tf}, \\    
-%      0 &    t > {\tf}.
-%   \end{cases}
-%\end{align*}
-A direct calculation shows that $\G(s)$ is the Laplace transformation of $\g(t)$. Similarly let $\g_r(t)$ 
-denote the time-limited version of $\h_r(t)$. Then its Laplace transformation is 
+To give more intuition about these resulting interpolation conditions, consider the time-limited function $\g(t)$ such that
+$\g(t)=\h(t)$ when   $t\leq {\tf}$ and $\g(t)=0$ when $t > {\tf}$.
+A direct calculation shows that $\G(s)$ is the Laplace transform of $\g(t)$. Similarly let $\g_r(t)$ 
+denote the time-limited version of $\h_r(t)$. Then its Laplace transform is 
  $\G_r(s)$.  Therefore, the optimality conditions  \eqref{eqn:cond1}--\eqref{eqn:cond3}
- correspond to optimal interpolation of $\G(s)$ (Laplace transform of a time-limited function)
- by $\G_r(s)$, (Laplace transform of another time limited function.) The fact that  
+ correspond to optimal interpolation of $\G(s)$ (Laplace transform of the time-limited function
+ $\g(t)$)
+ by $\G_r(s)$ (Laplace transform of the time limited function  $\g_r(t)$). The fact that  
  $\g(t)$ and   $\g_r(t)$ are both time-limited is the precise reason  why we cannot simply apply $\ch_2$ optimal reduction  to $\G(s)$.
@@
 In other words, if we simply apply an $\ch_2$ optimal algorithm to $\G(s)$, we would be approximating a finite horizon model by an infinite horizon one and we cannot extract $\HH_r(s)$.
-Therefore, a new  algorithmic framework  is needed as we discuss in more detail in Section \ref{sec:discuss}.}
+Therefore, a new  algorithmic framework  is needed as we discuss in more detail in Section \ref{sec:discuss}.
 \end{remark}
@@
 \begin{remark}
-{\color{blue}
 For an unstable dynamical system without any purely imaginary eigenvalues, one can work with the
@@
 This is not the framework we are interested in here and we  work with a finite-time interval.
-}
  \end{remark}
-
-%%remark
-%\begin{remark}
-%{\color{blue} In the infinite-horizon case, if $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
-%the interpolant is $\G_r(s)$, which is a function of $\HH_r(s)$, and the interpolated function is $\G(s)$, which is a function of $\HH(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$ in this case.
-However, the optimal interpolation points  and directions are still determined by the reduced model $\HH_r(s)$.
-The algorithmic implications of this result will be revisited in Section \ref{sec:discuss}.}
-%\end{remark}
-%
-%
-%%Proposition 
-%{\color{blue}
-%The following proposition sheds some light on the physical interpretation of 
-%\begin{Proposition}
-%Let  $\G(s)=-e^{-s{\tf}}\C(s\I-\A)^{-1}e^{\A {\tf}}\B+\HH(s)$ where $\HH(s)=\C(s\I-\A)^{-1}\B$  is the transfer function of \eqref{fom1}. Then, $\norm{\G}_{\ch_2({\tf})}=\norm{\HH}_{\ch_2({\tf})}$
-%\end{Proposition}
-%\begin{pf}
-%Note 
-%\begin{align*}
-%\G(s)&=-e^{-s{\tf}}\C(s\I-\A)^{-1}e^{\A {\tf}}\B+\HH(s)\\
-%&=-e^{-s{\tf}}\sum_{i=1}^n \dfrac{\phi_i e^{\lambda_i {\tf}}}{s-\lambda_i}+\HH(s)
-%\end{align*}
-%where $\lambda_i$-s and $\phi_i$-s are the poles and the residues of the system, respectively. 
-%Consider the inverse Laplace transform of $\G(s)$:
-%\begin{align*}
-%\mathcal{L}^{-1}\{\G(s)\}&=\h(t)+\sum_{i=1}^n u_{\tf}(t)\phi_i e^{\lambda_i {\tf}}e^{\lambda_i (t-{\tf})}\\
-%&=\h(t)+\sum_{i=1}^n u_{\tf}(t)\phi_i e^{\lambda_i t}
-%\end{align*}
-%where $u_{\tf}(t)$ is the unit step function
-%\begin{align*}
-%u_{\tf}(t)=
-% \begin{cases} 
-%      1 &    t\leq {\tf} \\    
-%      0 &    t > {\tf} 
-%   \end{cases}.
-%\end{align*}
-%Thus, 
-%\begin{align*}
-%\norm{\G}_{\ch_2({\tf})}&=\int_0^{\tf}\h(t)+\sum_{i=1}^n u_{\tf}(t)\phi_i e^{\lambda_i t} dt\\
-%&=\int_0^{\tf}\h(t)dt\\
-%&=\norm{\HH}_{\ch_2({\tf})}
-%\end{align*}
-%\end{pf}
-%This proposition and its proof suggest that we are matching the models up to $t_f$. For $t>t_f$, we treat the model as insignificant, and the impulse response vanishes after $t_f$.
-%}
-%
 
@@
 Gains were significantly better for POD, especially for larger $r$ values.
-{\color{blue} 
-}
-
-% \begin{figure}[H] 
-% \centering
-% \begin{subfigure}{.5\textwidth}
-% \centering
-% \includegraphics [scale=0.4]{allDataHeatF} 
-% \caption{\FH \ and other algorithms for a heat model\label{fig:hm}}
-% \end{subfigure}
-% \begin{subfigure}{.5\textwidth}
-% \centering
-% \includegraphics [scale=0.4]{ErrorVSrISSF}
-% \caption{\FH \ and other algorithms for the ISS model \label{fig:iss}}
-% \end{subfigure}
-%  \begin{subfigure}{.5\textwidth}
-% \centering
-%   \includegraphics [scale=0.4]{ErrorVSrUNSF}
-%      \caption{\FH \ and other algorithms for the unstable model \label{fig:uns}}
-% \end{subfigure}
-% \caption{Numerical results for three  models \label{mainfig}}
-% \end{figure}
-
- %This is the original figure
-%  \begin{figure}[H] 
-% \centering
-% \includegraphics [scale=0.4]{allDataHeatF} 
-% \caption{\FH \ and other algorithms for a heat model\label{fig:hm}}
-% \end{figure}
-% \begin{figure}[H]
-% \centering
-%  \includegraphics [scale=0.4]{ErrorVSrISSF}
-% \caption{\FH \ and other algorithms for the ISS model \label{fig:iss}}
-% \end{figure}
-%  \begin{figure}[H]
-% \centering
-%   \includegraphics [scale=0.4]{ErrorVSrUNSF}
-%      \caption{\FH \ and other algorithms for the unstable model \label{fig:uns}}
-% \end{figure}
-%
 
@@
   \includegraphics [scale=0.35]{fig2i}
- \caption{\FH \ and other algorithms for the heat model \label{fig:iss}}
+ \caption{\FH \ and other algorithms for the ISS model \label{fig:iss}}
  \end{figure}
@@
 
-\textcolor{blue}{Finally, in Figure \ref{fig:impulse}, we compare the error in the impulse responses due to POD and \FH \ for the ISS model.
+Finally, in Figure \ref{fig:impulse}, we compare the error in the impulse responses due to POD and \FH \ for the ISS model.
 For both methods, POD and \FH\  the reduced model was of order $r=14$.
-As we can see from the graph, \FH \ clearly outperforms POD on the time interval $[0, 1]$.}
+As we can see from the graph, \FH \ clearly outperforms POD on the time interval $[0, 1]$.
 
@@
    \includegraphics [scale=0.35]{ISS1RErrorSmooth}
-      \caption{\FH \ and POD for the ISS 1R model \label{fig:impulse}}
+      \caption{\FH \ and POD for the ISS model \label{fig:impulse}}
  \end{figure}
@@
 
-%\section*{References}
-%\bibliographystyle{elsarticle-num}
-
 \bibliographystyle{plain}

### 2019-09-02 Final Submission/FH Optimality Conditions PaperArxiv.tex
final version
The folder named "Final Submission" contains the final version of the paper

--- 2019-05-13 FH Optimality Conditions Paper.tex
@@
-%\documentclass[preprint]{elsarticle}
+%\documentclass[twocolumn][preprint]{elsarticle}
 \documentclass[twocolumn]{autart}  
@@
 %Simulation, design, and control of dynamical systems play an important role in numerous scientific and industrial tasks such as signal propagation in the nervous system\cite{kellems2009low}, heat dissipation, prediction of major weather events \cite{BoyP11,Sha03,Ant05}.
-Simulation, design, and control of dynamical systems play an important role in numerous scientific and industrial tasks such as \textcolor{blue}{signal propagation in the nervous system\cite{kellems2009low};  the synthesis  of
-interconnect \cite{BonD07} and semiconductor devices \cite{Hess2014}; large-scale inverse problem,
+Simulation, design, and control of dynamical systems play an important role in numerous scientific and industrial tasks such as \textcolor{black}{signal propagation in the nervous system\cite{kellems2009low};  the synthesis  of
+interconnect \cite{BonD07} and semiconductor devices \cite{Hess2014}; large-scale inverse problems,
  \cite{Druskin2011solution,Lieberman2010,de2015nonlinear}; and prediction of major weather events \cite{Ant05}.}
@@
 
-\textcolor{blue}{In this paper,  we explore optimal model reduction over a finite time horizon.
+\textcolor{black}{In this paper,  we explore optimal model reduction over a finite time horizon.
 We use a time-limited version of the $\ch_2$ norm, denoted by $\ch_2(\tf)$, to quantify the model reduction error.}
@@
 
-\textcolor{blue}{The advantage of the interpolation framework we will develop  is that 
+\textcolor{black}{The advantage of the interpolation framework we will develop  is that 
 we do not require the reduced-model to be obtained via projection, as usually assumed in model reduction \cite{Ant05}.
-Indeed, one observation we will make is that unlike in the infinite-horizon $\ch_2$ approximation problem, the optimal reduced model in the finite-horizon norm is not necessarily given by a projection and thus a projection-based approach will not be able to satisfy the optimality conditions.
+Indeed, one observation we will make is that unlike in the infinite-horizon $\ch_2$ approximation problem, the optimal reduced model in the finite-horizon case is not necessarily given by a projection and thus a projection-based approach will not be able to satisfy the optimality conditions.
 Therefore,
@@
 \HH(-\lambda_k)\rc_k&=\HH_r(-\lambda_k)\rc_k,\\
-\lc_k^T\HH'(-\lambda_k)\rc_k&=\lc_k^T\HH_r'(-\lambda_k)\rc_k.
+\lc_k^T\HH'(-\lambda_k)\rc_k&=\lc_k^T\HH_r'(-\lambda_k)\rc_k,
 \end{aligned}
 \end{equation}
+
+{\color{black} where $\HH'(s)$ denotes the derivative of $\HH(s)$ with respect to $s$.}
 \end{theorem}
-{\color{blue} where $\HH'(s)$ denotes the derivative of $\HH(s)$ with respect to $s$.}
-{\color{blue}For more details on Theorem \ref{h2thm}, see {\cite {GugBA08,AntBG10b}}.}
+{\color{black}For more details on Theorem \ref{h2thm}, see {\cite {GugBA08,AntBG10b}}.}
 This result states that the transfer function of the optimal $\ch_2$ approximation to $\HH(s)$ is a (tangential) Hermite interpolant where the interpolation points are the mirror images of the reduced-order poles $\{\lambda_k\}$, and the tangental directions are given by the corresponding residues $\{\lc_k\rc_k^T\}$.
@@
 %
-{\color{blue}
+{\color{black}
 In this section, we present  the main theoretical results of the paper i.e., the interpolatory $\ch_2(\tf)$  optimality conditions, and discuss their implications.}
@@
 For a finite-time horizon 
-$[0,t_f]$, the $\ch_2(\tf)$\footnote{\textcolor{blue}{Even though we are using an error measure specifically formulated in the time domain,   following \cite{RedK17,GoyR17} we choose to use  the $\ch_2(\tf)$ notation, despite the term $\ch_2$  is mostly associated with a measure defined in the frequency domain.
+$[0,t_f]$, the $\ch_2(\tf)$\footnote{\textcolor{black}{Even though the term  $\ch_2$ is mostly associated with a measure in the frequency domain, following \cite{RedK17,GoyR17} we are using the notation $\ch_2(\tf)$ here as well to denote the error measure specifically formulated in the time domain.
 Our main reason is to keep the connection to the infinite horizon problem where the measure in the frequency- and time-domains are equivalent.
@@
 \end{dmath}
-{\color{blue}
+{\color{black}
 Note that the first term in \eqref{eqn:diffl} corresponds to the derivative of the second term in \eqref{merror} and the second term in \eqref{eqn:diffl} corresponds to the derivative of the last four terms in \eqref{merror}.}
@@
 \begin{flushright} $\Box$  \end{flushright}
-{\color{blue} We note that the interval of interest is problem dependent and the choice of the interval, i.e., the choice of $t_f$, may depend on the model.
+{\color{black} We note that the interval of interest is problem dependent and the choice of the interval, i.e., the choice of $t_f$, may depend on the model.
 However, these optimality conditions hold for any choice of $t_f>0$.}
@@
 \begin{remark}
-{\color{blue} In the infinite-horizon case, if  $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
+{\color{black} In the infinite-horizon case, if  $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
 the interpolant is $\G_r(s)$, and the interpolated function is $\G(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$.
-To give more intuition about these resulting interpolation conditions, considere the time-limited function $\g(t)$ such that
-$\g(t)=\h(t)$ when   $t\leq {\tf}$ and $\g(t)=0$ when when $t > {\tf}$.
+To give more intuition about these resulting interpolation conditions, consider the time-limited function $\g(t)$ such that
+$\g(t)=\h(t)$ when   $t\leq {\tf}$ and $\g(t)=0$ when $t > {\tf}$.
 %\begin{align*}
@@
 %\end{align*}
-A direct calculation shows that $\G(s)$ is the Laplace transformation of $\g(t)$. Similarly let $\g_r(t)$ 
-denote the time-limited version of $\h_r(t)$. Then its Laplace transformation is 
+A direct calculation shows that $\G(s)$ is the Laplace transform of $\g(t)$. Similarly let $\g_r(t)$ 
+denote the time-limited version of $\h_r(t)$. Then its Laplace transform is 
  $\G_r(s)$.  Therefore, the optimality conditions  \eqref{eqn:cond1}--\eqref{eqn:cond3}
- correspond to optimal interpolation of $\G(s)$ (Laplace transform of a time-limited function)
- by $\G_r(s)$, (Laplace transform of another time limited function.) The fact that  
+ correspond to optimal interpolation of $\G(s)$ (Laplace transform of the time-limited function
+ $\g(t)$)
+ by $\G_r(s)$ (Laplace transform of the time limited function  $\g_r(t)$). The fact that  
  $\g(t)$ and   $\g_r(t)$ are both time-limited is the precise reason  why we cannot simply apply $\ch_2$ optimal reduction  to $\G(s)$.
@@
 \begin{remark}
-{\color{blue}
+{\color{black}
 For an unstable dynamical system without any purely imaginary eigenvalues, one can work with the
@@
 %\begin{remark}
-%{\color{blue} In the infinite-horizon case, if $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
+%{\color{black} In the infinite-horizon case, if $\HH_r(s)$ is the best $\ch_2$ approximation to $\HH(s)$, then $\HH_r(s)$ interpolates $\HH(s)$. However, in the finite-horizon case, 
 %the interpolant is $\G_r(s)$, which is a function of $\HH_r(s)$, and the interpolated function is $\G(s)$, which is a function of $\HH(s)$; thus $\HH_r(s)$ does not interpolate $\HH(s)$ in this case.
@@
 %%Proposition 
-%{\color{blue}
+%{\color{black}
 %The following proposition sheds some light on the physical interpretation of 
@@
 Gains were significantly better for POD, especially for larger $r$ values.
-{\color{blue} 
+{\color{black} 
 }
@@
 
-\textcolor{blue}{Finally, in Figure \ref{fig:impulse}, we compare the error in the impulse responses due to POD and \FH \ for the ISS model.
+\textcolor{black}{Finally, in Figure \ref{fig:impulse}, we compare the error in the impulse responses due to POD and \FH \ for the ISS model.
 For both methods, POD and \FH\  the reduced model was of order $r=14$.

### 2019-09-02 Final Submission/SinG19_Auto_CoverLetterRevision.tex
final version
The folder named "Final Submission" contains the final version of the paper

\documentclass[11pt]{letter}
\usepackage{vtmathlet}
%\input{vtmathlet.tex}
\usepackage{amsfonts}
\usepackage{color}
\usepackage{soul}
 \textheight 20.5cm
 \textwidth 16.0cm
% \topmargin -3.0cm
%\evensidemargin 1.0cm
 \oddsidemargin -0.1cm

 \begin{document}

\begin{letter} {Professor Richard Middleton\\
Editor (Systems and Control Theory) \\
Automatica
 }

 \signature{ 
 %Christopher Beattie, 
Klajdi Sinani \\
% and Volker Mehrmann
}

 \opening{Dear  Prof. Middleton:}

 Thank you very much for handling our manuscript ``$\mathcal{H}_2(t_f)$ Optimality Conditions for a Finite-time Horizon"  with submission number 18-0764, submitted to Automatica for a possible publication.

We have revised our manuscript following the referees' comments. In addition to the revised manuscript, we are also including a detailed response explaining our revisions and our answers to the referees' questions and comments.

We have uploaded the revised version, in which the revised parts are highlighted in blue so that the referees can easily spot the changes.

Moreover, following your suggestions, we put the paper in two-column format and now the length is around 10.5 pages.

Looking forward to hearing back from you.

\vspace{1.0cm}

\closing{Sincerely yours,}

\end{letter}

\end{document}
